### 2023-06-28 paper.qmd
First draft of paper sent to Julie & Margaux

---
title: "AdaptiveConformal: An R Package for Adaptive Conformal Inference"
subtitle: "AdaptiveConformal: An R Package for Adaptive Conformal Inference"
author:
  - name: Herbert Susmann
    corresponding: true
    email: herbert.susmann@dauphine.psl.eu
    url: https://herbsusmann.com
    orcid: 0000-0002-3540-8255
    #affiliations:
    #  - name: CNRS, Université Paris Dauphine - PSL
    #    url: https://cnrs.fr
    #  - name: PR[AI]RIE Institute
    #    url: https://prairie-institute.fr/
  - name: Margaux Zaffran
    email: margaux.zaffran@inria.fr
    url: https://mzaffran.github.io/
    orcid: 0000-0001-7560-8916
    #affiliations:
    #  - name: INRIA
    #    url: https://inria.fr
    #  - name: CMAP, Ecole Polytechnique
    #    url: https://inria.fr
    #  - name: EDF R\&D
    #    url: https://inria.fr
  - name: Antoine Chambaz
    email: antoine.chambaz@u-paris.fr
    url: https://helios2.mi.parisdescartes.fr/~chambaz/
    orcid: 0000-0002-5592-6471
    #affiliations:
    #  - name: Université Paris Cité
    #    department: MAP5
    #    url: https://map5.mi.parisdescartes.fr/
  - name: Julie Josse
    email: julie.josse@inria.fr
    url: http://juliejosse.com/
    orcid: 0000-0001-9547-891X
    #affiliations:
    #  - name: INRIA Sophia-Antipolis
    #    url: https://www.inria.fr/fr/centre-inria-universite-cote-azur
date: last-modified
date-modified: last-modified
description: |
  Conformal Inference is a popular approach for generating prediction intervals based on the output of any point prediction method. Adaptive Conformal Inference (ACI) algorithms extend classical approaches to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data. The common thread that unites algorithms in the ACI family is that they adaptively adjust the width of the generated prediction intervals in response to the observed data. We provide a detailed description of five ACI algorithms and their theoretical guarantees, and test their performance in simulation studies. We then present a case study of producing prediction intervals for influenza incidence in the United States based on black-box point forecasts. Implementations of all the algorithms are released as an open-source R package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
abstract: >+
  Conformal Inference is a popular approach for generating prediction intervals based on the output of any point prediction method. Adaptive Conformal Inference (ACI) algorithms extend classical approaches to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data. The common thread that unites algorithms in the ACI family is that they adaptively adjust the width of the generated prediction intervals in response to the observed data. We provide a detailed description of five ACI algorithms and their theoretical guarantees, and test their performance in simulation studies. We then present a case study of producing prediction intervals for influenza incidence in the United States based on black-box point forecasts. Implementations of all the algorithms are released as an open-source R package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
keywords: [Conformal inference, Adaptive conformal inference, time series, R]
citation:
  type: article-journal
  container-title: "Computo"
  doi: "xxxx"
  url: https://computo.sfds.asso.fr/template-computo-quarto
  publisher: "Société Française de Statistique"
  issn: "2824-7795"
bibliography: references.bib
github-user: computorg
repo: "template-computo-r"
draft: true # set to false once the build is running
published: false # will be set to true once accepted
format:
  computo-html: 
    code-fold: true
  computo-pdf: 
    include-in-header:
      - text: |
          \usepackage{stmaryrd}
---

# Introduction
Conformal inference methods are a family of algorithms for generating prediction intervals around point predictions [@shafer2008conformal, @angelopoulos2022gentle]. The input point predictions can be derived from any prediction method, making conformal inference a powerful tool for augmenting black-box prediction algorithms with prediction intervals. Classical conformal inference methods are able to yield marginally valid intervals with only the assumption that the observed data are _exchangeable_ (that is, the joint distribution of the data does not change based on the order of the observations).  However, in many real-world settings data are not exchangeable: for example, time series data usually can not be assumed to be exchangeable due to temporal dependence. A recent line of research examines the problem of generating prediction intervals for observations that are observed online (that is, one at a time) and for which exchangeability is not assumed to hold [@gibbs2021adaptive, @zaffran2022agaci, @gibbs2022faci, @bhatnagar2023saocp]. These methods, which we refer to generally as _Adaptive Conformal Inference_ (ACI) algorithms, work by adaptively adjusting the width of the generated prediction intervals in response to the observed data.

Informally, suppose a sequence of outcomes $y_t \in \mathbb{R}$, $t = 1, \dots, T$ are observed one at a time. Before seeing each observation, we have at our disposal a point prediction $\hat{\mu}_t \in \mathbb{R}$, which can come from any source. Our goal is to find an algorithm for producing prediction intervals $[\ell_t, u_t]$, $\ell_t \leq u_t$ such that, in the long run, the observations $y_t$ fall within the corresponding prediction intervals roughly $\alpha \times 100\%$ of the time. The original Adaptive Conformal Inference algorithm [@gibbs2021adaptive] is based on a simple idea: if the previous prediction interval at time $t-1$ did not cover the true observation, then the next prediction interval at time $t$ is made slightly wider. Conversely, if the previous prediction interval did include the observation, then the next prediction interval is made slightly narrower. It can be shown that, by choosing wisely how fast the intervals change size, this procedure yields prediction intervals that in the long run cover the true observations roughly the desired proportion of the time.

The main tuning parameter of the original ACI algorithm is a learning rate that controls how fast prediction interval width changes. If the learning rate is too low, then the prediction intervals will not be able to adapt fast enough to shifts in the data generating distribution; if it is too large, then the intervals will oscillate widely. The critical dependence of the original ACI algorithm on proper choice of the its learning rate spurred subsequent research into meta-algorithms that learn the correct learning rate (or an analogue thereof) in various ways, typically drawing on approaches from the online learning literature. In this paper, we present four such algorithms: Aggregated ACI [AgACI, @zaffran2022agaci], Fully Adaptive Conformal Inference [FACI, @gibbs2022faci], Scale-Free Online Gradient Descent [SF-OGD, @bhatnagar2023saocp], and Strongly Adaptive Online Conformal Prediction [SAOCP, @bhatnagar2023saocp].

Our primary practical contribution is an implementation of each algorithm in an open source R package, `AdaptiveConformal`, which is available at [https://github.com/herbps10/AdaptiveConformal](https://github.com/herbps10/AdaptiveConformal). The package also includes routines for visualization and summary of the prediction intervals. We note that Python versions of several algorithms were also made available by @zaffran2022agaci and @bhatnagar2023saocp, but to our knowledge this is the first R package to implement them.

The rest of the paper unfolds as follows. In @sec-theory, we present a unified theoretical framework for analyzing the ACI algorithms based on the online learning paradigm. In @sec-algorithms we provide descriptions of each algorithm along with their known theoretical properties. In @sec-simulations we compare the performance of the algorithms in several simulation studies. @sec-case-study gives a case study of generating prediction intervals for influenza incidence in the United States based on black-box point forecasts. Finally, @sec-discussion provides a discussion and ideas for future research in this rapidly expanding subfield.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(patchwork)
library(future)
library(furrr)
library(progressr)

# Contents of helpers.R included in the appendix
source("helpers.R")
```

# Theoretical Framework {#sec-theory}
*Notation*: for any integer $N \geq 1$ let $\llbracket N \rrbracket = \{ 1, \dots, N \}$. Let $\mathbb{I}$ be the indicator function. Let $\nabla f$ denote the subgradient of the function $f$.

We consider an online learning scenario in which we gain access to a sequence of observations $(y_t)_{t \geq 1}$ one at a time (see @cesabianchi2006games for an comprehensive account of online learning theory). Fix $\alpha \in (0, 1)$ to be the target empirical coverage of the prediction intervals. The goal is to output at time $t$ a prediction interval for the unseen observation $y_{t+1}$, with the prediction interval generated by an _interval construction function_ $\hat{C}_{t+1}$. Formally, let $\hat{C}_t$ be a function that takes as input a parameter $\theta_t \in \mathbb{R}$ and outputs a closed prediction interval $[\ell_t, u_t]$. The interval construction function must be nested: if $\theta^\prime > \theta$, then $\hat{C}_t(\theta) \subseteq \hat{C}_t(\theta^\prime)$. That is, larger values of $\theta$ imply wider prediction intervals. The interval constructor is indexed by $t$ to emphasize that it may use other information at each time point, such as a point prediction $\hat{\mu}_t \in \mathbb{R}$. We make no restrictions on how this external information is generated.

Define $r_t := \inf\{\theta \in \mathbb{R} : \mathbb{I}(y_t \in \hat{C}_t(\theta)) \}$ to be the _radius_ at time $t$. The radius is the smallest possible $\theta$ such that the prediction interval covers the observation $y_t$. A key assumption for the theoretical analysis of several of the algorithms is that the radii are bounded: that is, there exists a $D > 0$ such that $r_t < D$ for all $t$.

Next, we describe two possible definitions of the interval construction function.

## Linear Intervals
A simple method for forming the prediction intervals is to use the parameter $\theta_t$ to directly define the width of the interval. Suppose that at each time $t$ we have access to a point prediction $\hat{\mu}_t \in \mathbb{R}$. Then we can form a symmetric prediction interval around the point estimate as
$$
\begin{aligned}
  \hat{C}_t(\theta_t) := [\hat{\mu}_t - \theta_t, \hat{\mu}_t + \theta_t].
\end{aligned}
$$
We refer to this as the _linear interval constructor_. Note that in this case, the radius is simply the absolute residual $r_t = |\hat{\mu}_t - y_t|$. We focus primarily on this interval constructor in this paper, and it is the default used in the `Adaptive Conformal` R package. _Alternatively, we could also include results for all the algorithms run with conformal interval scores, perhaps in appendix_

## Conformal Intervals
The original ACI paper proposed constructing intervals based on the previously observed residuals [@gibbs2021adaptive]. Let $S(\hat{\mu}, y) : \mathbb{R}^2 \to \mathbb{R}$ be a function called a _conformity score_. A popular choice of conformity score is the absolute residual: $S(\hat{\mu}, y) := |y - \hat{\mu}|$.  Let $s_t := S(\hat{\mu}_t, y_t)$ be the conformity score of the $t$th-observation. The conformal interval construction function is then given by
$$
\begin{aligned}
  \hat{C}_t(\theta_t) := [\hat{\mu}_t - \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \}), \hat{\mu}_t + \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \})].
\end{aligned}
$$
where $\mathrm{Quantile}(\theta, A)$ denotes the empirical $\theta$-quantile of the elements in the set $A$. Note that $\hat{C}_t$ is indeed nested in $\theta_t$ because the Quantile function is non-decreasing in $\theta$.

The `AdaptiveConformal` package takes the absolute residual as the default conformity score, although the user may also specify any custom conformity score by supplying it as an R function.

## Online Learning Framework
We now introduce a loss function that defines the quality of a prediction interval with respect to a realized observation. Define the _pinball loss_ $L^\alpha$ as
$$
L^\alpha(\theta_t, r_t) := \begin{cases}
  (1-\alpha)(r_t - \theta), & r_t \leq \theta, \\
  \alpha(\theta - r_t), & \theta < r_t.
\end{cases}
$$
The way in which we observe the data and incur losses is as follows:

- For $t = 1, \dots, T$:
  - Predict radius $\theta_t$ and form prediction interval $\hat{C}_t(\theta_t)$.
  - Observe true outcome $y_t$ and calculate true radius $r_t$.
  - Record $\mathrm{err}_t := \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
  - Incur loss $L^\alpha(\theta_t, r_t)$.

There are two different perspectives we can take in measuring the quality of an ACI algorithm that generates a sequence $\theta_t$, $t \in \llbracket T \rrbracket$. First, we could look at how close the empirical coverage of the generated prediction intervals is to the desired coverage level $\alpha$. Formally, define the empirical coverage as the proportion of observations that fell within the corresponding prediction interval: $\mathrm{EmpCov}(T) := \frac{1}{T} \sum_{t=1}^T (1 - \mathrm{err}_t)$. The coverage error is then given by
$$
\begin{aligned}
  \mathrm{CovErr}(T) := \mathrm{EmpCov}(T) - \alpha.
\end{aligned}
$$
The second perspective is to look at how well the algorithm controls the incurred pinball losses. Following the classical framework from the online learning literature, we define the _regret_ as the difference between the cumulative loss yielded by a sequence $\theta_t$ versus the cumulative loss of the best possible fixed choice $\theta^*$:
$$
\begin{aligned}
  \mathrm{Reg}(T) := \sum_{t=1}^T L^\alpha(\theta_t, r_t) - \min_{\theta^* \in \mathbb{R}} \sum_{t=1}^T L^\alpha(\theta^*, r_t).
\end{aligned}
$$
In settings of distribution shift, it may not be appropriate to compare the cumulative loss of an algorithm to a fixed competitor. As such, stronger notions of regret have been defined. The _Strongly Adaptive Regret_ is the largest regret over any subperiod of length $k \in \llbracket T \rrbracket$:
$$
\begin{aligned}
  \mathrm{SAReg}(T, k) := \max_{[\tau, \tau + k - 1] \subseteq \llbracket T \rrbracket} \left( \sum_{t=\tau}^{\tau + k - 1} L^{\alpha}(\theta_t, r_t) - \inf_{\theta^*} \sum_{t=\tau}^{\tau + k - 1} L^\alpha(\theta^*, r_t) \right).
\end{aligned}
$$
Both ways of evaluating ACI methods are important because targeting only one or the other can lead to algorithms that yield prediction intervals that are not practically useful. As a simple pathological example of only targeting the coverage error, suppose we wish to generate $\alpha = 50\%$ prediction intervals. We could choose to alternate $\theta$ between 0 and $\infty$, such that $\mathrm{err}_t$ alternates between 0 and 1. The empirical coverage would then trivially converge to the desired level of 50\%. However, the same algorithm would yield infinite regret (see @bhatnagar2023saocp for a more in-depth example of an scenario in which coverage is optimal but the regret grows linearly). On the other hand, an algorithm that has arbitrarily small regret may not yield good empirical coverage. Suppose the observations and point predictions are constant: $y_t = 1$ and $\hat{\mu}_t = 0$ for all $t \geq 1$. Consider a simple class of algorithms that outputs constantly $\theta_t = \theta'$ for some $\theta' < 1$. With the linear interval construction function, the prediction intervals are then $\hat{C}_t(\theta_t) = [-\theta', \theta']$. The regret is given by $\mathrm{Reg}(T) = 2T\alpha(1-\theta')$, which approaches zero as $\theta'$ approaches 1. The empirical coverage is, however, always zero. In other words, the regret can be arbitrarily close to zero while at the same time the empirical coverage does not approach the desired level.

These simple examples illustrate that, unfortunately, bounds on the coverage error and bounds on the regret are not in general interchangeable. It is possible, however, to show equivalencies by either (1) making distributional assumptions on the data or (2) using additional information about how the algorithm produces the iterates $\theta_t$.

It may also be informative to summarize a set of prediction intervals in ways beyond their coverage error or by their regret. A common metric for prediction intervals is the _mean interval width_:
$$
\begin{aligned}
  \mathrm{MeanWidth}(T) := \frac{1}{T} \sum_{t=1}^T w_t,
\end{aligned}
$$
where $w_t := u_t - \ell-t$ is the interval width at time $t$. Finally, we introduce a metric that is intended to capture pathological behavior that can arise with ACI algorithms where the prediction intervals oscillate between being extremely narrow and extremely wide. Define the _path length_ of prediction intervals generated by an ACI algorithm as
$$
\begin{aligned}
  \mathrm{PathLength}(T) := \sum_{t=2}^T |w_t - w_{t-1}|.
\end{aligned}
$$
A high path length indicates that the prediction intervals were variable over time, and a low path length indicates the prediction intervals were stable.

# Algorithms {#sec-algorithms}
In this section we describe five ACI algorithms and summarize the theoretical guarantees that have been proven for each one. We have harmonized the algorithm descriptions to use a common structure and notation, in some cases departing from their original descriptions.

As a simple running example to illustrate each algorithm, we simulate $T = 500$ values $y_1, \dots, y_T$ following
$$
y_t \sim N(0, 0.2^2), \quad t \in \llbracket T \rrbracket.
$$
For demonstration purposes we assume to have access to unbiased predictions $\hat{\mu}_t = 0$ for all $t$.

```{r example_setup, echo = FALSE}
library(AdaptiveConformal)
set.seed(15321)
N <- 5e2
mu <- rep(0, N)
muhat <- mu
y <- rnorm(N, mean = mu, sd = 0.2)
```

## Adaptive Conformal Inference (ACI)

::: {#fig-algo-aci fig-align=left fig-pos=H}
```{.pseudocode}
\begin{algorithmic}
\State \textbf{Input:} starting value $\theta_1$, learning rate $\gamma > 0$.
\For{$t = 1, 2, \dots, T$}
  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$.
  \State Observe $y_t$.
  \State Set $\mathrm{err}_t = \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
  \State Update $\theta_{t+1} = \theta_t + \gamma (\mathrm{err}_t - (1 - \alpha))$.
\EndFor
\end{algorithmic}
```
Adaptive Conformal Inference algorithm [@gibbs2021adaptive].
:::

The original ACI algorithm (@fig-algo-aci) adaptively adjusts the width of the prediction intervals in response to the observations [@gibbs2021adaptive]. The updating rule for the estimated radius can be derived as an online subgradient descent scheme. The subgradient of the pinball loss function with respect to $\theta$ is given by
$$
\begin{aligned}
  \nabla L^\alpha(\theta_t, r_t) &= \begin{cases}
    \alpha - 1, & r_t < \theta, \\
    \alpha, &\theta < r_t, \\
    [\alpha - 1, \alpha], &\theta = r_t
  \end{cases} \\
\end{aligned}
$$
It follows that, for all $\theta_t \in \mathbb{R}$ and $r_t \in \mathbb{R}$,
$$
\alpha - 1 + \mathrm{err}_t \in \nabla L^\alpha(\theta_t, r_t).
$$
This leads to the following update rule for $\theta$ based on subgradient descent:
$$
\begin{aligned}
  \theta_{t+1} = \theta_{t} + \gamma (\mathrm{err}_t - (1 - \alpha)),
\end{aligned}
$$
where $\gamma > 0$ is a user-specified learning rate. For intuition, note that if $y_t$ fell outside of the prediction interval at time $t$ ($\mathrm{err}_t = 1$) then the next interval is widened ($\theta_{t+1} = \theta_t + \gamma \alpha$). On the other hand, if $y_t$ fell within the interval ($\mathrm{err}_t = 0$) then the next interval is shortened ($\theta_{t+1} = \theta_t - \gamma(1 - \alpha)$). The learning rate $\gamma$ controls how fast the width of the prediction intervals change in response to the data.

### Theoretical Guarantees
Standard results for online subgradient descent yield the following regret bound with the use of the linear interval constructor, assuming that the true radii are bounded by $D$:
$$
\mathrm{Reg}(T) \leq \mathcal{O}(D^2 / \gamma + \gamma T) \leq \mathcal{O}(D \sqrt{T}),
$$
where the second inequality follows if the optimal choice of $\gamma = D/\sqrt{T}$ is used [@bhatnagar2023saocp]. In addition, a finite sample bound on the coverage error can be established [@gibbs2021adaptive]:
$$
|\mathrm{CovErr}(T)| \leq \frac{D + \gamma}{\gamma T}.
$$

### Tuning Parameters
The only tuning parameter is the learning rate $\gamma$. Setting $\gamma$ too small will lead to intervals that do not adapt fast enough to distribution shifts, and setting $\gamma$ too large will lead to intervals with large oscillations as seen in @fig-aci. This is quantified in the path lengths, which increase significantly as $\gamma$ increases, even though the empirical coverage remains near the desired value. Thus, choosing a good value for $\gamma$ is essential. However, the optimal choice $\gamma = D / \sqrt{T}$ cannot be used directly in practice unless the time series length $T$ is fixed in advance, or the so called "doubling trick" is used to relax the need to know $T$ in advance [@cesabianchi2006games].

```{r aci_example_plot, echo = FALSE}
#| label: fig-aci
#| fig-height: 4.5
#| fig-cap: "Example prediction intervals from the ACI algorithm for different choices of learning rate $\\gamma$."
alpha <- 0.8
results <- list(
  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.01)),
  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.05)),
  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.1)),
  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.25))
)

coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))
mean_width <- scales::number_format(accuracy = 0.01)(unlist(lapply(results, function(result) result$metrics$mean_width)))

par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
for(i in 1:4) {
  plot(results[[i]], legend = FALSE, predictions = FALSE, cex = 0.5, main = bquote(gamma==.(results[[i]]$parameters$gamma)), ylim = c(-0.9, 0.8))
  text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
  text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[i]]) ), pos = 4)
}
par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
```

## Aggregated Adaptive Conformal Inference (AgACI)
::: {#fig-algo-agaci fig-align=left fig-pos=H}
```{.pseudocode}
\begin{algorithmic}
  \State \textbf{Input:} candidate learning rates $\{ \gamma_k \}_{1 \leq k \leq K }$, starting value $\theta_1$.
  \State Initialize BOA algorithms $\mathcal{B}^\ell$ and $\mathcal{B}^u$.
  \For{$k = 1, \dots, K$}
    \State Initialize ACI $\mathcal{A}_k = \texttt{ACI}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma_k, \theta_1 \leftarrow \theta_1)$.
  \EndFor
  \For{$t = 1, 2, \dots, T$}
    \For{$k = 1, \dots, K$}
      \State Generate candidate prediction interval $[\ell^k_{t}, u^k_{t}]$ from $\mathcal{A}_k$.
    \EndFor
    \State Compute aggregated lower and upper bounds $\tilde{\ell}_t$ and $\tilde{u}_t$ using $\mathcal{B}^\ell$ and $\mathcal{B}^u$.
    \State \textbf{Output} prediction interval $[\tilde{\ell}_t, \tilde{u}_t]$.
    \State Observe $y_t$.
    \For{$k = 1, \dots, K$}
      \State Update ACI $\mathcal{A}_k$ with $y_t$.
    \EndFor
    \State Update $\mathcal{B}^\ell$ with lower bounds $\ell^k_t$, $k = 1, \dots, K$ and observed outcome $y_t$. 
    \State Update $\mathcal{B}^u$ with upper bounds $u^k_t$, $k = 1, \dots, K$ and observed outcome $y_t$. 
  \EndFor
\end{algorithmic}
```
Aggregated Adaptive Conformal Inference (AgACI) algorithm [@zaffran2022agaci].
:::

The Aggregated ACI (AgACI; @fig-algo-agaci) algorithm solves the problem of choosing a learning rate for ACI by running multiple copies of the algorithm with different learning rates, and then combining their predictions using an online aggregation of experts algorithm [@zaffran2022agaci]. We follow the AgACI paper in using the Bernstein Online Aggregation (BOA) algorithm as implemented in the `opera` R package [@wintenberger2017boa; @opera2023].

### Theoretical Gaurantees
AgACI departs from our main theoretical framework in that it does not yield a set of iterates $\theta_1, \dots, \theta_T$ which are used to form prediction intervals using a set construction function $\hat{C}_t$. Rather, the upper and lower bounds from a set of candidate ACI algorithms are aggregated separately. Thus, theoretical results such as regret bounds in terms of $\theta$ similar to those for the other algorithms are not available. It would be possible, however, to establish regret bounds for the pinball loss applied separately to the lower and upper bounds of the prediction intervals. It's not clear, however, how to convert such regret bounds into a coverage bound.

### Tuning Parameters
The only tuning parameter for AgACI is the set of candidate learning rates. Beyond necessitating additional computational time, there isn't a drawback to having a large grid. As a basic check, we can also look at the weights assigned to each of the learning rates. If large weights are given to the smallest or largest learning rates, it is a sign that smaller (or larger) learning rates would perform well.

```{r agaci_example_plot, echo = FALSE}
#| label: fig-agaci
#| fig-height: 4.5
#| fig-cap: "Aggregated Adaptive Conformal Inference (AgACI) using the linear interval construction function and with initialization $\\theta_1 = 0$."
result <- aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear"))
coverage <- scales::percent_format()(result$metrics$coverage)
path_length <- scales::number_format(accuracy = 0.1)(result$metrics$path_length)
plot(result, ylim = c(-0.9, 0.8), legend = FALSE)
text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage) ), pos = 4)
text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length) ), pos = 4)
```

## Fully Adaptive Conformal Inference (FACI)
::: {#fig-algo-faci fig-align=left fig-pos=H}
```{.pseudocode}
\begin{algorithmic}
\State \textbf{Input:} starting value $\theta_1$, candidate learning rates $\{ \gamma_k \}_{1 \leq k \leq K }$, parameters $\sigma, \eta$.
\For{$k = 1, \dots, K$}
  \State Initialize expert $\mathcal{A}_k = \texttt{ACI}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma_k, \theta_1 \leftarrow \theta_1)$.
\EndFor
\For{$t = 1, 2, \dots, T$}
  \State Define $p_t^k := w_t^k / \sum_{i=1}^K w_t^i$, for all $1 \leq k \leq K$.
  \State Set $\theta_t = \sum_{k=1}^K \theta_t^k p_t^k$.
  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$.
  \State Observe $y_t$ and compute $r_t$.
  \State $\bar{w}_{t}^k \gets w_t^k \exp(-\eta L^\alpha(\theta_t^k, r_t))$, for all $1 \leq k \leq K$.
  \State $\bar{W}_t \gets \sum_{i=1}^K \bar{w}_t^i$.
  \State $w_{t+1}^k \gets (1 - \sigma) \bar{w}_t^k + \bar{W}_t \sigma / K$.
  \State Set $\mathrm{err}_t := \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
  \For{$k = 1, \dots, K$}
    \State Update ACI $\mathcal{A}_k$ with $y_t$ and obtain $\theta_{t+1}^k$.
  \EndFor
\EndFor
\end{algorithmic}
```
Fully Adaptive Conformal Inference (FACI) algorithm [@gibbs2022faci].
:::

The Fully Adaptive Conformal Inference method also aggregates predictions from multiple copies of ACI run with different learning rates, but differs from AgACI in that it aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci]. FACI uses the aggregation scheme proposed in [@gradu2022adaptive].

### Theoretical Guarantees
Let $\gamma_{\mathrm{max}} = \max_{1 \leq k \leq K} \gamma_k$ and assume that $\gamma_1 < \gamma_2 < \cdots < \gamma_K$ with $\gamma_{k+1}/\gamma \leq 2$ for all $1 \leq k < K$. Then, for any interval $I = [r, s] \subseteq \llbracket T \rrbracket$ and any sequence $\theta_r^*, \dots, \theta_s^*$, 
$$
\begin{aligned}
  \frac{1}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)] - \frac{1}{|I|} \sum_{t=r}^s \leq& \frac{\log(k / \sigma) + 2\sigma|I|}{\eta |I|} + \frac{\eta}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)^2] \\
  &+ 2\sqrt{3}(1 + \gamma_{\mathrm{max}})^2 \max\left\{ \sqrt{\frac{\sum_{t=r+1}^s |\beta_t^* - \beta_{t-1}^*}{|I|}}, \gamma_1 \right\},
\end{aligned}
$$
where the expectation is over the randomness in the randomized version of the algorithm. 
To convert the regret bound to a coverage bound, the authors took the path of making a distributional assumption on the radii $r_t$. Specifically, they assume that each of the conditional laws of $r_t | \{ r_s \}_{s < t}$ have a density that is bounded below by $p > 0$. For any interval $[s, r] := I \subseteq \llbracket T \rrbracket$ and any sequence $\theta_t^*$, $s \leq t \leq r$:
$$
\begin{aligned}
  \frac{1}{|I|} \sum_{t=r}^s \frac{p \mathbb{E}[(\theta_t - \theta^*_t)^2]}{2} \leq O\left( \sqrt{\frac{\log(|I|)}{|I|}} \right) + O\left( \sqrt{\frac{\sum_{t=r+1}^s |\theta_t^* - \theta_{t-1}^*|}{|I|}} \right).
\end{aligned}
$$
The term $\sum_{t=r+1}^s | \theta_t^* - \theta_{t-1}^*|$ is a path length providing a scalar summary of the degree of distribution shift in the data.

It has also been pointed out that FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]:
$$
\begin{aligned}
  \mathrm{SAReg}(T, k) \leq \widetilde{\mathcal{O}}(D^2 / \gamma + \gamma k).
\end{aligned}
$$
If $k$ is taken as fixed, then choosing $\gamma = D/\sqrt{k}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{k})$ (for a single choice of $k$).

### Tuning parameters
The parameter $|I|$, the time interval of interest, can be chosen arbitrarily. For the tuning parameter $\sigma$, the authors suggest the optimal choice $\sigma = 1 / (2 |I|)$. Choosing $\eta$ is more difficult. The authors suggest the following choice for $\eta$, which is optimal if there is no distribution shift:
$$
\begin{aligned}
  \eta = \sqrt{\frac{3}{|I|}} \sqrt{\frac{\log(K \cdot |I|) + 2}{(\alpha)^2 (1 - \alpha)^3 + (1-\alpha)^2 \alpha^3 }}
\end{aligned}.
$$
Note that this choice is optimal only for the conformal interval constructor, for which $\theta_t$ is a quantile of previous conformity scores. As an alternative, the authors point out that $\eta$ can be learned in an online fashion using the update rule
$$
\begin{aligned}
  \eta_t := \sqrt{\frac{\log(|I| K) + 2}{\sum_{s=t-|I|}^{t-1} L^\alpha(\theta_s, r_s)}}.
\end{aligned}
$$
Both ways of choosing $\eta$ led to very similar results in the original author's empirical studies. In the `AdaptiveConformal` package, the first approach is used when the conformal interval construction function is chosen, and the latter approach for the linear interval construction function.

```{r faci_example_plot, echo = FALSE}
#| label: fig-faci-linear
#| fig-height: 4.5
#| fig-cap: "Example prediction intervals generated by the FACI algorithm."
gamma_grid <- c(0.05, 0.25, 0.05)
result <- aci(y, muhat, alpha = alpha, method = "FACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "linear"))
coverage <- scales::percent_format()(result$metrics$coverage)
path_length <- scales::number_format(accuracy = 0.1)(result$metrics$path_length)

plot(result, ylim = c(-0.9, 0.8), legend = FALSE)
text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage) ), pos = 4)
text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length)), pos = 4)
```

## Scale-Free Online Gradient Descent (SF-OGD)
::: {#fig-algo-sfogd fig-align=left fig-pos=H}
```{.pseudocode}
\begin{algorithmic}
\State \textbf{Input:} starting value $\theta_1$, learning rate $\gamma > 0$.
\For{$t = 1, 2, \dots, T$}
  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$. 
  \State Observe $y_t$ and compute $r_t$. 
  \State Update $\theta_{t+1} = \theta_t - \gamma \frac{\nabla L^\alpha(\theta_t, r_t)}{\sqrt{\sum_{i=1}^t} \| \nabla L^\alpha(\theta_i, r_i) \|_2^2}$.
\EndFor
\end{algorithmic}
```
Scale-Free Online Gradient Descent (SF-OGD) algorithm [@orabona2018sfogd; @bhatnagar2023saocp].
:::

Scale-Free Online Gradient Descent is a general algorithm for online learning proposed by @orabona2018sfogd. The algorithm updates $\theta_t$ with a gradient descent step where the learning rate adapts to the scale of the observed gradients. SF-OGD was first used in the context of ACI as a sub-algorithm for SAOCP (described in the next section). However, it was found to have good performance by itself [@bhatnagar2023saocp] in real-world tasks, so we have made it available in the package as a stand-alone algorithm.

### Theoretical Guarantees
The SF-OGD algorithm has the following regret bound, which is called an _anytime regret bound_ because it holds for all $t \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
$$
\begin{aligned}
  \mathrm{Reg}(t) \leq \mathcal{O}(D \sqrt{t}) \text{ for all } t \in \llbracket T \rrbracket.
\end{aligned}
$$
A bound for the coverage error has also been established [@bhatnagar2023saocp]. For any learning rate $\gamma = \Theta(D)$ (where $\gamma = D / \sqrt{3}$ is optimal) and any initialization $\theta_1 \in [0, D]$, then it holds that for any $T > 1$,
$$
\begin{aligned}
  |\mathrm{CovErr}(T)| \leq \mathcal{O}\left( (1 - \alpha)^{-2} T^{-1/4} \log T \right).
\end{aligned}
$$

### Tuning parameters
[@fig-sf-ogd] compares results for several choices of $\gamma$ to illustrate its effect. The optimal choice of learning rate is $\gamma = D / \sqrt{3}$, where $D$ is the maximum possible radius. When $D$ is not known, it can be estimated by using an initial subset of the time series as a calibration set and estimating $D$ as the maximum of the absolute residuals of the observations and the predictions [@bhatnagar2023saocp].

```{r sfogd_example_plot, echo = FALSE}
#| label: fig-sf-ogd
#| fig-height: 4.5
#| fig-cap: "Example prediction intervals generated by the SF-OGD algorithm with different values of the maximum radius tuning parameter $D$."
alpha <- 0.8
results <- list(
  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.01)),
  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.1)),
  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.25)),
  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.5))
)

coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))

par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
for(i in 1:4) {
  plot(results[[i]], legend = FALSE, predictions = FALSE, cex = 0.5, main = bquote(gamma==.(results[[i]]$parameters$gamma)), ylim = c(-0.9, 0.8))
  text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
  text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[i]])), pos = 4)
}
par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
```

## Strongly Adaptive Online Conformal Prediction (SAOCP)
::: {#fig-algo-saocp fig-align=left fig-pos=H}
```{.pseudocode}
\begin{algorithmic}
\State \textbf{Input:} initial value $\theta_0$, learning rate $\gamma > 0$.
\For{$t = 1, 2, \dots, T$}
  \State Initialize expert $\mathcal{A}_t = \texttt{SF-OGD}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma, \theta_1 \leftarrow \theta_{t-1})$, set weight $w_t^t = 0$.
  \State Compute active set $\mathrm{Active}(t) = \{ i \in \llbracket T \rrbracket : t - L(i) < i \leq t \}$.
  \State Compute prior probability $\pi_i \propto i^{-2} (1 + \lfloor \log_2 i \rfloor )^{-1} \mathbb{I}[i \in \mathrm{Active}(t)]$.
  \State Compute un-normalized probability $\hat{p}_i = \pi_i [w_{t,i}]_+$ for all $i \in \llbracket t \rrbracket$.
  \State Normalize $p = \hat{p} / \| \hat{p} \|_1 \in \Delta^t$ if $\| \hat{p} \|_1 > 0$, else $p = \pi$.
  \State Set $\theta_t = \sum_{i \in \mathrm{Active}(t)} p_i \theta_t^i$ (for $t \geq 2$), and $\theta_t = 0$ for $t = 1$.
  \State \textbf{Output} prediction set $\hat{C}_t(\theta_t)$.
  \State Observe $y_t$ and compute $r_t$. 
  \For{$i \in \mathrm{Active}(t)$}
    \State Update expert $\mathcal{A}_t$ with $y_t$ and obtain $\theta_{t+1}^i$.
    \State Compute $g_t^i = \begin{cases}
      \frac{1}{D}\left(L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t)\right) & w_t^i > 0 \\
      \frac{1}{D}\left[L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t))\right]_+ & w_t^i \leq 0 \\
    \end{cases}$.
    \State Update expert weight $w_{t+1}^i = \frac{1}{t - i + 1}\left( \sum_{j=i}^t g_j^i \right) \left(1 + \sum_{j=i}^t w_j^i g_j^i \right)$.
  \EndFor
\EndFor
\end{algorithmic}
```
Strongly Adaptive Online Conformal Prediction (SAOCP) algorithm [@bhatnagar2023saocp]. 
:::

The Strongly Adaptive Online Conformal Prediction algorithm maintains a library of candidate online learning algorithms that generate prediction intervals which are then aggregated using a meta-algorithm [@bhatnagar2023saocp]. The candidate algorithm was chosen to be SF-OGD, although any algorithm that features anytime regret guarantees can be chosen. As opposed to AgACI and FACI, in which each candidate has a different learning rate, here each candidate has the same learning rate but is active over a different interval. At each time point, a new expert is instantiated which is active over a finite lifetime. Define the _lifetime_ of an expert instantiated at time $t$ as 
$$
\begin{aligned}
  L(t) := g \cdot \max_{n \in \mathbb{Z}} \{ 2^n t \equiv 0 \mod 2^n \},
\end{aligned}
$$
where $g \in \mathcal{Z}^*$ is a _lifetime multiplier_ parameter. The active experts are weighted according to their empirical performance with respect to the pinball loss function.

### Theoretical Guarantees
The following bound for the strongly adaptive regret holds for all lengths $k \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
$$
\begin{aligned}
  \mathrm{SAReg}(T, k) \leq 15 D \sqrt{k(\log T + 1)} \leq \tilde{\mathcal{O}}(D \sqrt k).
\end{aligned}
$$
It should be emphasized that this regret bounds holds simultaneously across all $k$, as opposed to FACI, where a similar bound holds only for a single $k$. A bound on the coverage error of SAOCP has also been established  as:
$$
\begin{aligned}
  |\mathrm{CovErr}(T)| \leq \mathcal{O}\left(\inf_\beta(T^{1/2 - \beta} + T^{\beta - 1} S_\beta(T))\right).
\end{aligned}
$$
where $S_{\beta}(T)$ is a technical measure of the smoothness of the cumulative gradients and expert weights for each of the candidate experts [@bhatnagar2023saocp].

### Tuning Parameters
The main tuning parameter for SAOCP is the learning rate $\gamma$ of the SF-OGD sub-algorithms, which we saw in the previous section has for optimal choice $\gamma = D / \sqrt{3}$. Values for $D$ that are too low lead to intervals that adapt slowly, and values that are too large lead to jagged intervals. In their experiments, the authors select a value for $D$ by picking the maximum residual from a calibration set. The second tuning parameter is the lifetime multiplier $g$ which controls the lifetime of each of the experts. We follow the original paper in setting $g = 8$.

```{r saocp_example_plot, echo = FALSE}
#| label: fig-saocp
#| fig-height: 4.5
#| fig-cap: "Example of prediction intervals generated by the SAOCP algorithm with different values of the maximum radius parameter $D$."
alpha <- 0.8
results <- list(
  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.01)),
  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.1)),
  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.25)),
  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.5))
)

coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))

par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
for(i in 1:4) {
  plot(results[[i]], legend = FALSE, predictions = FALSE, cex = 0.5, main = bquote(D==.(results[[i]]$parameters$D)), ylim = c(-0.9, 0.8))
  text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[i]])), pos = 4)
  text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[i]])), pos = 4)
}
par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
```

# Simulation Study {#sec-simulations}
We present two empirical studies in order to compare the performance of the ACI algorithms in simple simulated datasets. For both simulations we targeted empirical coverage of $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.

## Time series with ARMA errors
In this simulation we reproduce the setup described in @zaffran2022agaci (itself based on that of @friedman1983). The time series values $y_t$ for $t = 1, \dots, 600$ are simulated according to
$$
\begin{aligned}
  y_t = 10\sin(\pi X_{t,1}X_{t,2}) + 20(X_{t,3} - 0.5)^2 + 10X_{t,4} + 5 X_{t,5} + 0X_{t,6} + \epsilon_t,
\end{aligned}
$$
where $X_{t,i}$, $i = 1, \dots, 6$ are uniformly distributed on $[0, 1]$. The noise terms $\epsilon_t$ are generated according to an ARMA(1, 1) process:
$$
\begin{aligned}
  \epsilon_t &= \psi \epsilon_{t-1} + \xi_t + \theta \xi_{t-1}, \\
  \xi_t &\sim N(0, \sigma^2).
\end{aligned}
$$
We set $\psi$ and $\theta$ jointly to each value in $\{ 0.1, 0.8, 0.9, 0.95, 0.99 \}$ to simulate time series with increasing temporal dependence. The innovation variance was set to $\sigma^2 = (1 - \psi^2) / (1 + 2\psi \xi + \xi^2)$. For each setting, 25 simulated datasets were generated.

To provide point predictions for the ACI algorithms, at each time $t \geq 200$ a random forest model was fitted to the previously observed data using the `ranger` R package [@wright2017ranger]. The estimated model was then used to predict the subsequent time point. The maximum radius $D$ was estimated as the maximum residual observed between time points $t=200$ and $t=249$. The ACI models were then executed starting at time point $t = 250$. All metrics are based on time points $t \geq 300$ to allow time for the ACI methods to initialize.

```{r simulation_study_one, cache = TRUE}
simulate <- function(seed, psi, xi, N = 1e3) {
  set.seed(seed)

  s <- 10
  innov_scale <- sqrt(s * (1 - psi^2) / (1 + 2 * psi * xi + xi^2))

  X <- matrix(runif(6 * N), ncol = 6, nrow = N)
  colnames(X) <- paste0("X", 1:6)

  epsilon <- arima.sim(n = N, model = list(ar = psi, ma = xi), sd = innov_scale)

  mu <- 10 * sin(pi * X[,1] * X[,2]) + 20 * (X[,3] - 0.5)^2 + 10 * X[,4] + 5 * X[,5]
  y <- mu + epsilon
  as_tibble(X) %>% mutate(y = y)
}

estimate_model <- function(data, p = NULL) {
  if(!is.null(p)) p()
  preds <- numeric(nrow(data))
  for(t in 200:nrow(data)) {
    model <- ranger::ranger(y ~ X1 + X2 + X3 + X4 + X5 + X6, data = data[1:(t - 1),])
    preds[t] <- predict(model, data = data[t, ])$predictions
  }
  preds
}

metrics <- function(fit) {
  indices <- 300:length(fit$Y)
  aci_metrics(fit, indices)
}

fit <- function(data, preds, method, alpha, p = NULL) {
  if(!is.null(p)) p()

  D <- max(abs(data$y - preds)[200:249])
  gamma <- ifelse(method == "ACI", D / sqrt(length(250:nrow(data))), D / sqrt(3))

  parameters <- list(interval_constructor = "linear", D = D, gamma = gamma, gamma_grid = seq(0.1, 1, 0.1))
  aci(data$y[250:nrow(data)], preds[250:nrow(data)], method = method, alpha = alpha, parameters = parameters)
}

N_sims <- 25
simulation_data <- expand_grid(
  index = 1:N_sims,
  param =  c(0.1, 0.8, 0.9, 0.95, 0.99),
  N = 600
) %>%
  mutate(psi = param, xi = param)

# For each simulated dataset, fit multiple ACI methods
simulation_study_setup <- expand_grid(
  alpha = c(0.8, 0.9, 0.95),
  method = c("ACI", "AgACI", "SF-OGD", "SAOCP", "FACI")
)

simulation_study1 <- run_simulation_study1(
  simulation_data,
  simulation_study_setup,
  estimate_model,
  fit,
  workers = 8
)

```

The coverage error, mean interval width, and path lengths of each of the algorithms for $\alpha = 0.9$ are shown in @fig-simulation-one-results (results for $\alpha \in \{ 0.8, 0.95 \}$ were similar and are available in the appendix). All methods achieved near optimal empirical coverage, although SAOCP tended to slightly undercover. The mean interval widths were similar across methods, although again SAOCP had slightly shorter intervals (as expected given its tendency to under cover). The path length of SAOCP was larger than any of the other methods. To investigate why, @fig-simulation-one-example plots $w_t - w_{t-1}$, the difference in interval width between times $t-1$ and $t$, for each method in one of the simulations. The interval widths for ACI, AgACI, and FACI change slowly relative to those for SF-OGD and SAOCP. For SAOCP, we can see the interval widths have larger fluctuations than the other methods, explaining its higher path width.

```{r simulation_one_plot, message=FALSE, warning=FALSE}
#| label: fig-simulation-one-results
#| fig-height: 5.5
#| fig-cap: "Coverage errors, mean interval widths, and path lengths for the first simulation study with target coverage $\\alpha = 0.9$."

simulation_one_plot(simulation_study1$results %>% filter(alpha == 0.9))
```

```{r simulation_one_example_plot}
#| label: fig-simulation-one-example
#| fig-height: 6
#| fig-cap: "Difference in successive interval widths ($w_t - w_{t-1}$) from an illustrative simulation from the first simulation study."
fits <- simulation_study1$example_fits

par(mfrow = c(3, 2), mar = c(3, 4, 2, 1))
for(i in 1:5) {
  plot(
    diff(fits$fit[[i]]$intervals[,2] - fits$fit[[i]]$intervals[,1]), 
    main = fits$method[[i]], 
    xlab = "T", 
    ylab = expression(w[t] - w[t - 1]))
}
par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
```

## Distribution shift
This simulation study features time series that feature distribution shifts. The setup is quite simple in order to probe the basic performance of the methods in response to distribution shift. As a baseline, we simulate time series according to
$$
\begin{aligned}
  y_t &\sim N(0, \sigma_t^2), \\
  \sigma_t &= 0.2,
\end{aligned}
$$
for $t = 1, \dots, T$ and $T = 500$. In the second type of time series, the variance of the observations increases halfway through the time series:
$$
\begin{aligned}
y_t &\sim N(0, \sigma_t^2), \\
\sigma_t^2 &= 0.2 + 0.5 \mathbb{I}[t > 250]
\end{aligned}
$$
In each case, the ACI algorithms are provided with the unbiased predictions $\hat{\mu}_t = 0$, $t = 1, \dots, T$. 50 simulated datasets were generated for each type of time series.

```{r simulation_two, cache = TRUE}
simulate <- function(seed, distribution_shift = 0, N = 1e3, sigma = 0.2) {
  set.seed(seed)
  mu <- rep(0, N)
  shift <- 1:N > (N / 2)
  muhat <- mu
  y <- rnorm(n = length(mu), mean = mu, sd = sigma + ifelse(shift, distribution_shift, 0))

  tibble(y = y, muhat = muhat)
}

metrics <- function(fit) {
  N <- length(fit$Y)
  indices <- which(1:N > 50)
  aci_metrics(fit, indices)
}

fit <- function(data, method, alpha, p = NULL) {
  if(!is.null(p)) p()
  D <- max(abs(data$y - data$muhat)[1:50])

  if(method == "ACI") {
    gamma <- D / sqrt(nrow(data))
  }
  else {
    gamma <- D / sqrt(3)
  }

  parameters <- list(
    interval_constructor = "linear", 
    D = D, 
    gamma = gamma, 
    gamma_grid = seq(0.1, 2, 0.1)
  )

  aci(data$y, data$muhat, method = method, alpha = alpha, parameters = parameters)
}

N_sims <- 5e1
simulation_study_setup2 <- expand_grid(
  index = 1:N_sims,
  distribution_shift = c(0, 0.5),
  alpha = c(0.8, 0.9, 0.95),
  N = 500,
  method = c("ACI", "AgACI", "SF-OGD", "SAOCP", "FACI")
) %>%
  mutate(data = pmap(list(index, distribution_shift, N), simulate))

simulation_study2 <- run_simulation_study2(simulation_study_setup2, fit, workers = 8)
```

The coverage error, mean interval widths, and path lengths of each of the algorithms are shown in @fig-simulation-two-results. The coverage error of all the algorithms is near the desired value in the absence of distribution shift. On the contrary, all of the algorithms with the exception of ACI undercover when there is distributional shift. An illustrative example of prediction intervals generated by each ACI method for one of the simulated time series with distribution shift is shown in @fig-simulation-two-example.

```{r simulation_two_plot}
#| label: fig-simulation-two-results
#| fig-height: 9
#| fig-cap: "Coverage error, mean interval width, and path length for $\\alpha = 0.8, 0.9, 0.95$ and simulations with and without distributional shift."

simulation_two_plot(simulation_study2$results)
```

```{r simulation_two_example_plot}
#| label: fig-simulation-two-example
#| fig-height: 6
#| fig-cap: "Example prediction intervals (target coverage $\\alpha = 0.9$) from the second simulation study of time series with distributional shift."
fits <- simulation_study2$example_fits

coverage <- scales::percent_format()(unlist(lapply(fits$fit, function(result) result$metrics$coverage)))
path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(fits$fit, function(result) result$metrics$path_length)))

par(mfrow = c(3, 2), mar = c(3, 3, 2, 1))
for(i in 1:5) {
  plot(fits$fit[[i]], legend = FALSE, main = fits$method[[i]], index = 51:500)
  text(x = -10, y = -1.5, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
  text(x = -10, y = -2, labels = bquote(PathLength == .(path_length[[i]]) ), pos = 4)
}
par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
```

# Case Study: Influenza Forecasting {#sec-case-study}
Influenza is a highly infectious disease that is estimated to infect ~1 billion individuals each year around the world [@krammer2018influenza]. Influenza incidence in temperate climates tends to follow a seasonal pattern, with the highest number of infections during what is commonly referred to as the \textit{flu season} [@lofgren2007influenza]. Accurate forecasting of influenza is of significant interest to aid in public health planning and resource allocation. To investigate the accuracy of influenza forecasts, The US Centers for Disease Control (CDC) initiated a challenge, referred to as FluSight, in which teams from multiple institutions submitted weekly forecasts of influenza incidence [@biggerstaff2016flusight]. @reich2019influenza evaluated the accuracy of the forecasts over seven flu seasons from 2010 to 2017. As a case study, we investigate the use of ACI algorithms to augment the FluSight forecasts with prediction intervals.

The FluSight challenge collected forecasts for multiple prediction targets. For this case study, we focus on national (US) one-week ahead forecasts of weighted influenza-like illness (wILI), which is a population-weighted percentage of doctors visits where patients presented with influenza-like symptoms [@biggerstaff2016flusight]. The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, including both mechanistic and statistical approaches [@tushar2018flusightnetwork]. For our purposes, we treat the way the forecasts were produced as a black box.

Formally, let $y_{t}$, $t \in \llbracket T \rrbracket$ be the observed national wILI at time $t$, and let $\hat{\mu}_{j,t}$, $j = \llbracket J \rrbracket$ be the one-week ahead forecast of the wILI from model $j$ at time $t$. Two of the original 21 forecasting methods were excluded from this case study due to poor predictive performance (\texttt{Delphi\_Uniform} and \texttt{CUBMA}). The ACI methods were then applied to the log-observations and log-predictions, where the log-transformation was used to constrain the final prediction intervals to be positive. The first flu season (2010-2011) was used as a warm-up for each ACI method, and we report the empirical performance of the prediction intervals for the subsequent seasons (six seasons from 2012-2013 to 2016-2017). The ACI algorithms target prediction intervals with coverage of $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.

```{r case_study, cache = TRUE}
raw_data <- read_csv("data/point_ests.csv", show_col_types = FALSE)

fit <- function(data, method, alpha) {
  first_season <- data$Season == "2010/2011"
  D <- max(abs(data$obs_value - data$Value)[first_season])

  gamma <- ifelse(method == "ACI", D / sqrt(nrow(data)), D / sqrt(3))

  parameters <- list(
    interval_constructor = "linear", 
    D = D, 
    gamma = gamma, 
    gamma_grid = seq(0.01, 1, 0.05)
  )

  aci(
    Y = log(data$obs_value), 
    predictions = log(data$Value), 
    method = method, 
    parameters = parameters, 
    alpha = alpha
  )
}

metrics <- function(data, fit) {
  aci_metrics(fit, indices = which(data$Season != "2010/2011"))
}

analysis_data <- raw_data %>%
  filter(
    Target == "1 wk ahead", 
    Location == "US National", 
    !(model_name %in% c("Delphi_Uniform", "CUBMA"))
  ) %>%
  arrange(Year, Model.Week) %>%
  group_by(model_name) %>%
  nest()

fits <- expand_grid(
  analysis_data, 
  tibble(method = c("ACI", "AgACI", "FACI", "SF-OGD", "SAOCP")), 
  tibble(alpha = c(0.8, 0.9, 0.95))
) %>%
  mutate(fit = pmap(list(data, method, alpha), fit),
         metrics = map2(data, fit, metrics))

case_study_results <- fits %>%
  select(-data, -fit) %>%
  mutate(metrics = map(metrics, as_tibble)) %>%
  unnest(c(metrics)) 
```

The coverage errors, mean interval widths, and path lengths of the prediction intervals for each of the underlying forecast models is shown in @fig-case-study-metrics. In almost all cases the absolute coverage error was less than $0.1$. SF-OGD performed particularly well, with coverage errors close to zero for all forecasting models. Interval widths were similar across methods, with SAOCP slightly shorter. Path Lengths were shorter for ACI and SF-OGD and longer for SAOCP.

```{r case_study_metrics_plot}
#| label: fig-case-study-metrics
#| fig-height: 6
#| fig-cap: "Coverage error, mean interval width, and path lengths of prediction intervals generated with each ACI method based on forecasts from each of the 19 underlying influenza forecasting models."

case_study_plot(case_study_results)
```

As an illustrative example, in @fig-case-study-example we plot the point forecasts from one of the forecasting models (based on SARIMA with no seasonal differencing) and the associated ACI-generated 90\% prediction intervals for each season from 2011-2017. Interestingly, the forecasts in 2011-2012 underpredicted the observations for much of the season. The algorithm responds by making the intervals wider to cover the observations, and because the intervals are symmetric the lower bound then becomes unrealistically low. A similar phenomenon can be seen in the growth phase of the 2012/2013 season as well.

```{r case_study_example_plot}
#| label: fig-case-study-example
#| fig-height: 4
#| fig-cap: "Example conformal prediction intervals for six flu seasons based on forecasts from a SARIMA type model."

sarima_fits <- fits %>% filter(
  model_name == "ReichLab_sarima_seasonal_difference_FALSE", 
  alpha == 0.9
) %>%
  mutate(output = map(fit, extract_intervals)) %>%
  select(method, alpha, data, output) %>%
  unnest(c(data, output)) %>%
  filter(Season != "2010/2011")

sarima_fits %>%
  ggplot(aes(x = Model.Week, y = log(obs_value))) +
  geom_point(aes(shape = "Observed")) +
  geom_line(aes(y = pred, lty = "Forecast"), color = "black") +
  geom_line(aes(y = lower, color = method)) +
  geom_line(aes(y = upper, color = method)) +
  facet_wrap(~Season) +
  labs(
    x = "Flu Season Week", 
    y = "log(wILI)", 
    title = "SARIMA forecasts with ACI 90% prediction intervals"
  )
```

# Discussion {#sec-discussion}
The results of our simulations and case study show that, when tuning parameters are chosen well, Adaptive Conformal Inference algorithms yield well-performing prediction intervals. On the other hand, poor choice of tuning parameters can lead to intervals of very low utility. Furthermore, in some cases the prediction intervals may appear to perform well with respect to metrics like the empirical coverage error, while simultaneously being practically useless. The original ACI algorithm illustrates this phenomena: too small a value of its learning rate $\gamma$ yields prediction intervals that are not reactive enough, and too large yields intervals that change too fast. In both cases, the empirical coverage may appear well-calibrated, while the prediction intervals will not be useful. Thus, the core challenge in designing an ACI algorithm is in finding an optimal level of reactivity for the prediction intervals. As users of these algorithms, the challenge is in finding values for the tuning parametesr that avoid pathological behaviors.

Several of the algorithms investigated in this paper handle the problem of finding an optimal level of reactivity by aggregating prediction intervals generated by a set of underlying ACI algorithms. 
Our results show the algorithms can perform well in multiple difficult scenarios. However, the overall effect of these approaches is to shift the problem to a higher level of abstraction: we still need to set tuning parameters that control the amount of reactivity, we just do so at a higher level than the original ACI algorithm. It is desirable that these tuning parameters be easily interpretable, with simple strategies available for setting them.  An advantage of the SF-OGD and SAOCP algorithms in this respect are that their main tuning parameter, the maximum radius $D$, is easily interpretable as the maximum possible difference between the input predictions and the truth. It is also straightforward to choose this parameter based on a calibration set, although this strategy does not necessarily work well in cases of distribution shift. We also found that an advantage of the AgACI method is its robustness to the choice of its main tuning parameter, the set of candidate learning rates. Indeed, if AgACI does not perform well, one can simply increase the number of candidate learning rates.

A key challenge in tuning the algorithms arises in settings of distribution shift, where methods for choosing hyperparameters based on a calibration set from before the distribution shift will likely not perform well. The second simulation study we conducted probed this setting in a simple scenario. We found that several of the methods yielded prediction intervals that had non-optimal empirical coverage. As we picked hyperparameters based on a calibration set formed before the distribution shift, it is not surprising that the resulting tuning parameters are not optimal. This underscores the difficulty in designing  ACI algorithms that can adapt to distribution shifts, and in finding robust methods for choosing hyperparameters. In practice, it is possible the second simulation study does not accurately reflect real-world scenarios. Indeed, the benchmarks presented in [@bhatnagar2023saocp] using the datasets from the M4 competition, and using point predictions generated by diverse prediction algorithms, found that ACI algorithms exhibited good performance in terms of empirical coverage.  Regardless, our recommendation for future papers in this line of research is to include simulation studies for simple distributional shift scenarios as a benchmark.

Our case study results illustrate the dependence of the ACI algorithms on having access to high-quality point predictions. If the predictions are biased, for example, then the prediction intervals may be able to achieve optimal coverage at the expense of larger interval widths. Using ensemble methods to combine forecasts from several flexible machine-learning models is one strategy that can be used to hedge against model misspecification and improve the quality of forecasts.

There remain many possible extensions of ACI algorithms. The algorithms presented in this work primarily consider symmetric intervals evaluated using the pinball loss function (AgACI can yield asymmetric intervals because the aggregation rule is applied separately to the lower and upper bounds from the underlying experts, but those underlying experts only produce symmetric intervals). A simple extension would switch to using the interval loss function, which would allow for asymmetric intervals where two parameters are learned for the upper and lower bounds, respectively. It may also be of interest to generate prediction intervals that have regret bounds and coverage guarantees for arbitrary subsets of observations (for example, we may seek prediction intervals that have near optimal coverage for every day of the week, or month of the year). Another avenue of theoretical research is to find algorithms with provable bounds for the coverage and regret that do not depend on the outcome being bounded.

_Final wrap up paragraph?_

# References {.unnumbered}

::: {#refs}
:::

# Appendix

```{r simulation_one_plot_appendix}
#| fig-cap: "Coverage errors, mean interval widths, and path lengths for the first simulation study with target coverage $\\alpha \\in \\{ 0.8, 0.9, 0.95 \\}$."
#| fig-height: 9 
simulation_one_plot(simulation_study1$results)
```

## Additional R code
```{r}
```

### 2023-06-28 helpers.R

@@ -1,0 +1,200 @@
+label_alpha = label_bquote(cols = {alpha~"="~.(alpha)})
+
+extract_intervals <- function(fit) {
+  tibble(
+    lower = fit$intervals[,1],
+    upper = fit$intervals[,2],
+    pred  = fit$predictions[,1]
+  )
+}
+
+aci_theme <- theme(axis.title.y = element_text(size = 8))
+
+basic_theme <- theme(
+  axis.text.x = element_blank(), 
+  axis.ticks.x = element_blank(), 
+  legend.position = "none"
+)
+
+simulation_one_plot <- function(results) {
+  labels = ggplot2::label_bquote(
+    rows = {alpha~"="~.(alpha)}, 
+    cols = {psi~"="~theta~"="~.(param)}
+  )
+  
+  plot_coverr <- results %>%
+    ggplot(aes(x = method, y = coverage - alpha, color = method)) +
+    geom_boxplot() +
+    geom_hline(yintercept = 0, lty = 2, alpha = 0.5) +
+    facet_grid(alpha ~ param, labeller = labels) +
+    labs(x = "", y = expression(CovErr(T)), subtitle = "Coverage Error", title = "Simulation study: ARMA errors") +
+    basic_theme +
+    aci_theme
+  
+  plot_piwidth <- results %>%
+    ggplot(aes(x = method, y = mean_width, color = method)) +
+    geom_boxplot() +
+    facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
+    labs(x = "", y = expression(MeanWidth(T)), subtitle = "Mean Interval Width") +
+    basic_theme +
+    aci_theme
+  
+  plot_pathlength <- results %>%
+    ggplot(aes(x = method, y = path_length, color = method)) +
+    geom_boxplot() +
+    facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
+    labs(x = "", y = expression(PathLength(T)), subtitle = "Path Length") +
+    basic_theme +
+    aci_theme +
+    theme(axis.text.x = element_text(angle = 90, hjust = 1))
+  
+  plot_coverr / plot_piwidth / plot_pathlength
+}
+
+simulation_two_plot <- function(results) {
+  results <- results %>% mutate(
+    distribution_shift = ifelse(distribution_shift == 0.5, "Shift", "No shift")
+  )
+  
+  labels = ggplot2::label_bquote(
[…]

### 2023-07-17 paper.qmd
Refactoring heavy functions into separate file

@@
 ---
-title: "AdaptiveConformal: An R Package for Adaptive Conformal Inference"
-subtitle: "AdaptiveConformal: An R Package for Adaptive Conformal Inference"
+title: "AdaptiveConformal: An `R` Package for Adaptive Conformal Inference"
+subtitle: "AdaptiveConformal: An `R` Package for Adaptive Conformal Inference"
 author:
@@
 description: |
-  Conformal Inference is a popular approach for generating prediction intervals based on the output of any point prediction method.
-Adaptive Conformal Inference (ACI) algorithms extend classical approaches to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data.
+  Conformal Inference (CI) is a popular approach for generating finite sample prediction intervals based on the output of any point prediction method when data are exchangeable.
+Adaptive Conformal Inference (ACI) algorithms extend CI to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data.
 The common thread that unites algorithms in the ACI family is that they adaptively adjust the width of the generated prediction intervals in response to the observed data.
@@
 We then present a case study of producing prediction intervals for influenza incidence in the United States based on black-box point forecasts.
-Implementations of all the algorithms are released as an open-source R package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
+Implementations of all the algorithms are released as an open-source `R` package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
 abstract: >+
-  Conformal Inference is a popular approach for generating prediction intervals based on the output of any point prediction method.
-Adaptive Conformal Inference (ACI) algorithms extend classical approaches to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data.
+  Conformal Inference (CI) is a popular approach for generating finite sample prediction intervals based on the output of any point prediction method when data are exchangeable.
+Adaptive Conformal Inference (ACI) algorithms extend CI to the case of sequentially observed data, such as time series, and exhibit strong theoretical guarantees without having to assume exchangeability of the observed data.
 The common thread that unites algorithms in the ACI family is that they adaptively adjust the width of the generated prediction intervals in response to the observed data.
@@
 We then present a case study of producing prediction intervals for influenza incidence in the United States based on black-box point forecasts.
-Implementations of all the algorithms are released as an open-source R package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
+Implementations of all the algorithms are released as an open-source `R` package, `AdaptiveConformal`, which also includes tools for visualizing and summarizing conformal prediction intervals.
 keywords: [Conformal inference, Adaptive conformal inference, time series, R]
@@
           \usepackage{stmaryrd}
+          \usepackage{xfrac}
 ---
@@
 # Introduction
-Conformal inference methods are a family of algorithms for generating prediction intervals around point predictions [@shafer2008conformal, @angelopoulos2022gentle].
-The input point predictions can be derived from any prediction method, making conformal inference a powerful tool for augmenting black-box prediction algorithms with prediction intervals.
-Classical conformal inference methods are able to yield marginally valid intervals with only the assumption that the observed data are _exchangeable_ (that is, the joint distribution of the data does not change based on the order of the observations).
-However, in many real-world settings data are not exchangeable: for example, time series data usually can not be assumed to be exchangeable due to temporal dependence.
-A recent line of research examines the problem of generating prediction intervals for observations that are observed online (that is, one at a time) and for which exchangeability is not assumed to hold [@gibbs2021adaptive, @zaffran2022agaci, @gibbs2022faci, @bhatnagar2023saocp].
-These methods, which we refer to generally as _Adaptive Conformal Inference_ (ACI) algorithms, work by adaptively adjusting the width of the generated prediction intervals in response to the observed data.
+Conformal Inference (CI) is a family of methods for generating finite sample prediction intervals around point predictions when data are exchangeable [@shafer2008conformal; @angelopoulos2022gentle].
+The input point predictions can be derived from any prediction method, making CI a powerful tool for augmenting black-box prediction algorithms with prediction intervals.
+Classical CI methods are able to yield marginally valid intervals with only the assumption that the joint distribution of the data does not change based on the order of the observations (that is, they are exchangeable).
+However, in many real-world settings data are not exchangeable: for example, time series data usually cannot be assumed to be exchangeable due to temporal dependence.
+A recent line of research examines the problem of generating prediction intervals for observations that are observed online (that is, one at a time) and for which exchangeability is not assumed to hold [@gibbs2021adaptive; @zaffran2022agaci; @gibbs2022faci; @bhatnagar2023saocp].
+The methods from this literature, which we refer to generally as _Adaptive Conformal Inference_ (ACI) algorithms, work by adaptively adjusting the width of the generated prediction intervals in response to the observed data.
 
 Informally, suppose a sequence of outcomes $y_t \in \mathbb{R}$, $t = 1, \dots, T$ are observed one at a time.
-Before seeing each observation, we have at our disposal a point prediction $\hat{\mu}_t \in \mathbb{R}$, which can come from any source.
-Our goal is to find an algorithm for producing prediction intervals $[\ell_t, u_t]$, $\ell_t \leq u_t$ such that, in the long run, the observations $y_t$ fall within the corresponding prediction intervals roughly $\alpha \times 100\%$ of the time.
-The original Adaptive Conformal Inference algorithm [@gibbs2021adaptive] is based on a simple idea: if the previous prediction interval at time $t-1$ did not cover the true observation, then the next prediction interval at time $t$ is made slightly wider.
+Before seeing each observation, we have at our disposal a point prediction $\hat{\mu}_t \in \mathbb{R}$, which can be generated by any method.
+Our goal is to find an algorithm for producing prediction intervals $[\ell_t, u_t]$, $\ell_t \leq u_t$ such that, in the long run, the observations $y_t$ fall within the corresponding prediction intervals roughly $\alpha \times 100\%$ of the time: that is, $\lim_{T \to \infty} \sfrac{1}{T} \sum_{t=1}^T \mathbb{I}\{ y_t \in [\ell_t, u_t] \} = \alpha$.
+The original ACI algorithm [@gibbs2021adaptive] is based on a simple idea: if the previous prediction interval at time $(t-1)$ did not cover the true observation, then the next prediction interval at time $t$ is made slightly wider.
 Conversely, if the previous prediction interval did include the observation, then the next prediction interval is made slightly narrower.
-It can be shown that, by choosing wisely how fast the intervals change size, this procedure yields prediction intervals that in the long run cover the true observations roughly the desired proportion of the time.
+It can be shown that this procedure yields prediction intervals that in the long run cover the true observations the desired proportion of the time.
 
@@
 If the learning rate is too low, then the prediction intervals will not be able to adapt fast enough to shifts in the data generating distribution; if it is too large, then the intervals will oscillate widely.
-The critical dependence of the original ACI algorithm on proper choice of the its learning rate spurred subsequent research into meta-algorithms that learn the correct learning rate (or an analogue thereof) in various ways, typically drawing on approaches from the online learning literature.
+The critical dependence of the original ACI algorithm on proper choice of its learning rate spurred subsequent research into meta-algorithms that learn the correct learning rate (or an analogue thereof) in various ways, typically drawing on approaches from the online learning literature.
 In this paper, we present four such algorithms: Aggregated ACI [AgACI, @zaffran2022agaci], Fully Adaptive Conformal Inference [FACI, @gibbs2022faci], Scale-Free Online Gradient Descent [SF-OGD, @bhatnagar2023saocp], and Strongly Adaptive Online Conformal Prediction [SAOCP, @bhatnagar2023saocp].
 
-Our primary practical contribution is an implementation of each algorithm in an open source R package, `AdaptiveConformal`, which is available at [https://github.com/herbps10/AdaptiveConformal](https://github.com/herbps10/AdaptiveConformal).
+Our primary practical contribution is an implementation of each algorithm in an open source `R` package, `AdaptiveConformal`, which is available at [https://github.com/herbps10/AdaptiveConformal](https://github.com/herbps10/AdaptiveConformal).
 The package also includes routines for visualization and summary of the prediction intervals.
-We note that Python versions of several algorithms were also made available by @zaffran2022agaci and @bhatnagar2023saocp, but to our knowledge this is the first R package to implement them.
+We note that Python versions of several algorithms were also made available by @zaffran2022agaci and @bhatnagar2023saocp, but to our knowledge this is the first `R` package to implement them.
 
@@
 @sec-case-study gives a case study of generating prediction intervals for influenza incidence in the United States based on black-box point forecasts.
-Finally, @sec-discussion provides a discussion and ideas for future research in this rapidly expanding subfield.
+Finally, @sec-discussion provides a discussion and ideas for future research in this rapidly expanding field.
 
@@
 # Theoretical Framework {#sec-theory}
-*Notation*: for any integer $N \geq 1$ let $\llbracket N \rrbracket = \{ 1, \dots, N \}$. Let $\mathbb{I}$ be the indicator function. Let $\nabla f$ denote the subgradient of the function $f$.
+*Notation*: for any integer $N \geq 1$ let $\llbracket N \rrbracket := \{ 1, \dots, N \}$.
+Let $\mathbb{I}$ be the indicator function.
+Let $\nabla f$ denote the gradient (subgradient) of the differentiable (convex) function $f$.
 
@@
 Fix $\alpha \in (0, 1)$ to be the target empirical coverage of the prediction intervals.
-The goal is to output at time $t$ a prediction interval for the unseen observation $y_{t+1}$, with the prediction interval generated by an _interval construction function_ $\hat{C}_{t+1}$.
-Formally, let $\hat{C}_t$ be a function that takes as input a parameter $\theta_t \in \mathbb{R}$ and outputs a closed prediction interval $[\ell_t, u_t]$.
-The interval construction function must be nested: if $\theta^\prime > \theta$, then $\hat{C}_t(\theta) \subseteq \hat{C}_t(\theta^\prime)$.
-That is, larger values of $\theta$ imply wider prediction intervals.
+The goal is to output at time $t$ a prediction interval for the unseen observation $y_{t}$, with the prediction interval generated by an _interval construction function_ $\widehat{C}_{t+1}$.
+Formally, let $\widehat{C}_t$ be a function that takes as input a parameter $\theta_t \in \mathbb{R}$ and outputs a closed prediction interval $[\ell_t, u_t]$.
+The interval construction function must be nested: if $\theta^\prime > \theta$, then $\widehat{C}_t(\theta) \subseteq \widehat{C}_t(\theta^\prime)$.
+In words, larger values of $\theta$ imply wider prediction intervals.
 The interval constructor is indexed by $t$ to emphasize that it may use other information at each time point, such as a point prediction $\hat{\mu}_t \in \mathbb{R}$.
@@
 
-Define $r_t := \inf\{\theta \in \mathbb{R} : \mathbb{I}(y_t \in \hat{C}_t(\theta)) \}$ to be the _radius_ at time $t$.
+Define $r_t := \inf\{\theta \in \mathbb{R} : \mathbb{I}(y_t \in \widehat{C}_t(\theta)) \}$ to be the _radius_ at time $t$.
 The radius is the smallest possible $\theta$ such that the prediction interval covers the observation $y_t$.
@@
 Suppose that at each time $t$ we have access to a point prediction $\hat{\mu}_t \in \mathbb{R}$.
-Then we can form a symmetric prediction interval around the point estimate as
-$$
-\begin{aligned}
-  \hat{C}_t(\theta_t) := [\hat{\mu}_t - \theta_t, \hat{\mu}_t + \theta_t].
+Then we can form a symmetric prediction interval around the point estimate using
+$$
+\begin{aligned}
+  \theta \mapsto \widehat{C}_t(\theta) := [\hat{\mu}_t - \theta, \hat{\mu}_t + \theta].
 \end{aligned}
@@
 Note that in this case, the radius is simply the absolute residual $r_t = |\hat{\mu}_t - y_t|$.
-We focus primarily on this interval constructor in this paper, and it is the default used in the `Adaptive Conformal` R package.
+We focus primarily on this interval constructor in this paper, and it is the default used in the `AdaptiveConformal` `R` package.
 _Alternatively, we could also include results for all the algorithms run with conformal interval scores, perhaps in appendix_
@@
 The original ACI paper proposed constructing intervals based on the previously observed residuals [@gibbs2021adaptive].
-Let $S(\hat{\mu}, y) : \mathbb{R}^2 \to \mathbb{R}$ be a function called a _conformity score_.
-A popular choice of conformity score is the absolute residual: $S(\hat{\mu}, y) := |y - \hat{\mu}|$.
+Let $S : \mathbb{R}^2 \to \mathbb{R}$ be a function called a _conformity score_.
+A popular choice of conformity score is the absolute residual: $(\mu, y) \mapsto  S(\mu, y):= |\mu - y|$.
 Let $s_t := S(\hat{\mu}_t, y_t)$ be the conformity score of the $t$th-observation.
@@
 \begin{aligned}
-  \hat{C}_t(\theta_t) := [\hat{\mu}_t - \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \}), \hat{\mu}_t + \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \})].
+  \widehat{C}_t(\theta_t) := [\hat{\mu}_t - \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \}), \hat{\mu}_t + \mathrm{Quantile}(\theta, \{ s_1, \dots, s_{t-1} \})]
 \end{aligned}
@@
 where $\mathrm{Quantile}(\theta, A)$ denotes the empirical $\theta$-quantile of the elements in the set $A$.
-Note that $\hat{C}_t$ is indeed nested in $\theta_t$ because the Quantile function is non-decreasing in $\theta$.
-
-The `AdaptiveConformal` package takes the absolute residual as the default conformity score, although the user may also specify any custom conformity score by supplying it as an R function.
+Note that $\widehat{C}_t$ is indeed nested in $\theta_t$ because the Quantile function is non-decreasing in $\theta$.
+
+The `AdaptiveConformal` package takes the absolute residual as the default conformity score, although the user may also specify any custom conformity score by supplying it as an `R` function.
 
@@
 $$
-L^\alpha(\theta_t, r_t) := \begin{cases}
-  (1-\alpha)(r_t - \theta), & r_t \leq \theta, \\
-  \alpha(\theta - r_t), & \theta < r_t.
+(\theta, r) \mapsto L^\alpha(\theta, r) := \begin{cases}
+  \alpha(\theta - r), & \theta \geq r \\
+  (1-\alpha)(r - \theta), & \theta < r.
 \end{cases}
 $$
-The way in which we observe the data and incur losses is as follows:
-
-- For $t = 1, \dots, T$:
-  - Predict radius $\theta_t$ and form prediction interval $\hat{C}_t(\theta_t)$.
-  - Observe true outcome $y_t$ and calculate true radius $r_t$.
-  - Record $\mathrm{err}_t := \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
+The way in which the algorithm gains access to the data and incurs losses is as follows:
+
+- Sequentially, for $t = 1, \dots, T$:
+  - Predict radius $\theta_t$ and form prediction interval $\widehat{C}_t(\theta_t)$.
+  - Observe true outcome $y_t$ and calculate radius $r_t$.
+  - Record $\mathrm{err}_t := \mathbb{I}[y_t \not\in \widehat{C}_t(\theta_t)]$.
   - Incur loss $L^\alpha(\theta_t, r_t)$.
 
-There are two different perspectives we can take in measuring the quality of an ACI algorithm that generates a sequence $\theta_t$, $t \in \llbracket T \rrbracket$.
+## Assessing ACI algorithms
+There are two different perspectives we can take in measuring the quality of an ACI algorithm that generates a sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$.
 First, we could look at how close the empirical coverage of the generated prediction intervals is to the desired coverage level $\alpha$.
@@
 The second perspective is to look at how well the algorithm controls the incurred pinball losses.
-Following the classical framework from the online learning literature, we define the _regret_ as the difference between the cumulative loss yielded by a sequence $\theta_t$ versus the cumulative loss of the best possible fixed choice $\theta^*$:
-$$
-\begin{aligned}
-  \mathrm{Reg}(T) := \sum_{t=1}^T L^\alpha(\theta_t, r_t) - \min_{\theta^* \in \mathbb{R}} \sum_{t=1}^T L^\alpha(\theta^*, r_t).
+Following the classical framework from the online learning literature, we define the _regret_ as the difference between the cumulative loss yielded by a sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$ versus the cumulative loss of the best possible fixed choice:
+$$
+\begin{aligned}
+  \mathrm{Reg}(T) := \sum_{t=1}^T L^\alpha(\theta_t, r_t) - \inf_{\theta^* \in \mathbb{R}} \sum_{t=1}^T L^\alpha(\theta^*, r_t).
 \end{aligned}
@@
 As such, stronger notions of regret have been defined.
-The _Strongly Adaptive Regret_ is the largest regret over any subperiod of length $k \in \llbracket T \rrbracket$:
-$$
-\begin{aligned}
-  \mathrm{SAReg}(T, k) := \max_{[\tau, \tau + k - 1] \subseteq \llbracket T \rrbracket} \left( \sum_{t=\tau}^{\tau + k - 1} L^{\alpha}(\theta_t, r_t) - \inf_{\theta^*} \sum_{t=\tau}^{\tau + k - 1} L^\alpha(\theta^*, r_t) \right).
+The _strongly adaptive regret_ is the largest regret over any subperiod of length $k \in \llbracket T \rrbracket$:
+$$
+\begin{aligned}
+  \mathrm{SAReg}(T, k) := \max_{[\tau, \tau + k - 1] \subseteq \llbracket T \rrbracket} \left( \sum_{t=\tau}^{\tau + k - 1} L^{\alpha}(\theta_t, r_t) - \inf_{\theta^* \in \mathbb{R}} \sum_{t=\tau}^{\tau + k - 1} L^\alpha(\theta^*, r_t) \right).
 \end{aligned}
@@
 Consider a simple class of algorithms that outputs constantly $\theta_t = \theta'$ for some $\theta' < 1$.
-With the linear interval construction function, the prediction intervals are then $\hat{C}_t(\theta_t) = [-\theta', \theta']$.
+With the linear interval construction function, the prediction intervals are then $\widehat{C}_t(\theta_t) = [-\theta', \theta']$.
 The regret is given by $\mathrm{Reg}(T) = 2T\alpha(1-\theta')$, which approaches zero as $\theta'$ approaches 1.
@@
 These simple examples illustrate that, unfortunately, bounds on the coverage error and bounds on the regret are not in general interchangeable.
-It is possible, however, to show equivalencies by either (1) making distributional assumptions on the data or (2) using additional information about how the algorithm produces the iterates $\theta_t$.
-
-It may also be informative to summarize a set of prediction intervals in ways beyond their coverage error or by their regret. A common metric for prediction intervals is the _mean interval width_:
+It is possible, however, to show equivalencies by either (1) making distributional assumptions on the data or (2) using additional information about how the algorithm produces the sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$.
+
+It may also be informative to summarize a set of prediction intervals in ways beyond their coverage error or their regret. A common metric for prediction intervals is the _mean interval width_:
 $$
@@
 $$
-where $w_t := u_t - \ell-t$ is the interval width at time $t$.
+where $w_t := u_t - \ell_t$ is the interval width at time $t$.
 Finally, we introduce a metric that is intended to capture pathological behavior that can arise with ACI algorithms where the prediction intervals oscillate between being extremely narrow and extremely wide.
@@
 We have harmonized the algorithm descriptions to use a common structure and notation, in some cases departing from their original descriptions.
-
-As a simple running example to illustrate each algorithm, we simulate $T = 500$ values $y_1, \dots, y_T$ following
+As a quick reference, the algorithms are summarized in @tbl-aci.
+
+| Algorithm                                             | Tuning Parameters                                           | Citation            |
+| ------------------------------------------------------|-------------------------------------------------------------|---------------------|
+| Adaptive Conformal Inference (ACI)                    | Learning rate $\gamma$                                      | @gibbs2021adaptive  |
+| Aggregated Adaptive Conformal Inference (AgACI)       | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | @zaffran2022agaci   |
+| Fully Adaptive Conformal Inference (FACI)             | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | @gibbs2022faci      |
+| Scale-Free Online Gradient Descent (SF-OGD)           | Learning rate $\gamma$ or maximum radius $D$                | @bhatnagar2023saocp |
+| Strongly Adaptive Online Conformal Prediction (SAOCP) | Learning rate $\gamma$, lifetime multiplier $g$                 | @bhatnagar2023saocp |
+
+  : Summary of ACI algorithms {#tbl-aci}
+
+As a simple running example to illustrate each algorithm, we simulate ndependently $T = 500$ values $y_1, \dots, y_T$ following
 $$
@@
 $$
-For demonstration purposes we assume to have access to unbiased predictions $\hat{\mu}_t = 0$ for all $t$.
+For demonstration purposes we assume to have access to unbiased predictions $\hat{\mu}_t = 0$ for all $t \in \llbracket T \rrbracket$. Throughout we set a target empirical coverage of $\alpha = 0.8$.
 
@@
 \For{$t = 1, 2, \dots, T$}
-  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$.
+  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$.
-  \State Set $\mathrm{err}_t = \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
+  \State Evaluate $\mathrm{err}_t = \mathbb{I}[y_t \not\in \widehat{C}_t(\theta_t)]$.
   \State Update $\theta_{t+1} = \theta_t + \gamma (\mathrm{err}_t - (1 - \alpha))$.
@@
 \begin{aligned}
-  \nabla L^\alpha(\theta_t, r_t) &= \begin{cases}
-    \alpha - 1, & r_t < \theta, \\
-    \alpha, &\theta < r_t, \\
-    [\alpha - 1, \alpha], &\theta = r_t
+  \nabla L^\alpha(\theta, r) &= \begin{cases}
+    -\alpha, &\theta < r, \\
+    1 - \alpha, & \theta > r, \\
+    [-\alpha, 1 - \alpha], &\theta = r
   \end{cases} \\
@@
 $$
-\alpha - 1 + \mathrm{err}_t \in \nabla L^\alpha(\theta_t, r_t).
+1 - \alpha - \mathrm{err}_t \in \nabla L^\alpha(\theta_t, r_t).
 $$
@@
 For intuition, note that if $y_t$ fell outside of the prediction interval at time $t$ ($\mathrm{err}_t = 1$) then the next interval is widened ($\theta_{t+1} = \theta_t + \gamma \alpha$).
-On the other hand, if $y_t$ fell within the interval ($\mathrm{err}_t = 0$) then the next interval is shortened ($\theta_{t+1} = \theta_t - \gamma(1 - \alpha)$).
-The learning rate $\gamma$ controls how fast the width of the prediction intervals change in response to the data.
+On the contrary, if $y_t$ fell within the interval ($\mathrm{err}_t = 0$) then the next interval is shortened ($\theta_{t+1} = \theta_t - \gamma(1 - \alpha)$).
+The learning rate $\gamma$ controls how fast the width of the prediction intervals changes in response to the data.
 
@@
 where the second inequality follows if the optimal choice of $\gamma = D/\sqrt{T}$ is used [@bhatnagar2023saocp].
-In addition, a finite sample bound on the coverage error can be established [@gibbs2021adaptive]:
+A finite sample bound on the coverage error can also be established [@gibbs2021adaptive]:
 $$
@@
 $$
+ Taken together, the theoretical results imply that while the coverage error is guaranteed to converge to zero for any choice of $\gamma$, achieving sublinear regret requires choosing $\gamma$ more carefully.
+This highlights the importance of both ways of assessing ACI algorithms: if we only focused on controlling the coverage error, we might not achieve optimal control of regret, leading to intervals that are not practically useful.
 
 ### Tuning Parameters
-The only tuning parameter is the learning rate $\gamma$.
+The main tuning parameter is the learning rate $\gamma$.
 Setting $\gamma$ too small will lead to intervals that do not adapt fast enough to distribution shifts, and setting $\gamma$ too large will lead to intervals with large oscillations as seen in @fig-aci.
-This is quantified in the path lengths, which increase significantly as $\gamma$ increases, even though the empirical coverage remains near the desired value.
+This is quantified in the path length, which increases significantly as $\gamma$ increases, even though the empirical coverage remains near the desired value of 80\%.
 Thus, choosing a good value for $\gamma$ is essential.
 However, the optimal choice $\gamma = D / \sqrt{T}$ cannot be used directly in practice unless the time series length $T$ is fixed in advance, or the so called "doubling trick" is used to relax the need to know $T$ in advance [@cesabianchi2006games].
+
+The theoretical results guaranteeing the performance of the ACI algorithm do not depend on the choice of starting value $\theta_1$, and thus in practice any value can be chosen.
+Indeed, the effect of the choice of $\theta_1$ decays over time as a function of the chosen learning rate.
+In practice, substantive prior information can be used to pick a reasonable starting value.
+By default, the `AdaptiveConformal` package sets $\theta_1 = 0$.
 
@@
 \begin{algorithmic}
-  \State \textbf{Input:} candidate learning rates $\{ \gamma_k \}_{1 \leq k \leq K }$, starting value $\theta_1$.
-  \State Initialize BOA algorithms $\mathcal{B}^\ell$ and $\mathcal{B}^u$.
+  \State \textbf{Input:} candidate learning rates $(\gamma_k)_{1 \leq k \leq K }$, starting value $\theta_1$.
+  \State Initialize lower and upper BOA algorithms $\mathcal{B}^\ell := \texttt{BOA}(\alpha \leftarrow (1 - \alpha / 2))$ and $\mathcal{B}^u := \texttt{BOA}(\alpha \leftarrow (1 - (1 - \alpha)/2))$.
   \For{$k = 1, \dots, K$}
@@
     \For{$k = 1, \dots, K$}
-      \State Generate candidate prediction interval $[\ell^k_{t}, u^k_{t}]$ from $\mathcal{A}_k$.
+      \State Retrieve candidate prediction interval $[\ell^k_{t}, u^k_{t}]$ from $\mathcal{A}_k$.
     \EndFor
-    \State Compute aggregated lower and upper bounds $\tilde{\ell}_t$ and $\tilde{u}_t$ using $\mathcal{B}^\ell$ and $\mathcal{B}^u$.
+    \State Compute aggregated lower bound $\tilde{\ell}_t := \mathcal{B}^\ell((\ell^k_t : k \in \{ 1, \dots, K \}))$.
+    \State Compute aggregated upper bound $\tilde{u}_t := \mathcal{B}^u((u^k_t : k \in \{ 1, \dots, K \}))$.
     \State \textbf{Output} prediction interval $[\tilde{\ell}_t, \tilde{u}_t]$.
@@
     \For{$k = 1, \dots, K$}
-      \State Update ACI $\mathcal{A}_k$ with $y_t$.
+      \State Update $\mathcal{A}_k$ with observation $y_t$.
     \EndFor
-    \State Update $\mathcal{B}^\ell$ with lower bounds $\ell^k_t$, $k = 1, \dots, K$ and observed outcome $y_t$. 
-    \State Update $\mathcal{B}^u$ with upper bounds $u^k_t$, $k = 1, \dots, K$ and observed outcome $y_t$. 
+    \State Update $\mathcal{B}^\ell$ with observed outcome $y_t$. 
+    \State Update $\mathcal{B}^u$ with observed outcome $y_t$.
   \EndFor
@@
 
-The Aggregated ACI (AgACI; @fig-algo-agaci) algorithm solves the problem of choosing a learning rate for ACI by running multiple copies of the algorithm with different learning rates, and then combining their predictions using an online aggregation of experts algorithm [@zaffran2022agaci].
-We follow the AgACI paper in using the Bernstein Online Aggregation (BOA) algorithm as implemented in the `opera` R package [@wintenberger2017boa; @opera2023].
+The Aggregated ACI (AgACI; @fig-algo-agaci) algorithm solves the problem of choosing a learning rate for ACI by running multiple copies of the algorithm with different learning rates, and then combining separately the lower and upper interval bounds using an online aggregation of experts algorithm [@zaffran2022agaci].
+That is, one aggreation algorithm seeks to estimate the lower $(1-\alpha)/2$ quantile, and the other seeks to estimate the upper $1 - (1 - \alpha) / 2$ quantile.
+They experimented with multiple online aggregation algorithms, and found that they yielded similar results.
+Thus, we follow [@zaffran2022agaci] in using the Bernstein Online Aggregation (BOA) algorithm as implemented in the `opera` `R` package [@wintenberger2017boa; @opera2023].
+BOA is an online algorithm that forms predictions for the lower (or upper) prediction interval bound as a weighted mean of the candidate ACI prediction interval lower (upper) bound, where the weights are determined by each candidate's past performance with respect to the quantile loss.
+As a consequence, the prediction intervals generated by AgACI are not necessarily symmetric around the point prediction, as the weights for the lower and upper bounds are separate.
 
 ### Theoretical Gaurantees
-AgACI departs from our main theoretical framework in that it does not yield a set of iterates $\theta_1, \dots, \theta_T$ which are used to form prediction intervals using a set construction function $\hat{C}_t$.
-Rather, the upper and lower bounds from a set of candidate ACI algorithms are aggregated separately.
+AgACI departs from our main theoretical framework in that it does not yield a sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$ whose elements yield prediction intervals via a set construction function $\widehat{C}_t$.
+Rather, the upper and lower interval bounds from a set of candidate ACI algorithms are aggregated separately.
 Thus, theoretical results such as regret bounds in terms of $\theta$ similar to those for the other algorithms are not available.
 It would be possible, however, to establish regret bounds for the pinball loss applied separately to the lower and upper bounds of the prediction intervals.
-It's not clear, however, how to convert such regret bounds into a coverage bound.
+It is unclear, however, how to convert such regret bounds into a coverage bound.
 
 ### Tuning Parameters
-The only tuning parameter for AgACI is the set of candidate learning rates.
-Beyond necessitating additional computational time, there isn't a drawback to having a large grid.
-As a basic check, we can also look at the weights assigned to each of the learning rates.
-If large weights are given to the smallest or largest learning rates, it is a sign that smaller (or larger) learning rates would perform well.
+The main tuning parameter for AgACI is the set of candidate learning rates.
+Beyond necessitating additional computational time, there is no drawback to having a large grid.
+As a default, `AdaptiveConformal` uses learning rates $\gamma \in \{ 0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128 \}$.
+As a basic check, we can look at the weights assigned to each of the learning rates.
+If large weights are given to the smallest (largest) learning rates, it is a sign that smaller (or larger) learning rates would perform well.
+In addition each of the candidate ACI algorithms requires a starting value, which can be set to any value as discussed in the ACI section.
+@fig-agaci illustrates AgACI applied to the running example with two sets of learning grids.
+The first grid is the default, and the second grid is double the size, given by $\gamma_i = 2^i / 1000$ for $i = 0, 0.5, 1, \dots, 7$.
+For the first grid, we can see that for the lower bound AgACI assigns high weight to only one learning rate ($\gamma = 0.032$).
+Based on this observation, as a sensitivity check, we reran the algorithm with the second, denser learning grid, which yields weights that are less concentrated on a single learning rate.
+The output prediction intervals are similar, however, for both grid choices.
 
@@
 #| fig-height: 4.5
-#| fig-cap: "Aggregated Adaptive Conformal Inference (AgACI) using the linear interval construction function and with initialization $\\theta_1 = 0$."
-result <- aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear"))
-coverage <- scales::percent_format()(result$metrics$coverage)
-path_length <- scales::number_format(accuracy = 0.1)(result$metrics$path_length)
-plot(result, ylim = c(-0.9, 0.8), legend = FALSE)
-text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage) ), pos = 4)
-text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length) ), pos = 4)
+#| fig-cap: "Aggregated Adaptive Conformal Inference (AgACI) using the linear interval construction function, starting values $\\theta_1 = 0$, and two different learning rate grids, where the second grid is double the size of the first."
+grid1 <- 2^(seq(0, 7, 1)) / 1e3
+grid2 <- 2^(seq(0, 7, 0.5)) / 1e3
+results <- list(
+  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear", gamma_grid = grid1)),
+  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear", gamma_grid = grid2))
+)
+
+coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
+path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))
+mean_width <- scales::number_format(accuracy = 0.01)(unlist(lapply(results, function(result) result$metrics$mean_width)))
+
+par(mfrow = c(2, 2), mar = c(4, 3, 2, 1))
+for(i in 1:2) {
+  plot(results[[i]], legend = FALSE, predictions = FALSE, cex = 0.5, main = paste("Grid", i), ylim = c(-0.9, 0.8))
+  text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
+  text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[i]]) ), pos = 4)
+
+  plot_agaci_weights(results[[i]], main = "Weights", legend = "topleft")
+}
+par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
+```
+
+```{r agaci_example_plot2, echo = FALSE}
+#| label: fig-agaci2
+#| fig-height: 3
+#| fig-cap: "Weights assigned to each of the candidate ACI algorithms by AgACI at the final timepoint."
 ```
@@
 \begin{algorithmic}
-\State \textbf{Input:} starting value $\theta_1$, candidate learning rates $\{ \gamma_k \}_{1 \leq k \leq K }$, parameters $\sigma, \eta$.
+\State \textbf{Input:} starting value $\theta_1$, candidate learning rates $() \gamma_k )_{1 \leq k \leq K }$, parameters $\sigma, \eta$.
 \For{$k = 1, \dots, K$}
@@
   \State Set $\theta_t = \sum_{k=1}^K \theta_t^k p_t^k$.
-  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$.
+  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$ and compute $r_t$.
@@
   \State $w_{t+1}^k \gets (1 - \sigma) \bar{w}_t^k + \bar{W}_t \sigma / K$.
-  \State Set $\mathrm{err}_t := \mathbb{I}[y_t \not\in \hat{C}_t(\theta_t)]$.
+  \State Set $\mathrm{err}_t := \mathbb{I}[y_t \not\in \widehat{C}_t(\theta_t)]$.
   \For{$k = 1, \dots, K$}
@@
 
-The Fully Adaptive Conformal Inference method also aggregates predictions from multiple copies of ACI run with different learning rates, but differs from AgACI in that it aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci].
-FACI uses the aggregation scheme proposed in [@gradu2022adaptive].
+The Fully Adaptive Conformal Inference method also aggregates predictions from multiple copies of ACI run with different learning rates, but differs from AgACI in that it directly aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci].
+FACI uses the aggregation scheme proposed in @gradu2022adaptive.
 
 ### Theoretical Guarantees
-Let $\gamma_{\mathrm{max}} = \max_{1 \leq k \leq K} \gamma_k$ and assume that $\gamma_1 < \gamma_2 < \cdots < \gamma_K$ with $\gamma_{k+1}/\gamma \leq 2$ for all $1 \leq k < K$.
-Then, for any interval $I = [r, s] \subseteq \llbracket T \rrbracket$ and any sequence $\theta_r^*, \dots, \theta_s^*$,
-$$
-\begin{aligned}
-  \frac{1}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)] - \frac{1}{|I|} \sum_{t=r}^s \leq& \frac{\log(k / \sigma) + 2\sigma|I|}{\eta |I|} + \frac{\eta}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)^2] \\
-  &+ 2\sqrt{3}(1 + \gamma_{\mathrm{max}})^2 \max\left\{ \sqrt{\frac{\sum_{t=r+1}^s |\beta_t^* - \beta_{t-1}^*}{|I|}}, \gamma_1 \right\},
-\end{aligned}
-$$
-where the expectation is over the randomness in the randomized version of the algorithm. 
-To convert the regret bound to a coverage bound, the authors took the path of making a distributional assumption on the radii $r_t$.
-Specifically, they assume that each of the conditional laws of $r_t | \{ r_s \}_{s < t}$ have a density that is bounded below by $p > 0$.
-For any interval $[s, r] := I \subseteq \llbracket T \rrbracket$ and any sequence $\theta_t^*$, $s \leq t \leq r$:
-$$
-\begin{aligned}
-  \frac{1}{|I|} \sum_{t=r}^s \frac{p \mathbb{E}[(\theta_t - \theta^*_t)^2]}{2} \leq O\left( \sqrt{\frac{\log(|I|)}{|I|}} \right) + O\left( \sqrt{\frac{\sum_{t=r+1}^s |\theta_t^* - \theta_{t-1}^*|}{|I|}} \right).
-\end{aligned}
-$$
-The term $\sum_{t=r+1}^s | \theta_t^* - \theta_{t-1}^*|$ is a path length providing a scalar summary of the degree of distribution shift in the data.
-
-It has also been pointed out that FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]:
+FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]:
 $$
@@
 $$
-If $k$ is taken as fixed, then choosing $\gamma = D/\sqrt{k}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{k})$ (for a single choice of $k$).
+If $k$ is fixed a-priori, then choosing $\gamma = D/\sqrt{k}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{k})$ (for a single choice of $k$).
+Practically, this result implies that, if we know in advance the time length for which we would like to control the regret, it is possible to choose an optimal tuning parameter value.
+However, we cannot control the regret simultaneously for all possible time lengths.
+
+To establish a bound on the coverage error, the authors investigated a slightly modified version of FACI in which $\theta_t$ is chosen randomly from the candidate $\theta_{t_k}$ with weights given by $p_{t,k}$, instead of taking a weighted average.
+This is a common trick used in the literature as it facilitates theoretical analysis.
+In practice, the authors comment that this randomized version of FACI and the deterministic version lead very similar results.
+The coverage error result also assumes that the hyperparameters can change over time: that is, we have $\eta_{t}$ and $\sigma_t$, rather than $\eta$ and $\sigma$.
+The coverage error then has the following bound (Theorem 3.2, @gibbs2022faci), where $\gamma_{\min}$ and $\gamma_{\max}$ are the smallest and largest learning rates in the grid, respectively:
+$$
+|\mathrm{CovErr}(T)| \leq \frac{1 + 2\gamma_{\max}}{T \gamma_{\min}} + \frac{(1 + 2\gamma_{\max})^2}{\gamma_{\min}} \exp(\eta_t(1 + 2\gamma_{\max})) \frac{1}{T}\sum_{t=1}^T \eta_t + 2 \frac{1+\gamma_{\max}}{\gamma_{\min}} \frac{1}{T} \sum_{t=1}^T \sigma_t.
+$$
+Thus, if $\eta_t$ and $\sigma_t$ both converge to zero as $t \to \infty$, then the coverage error will also converge to zero.
+In addition, under mild distributional assumptions the authors provide a type of short-term coverage error bound for arbitrary time spans, for which we refer to Proposition 3.2 [@gibbs2022faci].
 
@@
 Choosing $\eta$ is more difficult.
-The authors suggest the following choice for $\eta$, which is optimal if there is no distribution shift:
+The authors suggest the following choice for $\eta$, which they show is optimal if there is in fact no distribution shift:
 $$
@@
 In the `AdaptiveConformal` package, the first approach is used when the conformal interval construction function is chosen, and the latter approach for the linear interval construction function.
+
+@fig-faci-linear illustrates FACI with a linear interval construction function and with the learning rate grid $\gamma \in \{ 0.05, 0.1, 0.15, 0.2, 0.25 \}$.
 
@@
 \For{$t = 1, 2, \dots, T$}
-  \State \textbf{Output} prediction interval $\hat{C}_t(\theta_t)$. 
+  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$. 
   \State Observe $y_t$ and compute $r_t$. 
@@
 Scale-Free Online Gradient Descent is a general algorithm for online learning proposed by @orabona2018sfogd.
-The algorithm updates $\theta_t$ with a gradient descent step where the learning rate adapts to the scale of the observed gradients.
+The algorithm updates $\theta_t$ with a gradient descent step where the learning rate adapts to the scale of the previously observed gradients.
 SF-OGD was first used in the context of ACI as a sub-algorithm for SAOCP (described in the next section).
@@
 A bound for the coverage error has also been established [@bhatnagar2023saocp].
-For any learning rate $\gamma = \Theta(D)$ (where $\gamma = D / \sqrt{3}$ is optimal) and any initialization $\theta_1 \in [0, D]$, then it holds that for any $T > 1$,
+For any learning rate $\gamma = \Theta(D)$ (where $\gamma = D / \sqrt{3}$ is optimal) and any starting value $\theta_1 \in [0, D]$, then it holds that for any $T > 1$,
 $$
@@
 When $D$ is not known, it can be estimated by using an initial subset of the time series as a calibration set and estimating $D$ as the maximum of the absolute residuals of the observations and the predictions [@bhatnagar2023saocp].
+@fig-sf-ogd illustrates SF-OGD for several values of $\gamma$.
 
@@
   \State Set $\theta_t = \sum_{i \in \mathrm{Active}(t)} p_i \theta_t^i$ (for $t \geq 2$), and $\theta_t = 0$ for $t = 1$.
-  \State \textbf{Output} prediction set $\hat{C}_t(\theta_t)$.
+  \State \textbf{Output} prediction set $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$ and compute $r_t$. 
@@
 We follow the original paper in setting $g = 8$.
+@fig-saocp illustrates the SAOCP algorithm for choices of $D \in \{0.01, 0.1, 0.25, 0.5 \}$.
 
@@
 
-To provide point predictions for the ACI algorithms, at each time $t \geq 200$ a random forest model was fitted to the previously observed data using the `ranger` R package [@wright2017ranger].
+To provide point predictions for the ACI algorithms, at each time $t \geq 200$ a random forest model was fitted to the previously observed data using the `ranger` `R` package [@wright2017ranger].
 The estimated model was then used to predict the subsequent time point.
@@
 ## Distribution shift
-This simulation study features time series that feature distribution shifts.
+This simulation study features time series with distribution shifts.
 The setup is quite simple in order to probe the basic performance of the methods in response to distribution shift.
@@
 
+-- Different AgACI weights for lower and upper bounds -- a step towards asymmetry, more flexible
+
 _Final wrap up paragraph?_
@@
 
-## Additional R code
+## Additional `R` code
 ```{r}

### 2023-07-17 helpers.R

@@ -69,7 +69,7 @@ simulation_two_plot <- function(results) {
     facet_grid(distribution_shift ~ alpha, labeller = labels) +
     labs(x = "", y = expression(CovErr(T)), subtitle = "Coverage Error", 
          title = "Simulation Study: Distribution Shift") +
-    basic_theme +
+    theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) +
     aci_theme
   
   plot_piwidth <- results %>%
@@ -79,8 +79,7 @@ simulation_two_plot <- function(results) {
     facet_grid(distribution_shift ~ alpha, labeller = labels, scales = "free_y") +
     labs(x = "", y = expression(MeanWidth(T)), subtitle = "Mean Interval Width") +
     aci_theme +
-    basic_theme +
-    theme(legend.position = "none")
+    theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))
   
   plot_pathlength <- results %>%
     ggplot(aes(x = method,
@@ -88,8 +87,8 @@ simulation_two_plot <- function(results) {
     geom_boxplot() +
     facet_grid(distribution_shift ~ alpha, labeller = labels, scales = "free_y") +
     labs(x = "ACI Method", y = expression(PathLength(T)), subtitle = "Path Length") +
-    aci_theme +
-    theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))
+    theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) +
+    aci_theme
   
   plot_coverr / plot_piwidth / plot_pathlength
 }

### 2023-10-09 paper.qmd
Updated paper draft

@@
     orcid: 0000-0002-3540-8255
-    #affiliations:
-    #  - name: CNRS, Université Paris Dauphine - PSL
-    #    url: https://cnrs.fr
-    #  - name: PR[AI]RIE Institute
-    #    url: https://prairie-institute.fr/
-  - name: Margaux Zaffran
-    email: margaux.zaffran@inria.fr
-    url: https://mzaffran.github.io/
-    orcid: 0000-0001-7560-8916
-    #affiliations:
-    #  - name: INRIA
-    #    url: https://inria.fr
-    #  - name: CMAP, Ecole Polytechnique
-    #    url: https://inria.fr
-    #  - name: EDF R\&D
-    #    url: https://inria.fr
+    affiliations:
+      - name: CEREMADE (UMR 7534), Université Paris-Dauphine PSL, Place du Maréchal de Lattre de Tassigny, Paris, 75016, France
+        url: https://www.ceremade.dauphine.fr/
   - name: Antoine Chambaz
@@
     orcid: 0000-0002-5592-6471
-    #affiliations:
-    #  - name: Université Paris Cité
-    #    department: MAP5
-    #    url: https://map5.mi.parisdescartes.fr/
+    affiliations:
+      - name: MAP5 (UMR 8145), Université Paris Cité
+        department: 
+        url: https://map5.mi.parisdescartes.fr/
   - name: Julie Josse
@@
     orcid: 0000-0001-9547-891X
-    #affiliations:
-    #  - name: INRIA Sophia-Antipolis
-    #    url: https://www.inria.fr/fr/centre-inria-universite-cote-azur
+    affiliations:
+      - name: Inria PreMeDICaL team, Université de Montpellier
+        url: https://team.inria.fr/premedical/
 date: last-modified
@@
 # Introduction
-Conformal Inference (CI) is a family of methods for generating finite sample prediction intervals around point predictions when data are exchangeable [@shafer2008conformal; @angelopoulos2022gentle].
+Conformal Inference (CI) is a family of methods for generating finite sample prediction intervals around point predictions when data are exchangeable [@vovk2005; @shafer2008conformal; @angelopoulos2022gentle].
 The input point predictions can be derived from any prediction method, making CI a powerful tool for augmenting black-box prediction algorithms with prediction intervals.
@@
 Informally, suppose a sequence of outcomes $y_t \in \mathbb{R}$, $t = 1, \dots, T$ are observed one at a time.
-Before seeing each observation, we have at our disposal a point prediction $\hat{\mu}_t \in \mathbb{R}$, which can be generated by any method.
+Before seeing each observation, we have at our disposal a point prediction $\hat{\mu}_t \in \mathbb{R}$ that can be generated by any method.
 Our goal is to find an algorithm for producing prediction intervals $[\ell_t, u_t]$, $\ell_t \leq u_t$ such that, in the long run, the observations $y_t$ fall within the corresponding prediction intervals roughly $\alpha \times 100\%$ of the time: that is, $\lim_{T \to \infty} \sfrac{1}{T} \sum_{t=1}^T \mathbb{I}\{ y_t \in [\ell_t, u_t] \} = \alpha$.
@@
 In this paper, we present four such algorithms: Aggregated ACI [AgACI, @zaffran2022agaci], Fully Adaptive Conformal Inference [FACI, @gibbs2022faci], Scale-Free Online Gradient Descent [SF-OGD, @bhatnagar2023saocp], and Strongly Adaptive Online Conformal Prediction [SAOCP, @bhatnagar2023saocp].
+We note that the adaption of conformal inference techniques is an active area of research and the algorithms we focus on in this work are not exhaustive; see among others @feldman2023achieving, @bastani2022practical, @xu2021enbpi, and @xu2023spci.
 
@@
 The package also includes routines for visualization and summary of the prediction intervals.
-We note that Python versions of several algorithms were also made available by @zaffran2022agaci and @bhatnagar2023saocp, but to our knowledge this is the first `R` package to implement them.
+We note that Python versions of several algorithms were also made available by @zaffran2022agaci and @bhatnagar2023saocp, but to our knowledge this is the first package implementing them in `R`.
+In addition, several `R` packages exist for conformal inference in other contexts, including `conformalInference` focusing on regression [@tibshirani2019ci], `conformalInference.fd`, with methods for functional responses [@diquigiovanni2022fd], and `cfcausal` for causal inference related functionals [@lei2020cfcausal].
+Our second practical contribution is to compare the performance of the algorithms in simulation studies and in a case study generating prediction intervals for influenza incidence in the United States based on black-box point forecasts.
 
@@
 In @sec-simulations we compare the performance of the algorithms in several simulation studies.
-@sec-case-study gives a case study of generating prediction intervals for influenza incidence in the United States based on black-box point forecasts.
+@sec-case-study gives a case study based on forecasting influenza in the United States.
 Finally, @sec-discussion provides a discussion and ideas for future research in this rapidly expanding field.
 
-```{r, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
 library(tidyverse)
@@
 
-# Contents of helpers.R included in the appendix
+knitr::opts_chunk$set(echo = FALSE)
+
 source("helpers.R")
@@
 Fix $\alpha \in (0, 1)$ to be the target empirical coverage of the prediction intervals.
-The goal is to output at time $t$ a prediction interval for the unseen observation $y_{t}$, with the prediction interval generated by an _interval construction function_ $\widehat{C}_{t+1}$.
+The goal is to output at time $t$ a prediction interval for the unseen observation $y_{t}$, with the prediction interval generated by an _interval construction function_ $\widehat{C}_{t}$.
 Formally, let $\widehat{C}_t$ be a function that takes as input a parameter $\theta_t \in \mathbb{R}$ and outputs a closed prediction interval $[\ell_t, u_t]$.
@@
 The radius is the smallest possible $\theta$ such that the prediction interval covers the observation $y_t$.
-A key assumption for the theoretical analysis of several of the algorithms is that the radii are bounded: that is, there exists a $D > 0$ such that $r_t < D$ for all $t$.
-
-Next, we describe two possible definitions of the interval construction function.
+A key assumption for the theoretical analysis of several of the algorithms is that the radii are bounded:
+
+**Assumption**: there exists a $D > 0$ such that $r_t < D$ for all $t$.
+
+Next, we describe two existing definitions of interval construction functions.
 
@@
 $$
-We refer to this as the _linear interval constructor_.
-Note that in this case, the radius is simply the absolute residual $r_t = |\hat{\mu}_t - y_t|$.
-We focus primarily on this interval constructor in this paper, and it is the default used in the `AdaptiveConformal` `R` package.
-_Alternatively, we could also include results for all the algorithms run with conformal interval scores, perhaps in appendix_
-
-## Conformal Intervals
+We refer to this as the _linear interval constructor_. Note that in this case, the radius is simply the absolute residual $r_t = |\hat{\mu}_t - y_t|$.
+
+## Quantile Intervals
 The original ACI paper proposed constructing intervals based on the previously observed residuals [@gibbs2021adaptive].
-Let $S : \mathbb{R}^2 \to \mathbb{R}$ be a function called a _conformity score_.
-A popular choice of conformity score is the absolute residual: $(\mu, y) \mapsto  S(\mu, y):= |\mu - y|$.
-Let $s_t := S(\hat{\mu}_t, y_t)$ be the conformity score of the $t$th-observation.
-The conformal interval construction function is then given by
+Let $S : \mathbb{R}^2 \to \mathbb{R}$ be a function called a _nonconformity score_.
+A popular choice of nonconformity score is the absolute residual: $(\mu, y) \mapsto  S(\mu, y):= |\mu - y|$.
+Let $s_t := S(\hat{\mu}_t, y_t)$ be the nonconformity score of the $t$th-observation.
+The quantile interval construction function is then given by
 $$
@@
 
-The `AdaptiveConformal` package takes the absolute residual as the default conformity score, although the user may also specify any custom conformity score by supplying it as an `R` function.
+Our proposed `AdaptiveConformal` package takes the absolute residual as the default nonconformity score, although the user may also specify any custom nonconformity score by supplying it as an `R` function.
 
@@
 
+This iterative procedure is at the core of the online learning theoretical framework in which theoretical results have been derived.
+
 ## Assessing ACI algorithms
@@
 \begin{aligned}
-  \mathrm{SAReg}(T, k) := \max_{[\tau, \tau + k - 1] \subseteq \llbracket T \rrbracket} \left( \sum_{t=\tau}^{\tau + k - 1} L^{\alpha}(\theta_t, r_t) - \inf_{\theta^* \in \mathbb{R}} \sum_{t=\tau}^{\tau + k - 1} L^\alpha(\theta^*, r_t) \right).
+  \mathrm{SAReg}(T, m) := \max_{[\tau, \tau + m - 1] \subseteq \llbracket T \rrbracket} \left( \sum_{t=\tau}^{\tau + m - 1} L^{\alpha}(\theta_t, r_t) - \inf_{\theta^* \in \mathbb{R}} \sum_{t=\tau}^{\tau + m - 1} L^\alpha(\theta^*, r_t) \right).
 \end{aligned}
@@
 These simple examples illustrate that, unfortunately, bounds on the coverage error and bounds on the regret are not in general interchangeable.
-It is possible, however, to show equivalencies by either (1) making distributional assumptions on the data or (2) using additional information about how the algorithm produces the sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$.
+It is possible, however, to show equivalencies by either (1) making distributional assumptions on the data or (2) using additional information about how the algorithm produces the sequence $(\theta_t)_{t \in \llbracket T \rrbracket}$ [@bhatnagar2023saocp].
 
@@
 where $w_t := u_t - \ell_t$ is the interval width at time $t$.
+
 Finally, we introduce a metric that is intended to capture pathological behavior that can arise with ACI algorithms where the prediction intervals oscillate between being extremely narrow and extremely wide.
@@
 # Algorithms {#sec-algorithms}
-In this section we describe five ACI algorithms and summarize the theoretical guarantees that have been proven for each one.
-We have harmonized the algorithm descriptions to use a common structure and notation, in some cases departing from their original descriptions.
-As a quick reference, the algorithms are summarized in @tbl-aci.
-
-| Algorithm                                             | Tuning Parameters                                           | Citation            |
+
+| Algorithm                                             | Tuning Parameters                                           | Original interval constructor | Citation            |
 | ------------------------------------------------------|-------------------------------------------------------------|---------------------|
-| Adaptive Conformal Inference (ACI)                    | Learning rate $\gamma$                                      | @gibbs2021adaptive  |
-| Aggregated Adaptive Conformal Inference (AgACI)       | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | @zaffran2022agaci   |
-| Fully Adaptive Conformal Inference (FACI)             | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | @gibbs2022faci      |
-| Scale-Free Online Gradient Descent (SF-OGD)           | Learning rate $\gamma$ or maximum radius $D$                | @bhatnagar2023saocp |
-| Strongly Adaptive Online Conformal Prediction (SAOCP) | Learning rate $\gamma$, lifetime multiplier $g$                 | @bhatnagar2023saocp |
+| Adaptive Conformal Inference (ACI)                    | Learning rate $\gamma$                                      | Quantile | @gibbs2021adaptive  |
+| Aggregated Adaptive Conformal Inference (AgACI)       | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | Quantile | @zaffran2022agaci   |
+| Fully Adaptive Conformal Inference (FACI)             | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | Quantile | @gibbs2022faci      |
+| Scale-Free Online Gradient Descent (SF-OGD)           | Learning rate $\gamma$ or maximum radius $D$             | Linear | @bhatnagar2023saocp |
+| Strongly Adaptive Online Conformal Prediction (SAOCP) | Learning rate $\gamma$, lifetime multiplier $g$          | Linear | @bhatnagar2023saocp |
 
@@
 
-As a simple running example to illustrate each algorithm, we simulate ndependently $T = 500$ values $y_1, \dots, y_T$ following
+As a simple running example to illustrate each algorithm, we simulate independently $T = 500$ values $y_1, \dots, y_T$ following
 $$
@@
 $$
-For demonstration purposes we assume to have access to unbiased predictions $\hat{\mu}_t = 0$ for all $t \in \llbracket T \rrbracket$. Throughout we set a target empirical coverage of $\alpha = 0.8$.
+For demonstration purposes we assume we have access to unbiased predictions $\hat{\mu}_t = 0$ for all $t \in \llbracket T \rrbracket$.
+Throughout we set the target empirical coverage to $\alpha = 0.8$.
 
@@
 
-::: {#fig-algo-aci fig-align=left fig-pos=H}
-```{.pseudocode}
+```pseudocode
+#| label: algo-aci
+#| html-indent-size: "1.2em"
+#| html-comment-delimiter: "//"
+#| html-line-number: true
+#| html-line-number-punc: ":"
+#| html-no-end: false
+#| pdf-placement: "htb!"
+#| pdf-line-number: true
+
+\begin{algorithm}
+\caption{Adaptive Conformal Inference}
 \begin{algorithmic}
@@
 \For{$t = 1, 2, \dots, T$}
-  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$.
+  \State \textbf{Output:} prediction interval $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$.
@@
 \end{algorithmic}
-```
-Adaptive Conformal Inference algorithm [@gibbs2021adaptive].
-:::
-
-The original ACI algorithm (@fig-algo-aci) adaptively adjusts the width of the prediction intervals in response to the observations [@gibbs2021adaptive].
+\end{algorithm}
+```
+
+The original ACI algorithm (@gibbs2021adaptive; @algo-aci) adaptively adjusts the width of the prediction intervals in response to the observations.
 The updating rule for the estimated radius can be derived as an online subgradient descent scheme.
@@
   \nabla L^\alpha(\theta, r) &= \begin{cases}
-    -\alpha, &\theta < r, \\
-    1 - \alpha, & \theta > r, \\
+    \{ -\alpha \}, &\theta < r, \\
+    \{ 1 - \alpha \}, & \theta > r, \\
     [-\alpha, 1 - \alpha], &\theta = r
@@
 ### Theoretical Guarantees
-Standard results for online subgradient descent yield the following regret bound with the use of the linear interval constructor, assuming that the true radii are bounded by $D$:
+The ACI algorithm has the following finite sample bound on the coverage error  [@gibbs2021adaptive]. For all $\gamma > 0$, 
+$$
+|\mathrm{CovErr}(T)| \leq \frac{D + \gamma}{\gamma T}.
+$$
+This result was originally shown for ACI with the choice of the quantile interval constructor, although it can also be extended to other interval constructors [@bhatnagar2023saocp, @feldman2023achieving].
+In addition, standard results for online subgradient descent yield the following regret bound with the use of the linear interval constructor, assuming that the true radii are bounded by $D$:
 $$
@@
 where the second inequality follows if the optimal choice of $\gamma = D/\sqrt{T}$ is used [@bhatnagar2023saocp].
-A finite sample bound on the coverage error can also be established [@gibbs2021adaptive]:
-$$
-|\mathrm{CovErr}(T)| \leq \frac{D + \gamma}{\gamma T}.
-$$
- Taken together, the theoretical results imply that while the coverage error is guaranteed to converge to zero for any choice of $\gamma$, achieving sublinear regret requires choosing $\gamma$ more carefully.
+Taken together, these theoretical results imply that while the coverage error is guaranteed to converge to zero for any choice of $\gamma$, achieving sublinear regret requires choosing $\gamma$ more carefully.
 This highlights the importance of both ways of assessing ACI algorithms: if we only focused on controlling the coverage error, we might not achieve optimal control of regret, leading to intervals that are not practically useful.
@@
 ### Tuning Parameters
-The main tuning parameter is the learning rate $\gamma$.
-Setting $\gamma$ too small will lead to intervals that do not adapt fast enough to distribution shifts, and setting $\gamma$ too large will lead to intervals with large oscillations as seen in @fig-aci.
+Therefore, the main tuning parameter is the learning rate $\gamma$.
+The theoretical bounds on the coverage error suggest setting a large $\gamma$ such that the coverage error decays quickly in $T$; however, in practice and setting $\gamma$ too large will lead to intervals with large oscillations as seen in @fig-aci.
 This is quantified in the path length, which increases significantly as $\gamma$ increases, even though the empirical coverage remains near the desired value of 80\%.
+On the other hand, setting $\gamma$ too small will lead to intervals that do not adapt fast enough to distribution shifts.
 Thus, choosing a good value for $\gamma$ is essential.
@@
 In practice, substantive prior information can be used to pick a reasonable starting value.
-By default, the `AdaptiveConformal` package sets $\theta_1 = 0$.
+By default, the `AdaptiveConformal` package sets $\theta_1 = \alpha$ when the quantile interval predictor is used, and $\theta_1 = 0$ otherwise, although in both cases the user can supply their own starting value.
+The behavior of the early prediction intervals in the examples (@fig-aci) is driven by the small number of residuals available, which makes the output of the quantile interval constructor sensitive to small changes in $\theta$.
+In practice, a warm-up period can be used before starting to produce prediction intervals so that the quantiles of the residuals are more stable.
 
@@
 #| fig-height: 4.5
-#| fig-cap: "Example prediction intervals from the ACI algorithm for different choices of learning rate $\\gamma$."
+#| fig-cap: "Example 80% prediction intervals from the ACI algorithm for different choices of learning rate $\\gamma$ and with $\\theta_1 = 0.8$.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 alpha <- 0.8
+gamma_grid <- c(0.016, 0.032, 0.064, 0.128)
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.01)),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.05)),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.1)),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "linear", gamma = 0.25))
-)
-
-coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
-path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))
-mean_width <- scales::number_format(accuracy = 0.01)(unlist(lapply(results, function(result) result$metrics$mean_width)))
+  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[1])),
+  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[2])),
+  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[3])),
+  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[4]))
+)
+
+coverage    <- format_coverage(extract_metric(results, "coverage"))
+path_length <- format_path_length(extract_metric(results, "path_length"))
 
@@
 ## Aggregated Adaptive Conformal Inference (AgACI)
-::: {#fig-algo-agaci fig-align=left fig-pos=H}
-```{.pseudocode}
+```pseudocode
+#| label: algo-agaci
+#| html-indent-size: "1.2em"
+#| html-comment-delimiter: "//"
+#| html-line-number: true
+#| html-line-number-punc: ":"
+#| html-no-end: false
+#| pdf-placement: "htb!"
+#| pdf-line-number: true
+
+\begin{algorithm}
+\caption{Aggregated Adaptive Conformal Inference}
 \begin{algorithmic}
   \State \textbf{Input:} candidate learning rates $(\gamma_k)_{1 \leq k \leq K }$, starting value $\theta_1$.
-  \State Initialize lower and upper BOA algorithms $\mathcal{B}^\ell := \texttt{BOA}(\alpha \leftarrow (1 - \alpha / 2))$ and $\mathcal{B}^u := \texttt{BOA}(\alpha \leftarrow (1 - (1 - \alpha)/2))$.
+  \State Initialize lower and upper BOA algorithms $\mathcal{B}^\ell := \texttt{BOA}(\alpha \leftarrow (1 - \alpha) / 2)$ and $\mathcal{B}^u := \texttt{BOA}(\alpha \leftarrow (1 - (1 - \alpha)/2))$.
   \For{$k = 1, \dots, K$}
@@
     \State Compute aggregated upper bound $\tilde{u}_t := \mathcal{B}^u((u^k_t : k \in \{ 1, \dots, K \}))$.
-    \State \textbf{Output} prediction interval $[\tilde{\ell}_t, \tilde{u}_t]$.
+    \State \textbf{Output:} prediction interval $[\tilde{\ell}_t, \tilde{u}_t]$.
     \State Observe $y_t$.
@@
 \end{algorithmic}
-```
-Aggregated Adaptive Conformal Inference (AgACI) algorithm [@zaffran2022agaci].
-:::
-
-The Aggregated ACI (AgACI; @fig-algo-agaci) algorithm solves the problem of choosing a learning rate for ACI by running multiple copies of the algorithm with different learning rates, and then combining separately the lower and upper interval bounds using an online aggregation of experts algorithm [@zaffran2022agaci].
-That is, one aggreation algorithm seeks to estimate the lower $(1-\alpha)/2$ quantile, and the other seeks to estimate the upper $1 - (1 - \alpha) / 2$ quantile.
-They experimented with multiple online aggregation algorithms, and found that they yielded similar results.
-Thus, we follow [@zaffran2022agaci] in using the Bernstein Online Aggregation (BOA) algorithm as implemented in the `opera` `R` package [@wintenberger2017boa; @opera2023].
+\end{algorithm}
+```
+
+The Aggregated ACI (AgACI; @algo-agaci) algorithm solves the problem of choosing a learning rate for ACI by running multiple copies of the algorithm with different learning rates, and then separately combining the lower and upper interval bounds using an online aggregation of experts algorithm [@zaffran2022agaci].
+That is, one aggregation algorithm seeks to estimate the lower $(1-\alpha)/2$ quantile, and the other seeks to estimate the upper $1 - (1 - \alpha) / 2$ quantile.
+@zaffran2022agaci experimented with multiple online aggregation algorithms, and found that they yielded similar results.
+Thus, we follow their lead in using the Bernstein Online Aggregation (BOA) algorithm as implemented in the `opera` `R` package [@wintenberger2017boa; @opera2023].
 BOA is an online algorithm that forms predictions for the lower (or upper) prediction interval bound as a weighted mean of the candidate ACI prediction interval lower (upper) bound, where the weights are determined by each candidate's past performance with respect to the quantile loss.
@@
 Rather, the upper and lower interval bounds from a set of candidate ACI algorithms are aggregated separately.
-Thus, theoretical results such as regret bounds in terms of $\theta$ similar to those for the other algorithms are not available.
+Thus, theoretical results such as regret bounds similar to those for the other algorithms are not available.
 It would be possible, however, to establish regret bounds for the pinball loss applied separately to the lower and upper bounds of the prediction intervals.
@@
 As a basic check, we can look at the weights assigned to each of the learning rates.
-If large weights are given to the smallest (largest) learning rates, it is a sign that smaller (or larger) learning rates would perform well.
+If large weights are given to the smallest (largest) learning rates, it is a sign that smaller (or larger) learning rates may perform well.
 In addition each of the candidate ACI algorithms requires a starting value, which can be set to any value as discussed in the ACI section.
 @fig-agaci illustrates AgACI applied to the running example with two sets of learning grids.
-The first grid is the default, and the second grid is double the size, given by $\gamma_i = 2^i / 1000$ for $i = 0, 0.5, 1, \dots, 7$.
-For the first grid, we can see that for the lower bound AgACI assigns high weight to only one learning rate ($\gamma = 0.032$).
-Based on this observation, as a sensitivity check, we reran the algorithm with the second, denser learning grid, which yields weights that are less concentrated on a single learning rate.
-The output prediction intervals are similar, however, for both grid choices.
+The first grid is the default, and the second grid includes the additional value $\gamma = \{ 0.0005 \}$.
+For the first grid, we can see that for the lower bound AgACI assigns high weight to the lowest learning rate ($\gamma = 0.001$).
+Based on this observation, as a sensitivity check, we reran the algorithm with the second learning grid, which yields weights that are less concentrated on a single learning rate.
+The output prediction intervals are similar, although slightly smoother in the second grid, reflecting the lack of distributional shift in the simulated data.
 
@@
 #| fig-height: 4.5
-#| fig-cap: "Aggregated Adaptive Conformal Inference (AgACI) using the linear interval construction function, starting values $\\theta_1 = 0$, and two different learning rate grids, where the second grid is double the size of the first."
+#| fig-cap: "Example 80% prediction intervals from the AgACI algorithm with starting values $\\theta_1 = 0.8$ and two different learning rate grids, where the second grid is double the size of the first.
+In the left column, blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 grid1 <- 2^(seq(0, 7, 1)) / 1e3
-grid2 <- 2^(seq(0, 7, 0.5)) / 1e3
+grid2 <- 2^(seq(-1, 5, 1)) / 1e3
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear", gamma_grid = grid1)),
-  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "linear", gamma_grid = grid2))
-)
-
-coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
-path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))
-mean_width <- scales::number_format(accuracy = 0.01)(unlist(lapply(results, function(result) result$metrics$mean_width)))
+  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid1, theta0 = 0.8)),
+  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid2, theta0 = 0.8))
+)
+
+coverage    <- format_coverage(extract_metric(results, "coverage"))
+path_length <- format_path_length(extract_metric(results, "path_length"))
 
@@
 
-  plot_agaci_weights(results[[i]], main = "Weights", legend = "topleft")
+  plot_agaci_weights(results[[i]], main = "Final Aggregation Weights", legend = "topright")
 }
@@
 ## Fully Adaptive Conformal Inference (FACI)
-::: {#fig-algo-faci fig-align=left fig-pos=H}
-```{.pseudocode}
+```pseudocode
+#| label: algo-faci
+#| html-indent-size: "1.2em"
+#| html-comment-delimiter: "//"
+#| html-line-number: true
+#| html-line-number-punc: ":"
+#| html-no-end: false
+#| pdf-placement: "htb!"
+#| pdf-line-number: true
+
+\begin{algorithm}
+\caption{Fully Adaptive Conformal Inference}
 \begin{algorithmic}
-\State \textbf{Input:} starting value $\theta_1$, candidate learning rates $() \gamma_k )_{1 \leq k \leq K }$, parameters $\sigma, \eta$.
+\State \textbf{Input:} starting value $\theta_1$, candidate learning rates $(\gamma_k)_{1 \leq k \leq K }$, parameters $\sigma, \eta$.
 \For{$k = 1, \dots, K$}
@@
   \State Set $\theta_t = \sum_{k=1}^K \theta_t^k p_t^k$.
-  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$.
+  \State \textbf{Output:} prediction interval $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$ and compute $r_t$.
@@
 \end{algorithmic}
-```
-Fully Adaptive Conformal Inference (FACI) algorithm [@gibbs2022faci].
-:::
-
-The Fully Adaptive Conformal Inference method also aggregates predictions from multiple copies of ACI run with different learning rates, but differs from AgACI in that it directly aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci].
-FACI uses the aggregation scheme proposed in @gradu2022adaptive.
+\end{algorithm}
+```
+
+The Fully Adaptive Conformal Inference (FACI; @algo-faci) algorithm was developed by the authors of the original ACI algorithm in part to address the issue of how to choose the learning rate parameter $\gamma$.
+In this respect the goal of the algorithm is similar to that of AgACI, although it is achieved slightly differently.
+FACI also aggregates predictions from multiple copies of ACI run with different learning rates, but differs in that it directly aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci] using an exponential reweighting scheme [@gradu2022adaptive].
+As opposed to AgACI, this construction allows for more straightforward development of theoretical guarantees on the algorithm's performance, because the upper and lower bounds of the intervals are not aggregated separately.
 
 ### Theoretical Guarantees
-FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]:
-$$
-\begin{aligned}
-  \mathrm{SAReg}(T, k) \leq \widetilde{\mathcal{O}}(D^2 / \gamma + \gamma k).
-\end{aligned}
-$$
-If $k$ is fixed a-priori, then choosing $\gamma = D/\sqrt{k}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{k})$ (for a single choice of $k$).
+FACI was originally proposed with the choice of the quantile interval constructor.
+FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]: for all $\gamma > 0$ and subperiod lengths $m$,
+$$
+\begin{aligned}
+  \mathrm{SAReg}(T, m) \leq \widetilde{\mathcal{O}}(D^2 / \gamma + \gamma m).
+\end{aligned}
+$$
+If $m$ is fixed a-priori, then choosing $\gamma = D/\sqrt{m}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{m})$ (for a single choice of $m$).
 Practically, this result implies that, if we know in advance the time length for which we would like to control the regret, it is possible to choose an optimal tuning parameter value.
@@
 This is a common trick used in the literature as it facilitates theoretical analysis.
-In practice, the authors comment that this randomized version of FACI and the deterministic version lead very similar results.
-The coverage error result also assumes that the hyperparameters can change over time: that is, we have $\eta_{t}$ and $\sigma_t$, rather than $\eta$ and $\sigma$.
-The coverage error then has the following bound (Theorem 3.2, @gibbs2022faci), where $\gamma_{\min}$ and $\gamma_{\max}$ are the smallest and largest learning rates in the grid, respectively:
+In practice, the authors comment that this randomized version of FACI and the deterministic version lead to very similar results.
+The coverage error result also assumes that the hyperparameters can change over time: that is, we have $t$-specific $\eta_{t}$ and $\sigma_t$, rather than fixed $\eta$ and $\sigma$.
+The coverage error then has the following bound [@gibbs2022faci], where $\gamma_{\min}$ and $\gamma_{\max}$ are the smallest and largest learning rates in the grid, respectively:
 $$
@@
 Thus, if $\eta_t$ and $\sigma_t$ both converge to zero as $t \to \infty$, then the coverage error will also converge to zero.
-In addition, under mild distributional assumptions the authors provide a type of short-term coverage error bound for arbitrary time spans, for which we refer to Proposition 3.2 [@gibbs2022faci].
+In addition, under mild distributional assumptions the authors provide a type of short-term coverage error bound for arbitrary time spans, for which we refer to [@gibbs2022faci].
+
+We note one additional result established by @gibbs2022faci on a slightly different regret bound in terms of the pinball loss, as it informs the choice of tuning parameters.
+Let $\gamma_{\mathrm{max}} = \max_{1 \leq k \leq K} \gamma_k$ be the largest learning rate in the grid and assume that $\gamma_1 < \gamma_2 < \cdots < \gamma_K$ with $\gamma_{k+1}/\gamma \leq 2$ for all $1 \leq k < K$.
+Then, for any interval $I = [r, s] \subseteq \llbracket T \rrbracket$ and any sequence $\theta_r^*, \dots, \theta_s^*$, under the assumption that $\gamma_k \geq \sqrt{1 + 1 / |I|}$,
+$$
+\begin{aligned}
+  \frac{1}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)] - \frac{1}{|I|} \sum_{t=r}^s L^\alpha(\theta_t, \theta_t^*) \leq& \frac{\log(k / \sigma) + 2\sigma|I|}{\eta |I|} + \frac{\eta}{|I|} \sum_{t=r}^s \mathbb{E}[L^\alpha(\theta_t, r_t)^2] \\
+  &+ 2\sqrt{3}(1 + \gamma_{\mathrm{max}})^2 \max\left\{ \sqrt{\frac{\sum_{t=r+1}^s |\theta_t^* - \theta_{t-1}^*| + 1}{|I|}}, \gamma_1 \right\},
+\end{aligned}
+$$
+where the expectation is over the randomness in the randomized version of the algorithm.
+Here the time interval $I$ (with length $|I|$) is comparable to the time period length $m$ for the strongly adaptive regret.
+The parameter $|I|$, the time interval of interest for which we would like to control, can be chosen arbitrarily.
 
 ### Tuning parameters
-The parameter $|I|$, the time interval of interest, can be chosen arbitrarily.
+The recommended settings for the tuning parameters depend on choosing a time interval length $|I|$ for which we would like to control the pinball loss. The choice of $|I|$ can be chosen arbitrarily.
 For the tuning parameter $\sigma$, the authors suggest the optimal choice $\sigma = 1 / (2 |I|)$.
@@
 $$
-Note that this choice is optimal only for the conformal interval constructor, for which $\theta_t$ is a quantile of previous conformity scores.
+Note that this choice is optimal only for the quantile interval constructor, for which $\theta_t$ is a quantile of previous nonconformity scores.
 As an alternative, the authors point out that $\eta$ can be learned in an online fashion using the update rule
@@
 Both ways of choosing $\eta$ led to very similar results in the original author's empirical studies.
-In the `AdaptiveConformal` package, the first approach is used when the conformal interval construction function is chosen, and the latter approach for the linear interval construction function.
-
-@fig-faci-linear illustrates FACI with a linear interval construction function and with the learning rate grid $\gamma \in \{ 0.05, 0.1, 0.15, 0.2, 0.25 \}$.
+In our proposed `AdaptiveConformal` package, the first approach is used when the quantile interval construction function is chosen, and the latter approach for the linear interval construction function.
+
+@fig-faci-example illustrates FACI with the quantile interval construction function and with the learning rate grid $\gamma \in \{ 0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128 \}$.
 
 ```{r faci_example_plot, echo = FALSE}
-#| label: fig-faci-linear
+#| label: fig-faci-example
 #| fig-height: 4.5
-#| fig-cap: "Example prediction intervals generated by the FACI algorithm."
-gamma_grid <- c(0.05, 0.25, 0.05)
-result <- aci(y, muhat, alpha = alpha, method = "FACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "linear"))
-coverage <- scales::percent_format()(result$metrics$coverage)
-path_length <- scales::number_format(accuracy = 0.1)(result$metrics$path_length)
+#| fig-cap: "Example 80% prediction intervals generated by the FACI algorithm with starting values $\\theta_1 = 0.8$.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
+gamma_grid <- c(0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128)
+result <- aci(y, muhat, alpha = alpha, method = "FACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal"))
+
+coverage    <- format_coverage(extract_metric(list(result), "coverage"))
+path_length <- format_path_length(extract_metric(list(result), "path_length"))
 
 plot(result, ylim = c(-0.9, 0.8), legend = FALSE)
-text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage) ), pos = 4)
-text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length)), pos = 4)
+text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[1]]) ), pos = 4)
+text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[1]])), pos = 4)
 ```
@@
 ## Scale-Free Online Gradient Descent (SF-OGD)
-::: {#fig-algo-sfogd fig-align=left fig-pos=H}
-```{.pseudocode}
+```pseudocode
+#| label: algo-sfogd
+#| html-indent-size: "1.2em"
+#| html-comment-delimiter: "//"
+#| html-line-number: true
+#| html-line-number-punc: ":"
+#| html-no-end: false
+#| pdf-placement: "htb!"
+#| pdf-line-number: true
+
+\begin{algorithm}
+\caption{Scale-Free Online Gradient Descent}
 \begin{algorithmic}
@@
 \For{$t = 1, 2, \dots, T$}
-  \State \textbf{Output} prediction interval $\widehat{C}_t(\theta_t)$. 
+  \State \textbf{Output:} prediction interval $\widehat{C}_t(\theta_t)$. 
   \State Observe $y_t$ and compute $r_t$. 
@@
 \end{algorithmic}
-```
-Scale-Free Online Gradient Descent (SF-OGD) algorithm [@orabona2018sfogd; @bhatnagar2023saocp].
-:::
-
-Scale-Free Online Gradient Descent is a general algorithm for online learning proposed by @orabona2018sfogd.
+\end{algorithm}
+```
+
+Scale-Free Online Gradient Descent (SF-OGD; @algo-sfogd) is a general algorithm for online learning proposed by @orabona2018sfogd.
 The algorithm updates $\theta_t$ with a gradient descent step where the learning rate adapts to the scale of the previously observed gradients.
@@
 ### Theoretical Guarantees
-The SF-OGD algorithm has the following regret bound, which is called an _anytime regret bound_ because it holds for all $t \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
+The SF-OGD algorithm with linear interval constructor has the following regret bound, which is called an _anytime regret bound_ because it holds for all $t \in \llbracket T \rrbracket$ [@bhatnagar2023saocp].
+For any $\gamma > 0$,
 $$
@@
 #| fig-height: 4.5
-#| fig-cap: "Example prediction intervals generated by the SF-OGD algorithm with different values of the maximum radius tuning parameter $D$."
+#| fig-cap: "Example 80% prediction intervals generated by the SF-OGD algorithm with different values of the maximum radius tuning parameter $D$.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 alpha <- 0.8
@@
 
-coverage <- scales::percent_format()(unlist(lapply(results, function(result) result$metrics$coverage)))
-path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(results, function(result) result$metrics$path_length)))
+coverage    <- format_coverage(extract_metric(results, "coverage"))
+path_length <- format_path_length(extract_metric(results, "path_length"))
 
@@
 ## Strongly Adaptive Online Conformal Prediction (SAOCP)
-::: {#fig-algo-saocp fig-align=left fig-pos=H}
-```{.pseudocode}
+```pseudocode
+#| label: algo-saocp
+#| html-indent-size: "1.2em"
+#| html-comment-delimiter: "//"
+#| html-line-number: true
+#| html-line-number-punc: ":"
+#| html-no-end: false
+#| pdf-placement: "htb!"
+#| pdf-line-number: true
+
+\begin{algorithm}
+\caption{Strongly Adaptive Online Conformal Prediction}
 \begin{algorithmic}
@@
   \State Initialize expert $\mathcal{A}_t = \texttt{SF-OGD}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma, \theta_1 \leftarrow \theta_{t-1})$, set weight $w_t^t = 0$.
-  \State Compute active set $\mathrm{Active}(t) = \{ i \in \llbracket T \rrbracket : t - L(i) < i \leq t \}$.
+  \State Compute active set $\mathrm{Active}(t) = \{ i \in \llbracket T \rrbracket : t - L(i) < i \leq t \}$ (see below for definition of $L(t)$).
   \State Compute prior probability $\pi_i \propto i^{-2} (1 + \lfloor \log_2 i \rfloor )^{-1} \mathbb{I}[i \in \mathrm{Active}(t)]$.
@@
   \State Set $\theta_t = \sum_{i \in \mathrm{Active}(t)} p_i \theta_t^i$ (for $t \geq 2$), and $\theta_t = 0$ for $t = 1$.
-  \State \textbf{Output} prediction set $\widehat{C}_t(\theta_t)$.
+  \State \textbf{Output:} prediction set $\widehat{C}_t(\theta_t)$.
   \State Observe $y_t$ and compute $r_t$. 
@@
 \end{algorithmic}
-```
-Strongly Adaptive Online Conformal Prediction (SAOCP) algorithm [@bhatnagar2023saocp]. 
-:::
-
-The Strongly Adaptive Online Conformal Prediction algorithm maintains a library of candidate online learning algorithms that generate prediction intervals which are then aggregated using a meta-algorithm [@bhatnagar2023saocp].
+\end{algorithm}
+```
+
+The Strongly Adaptive Online Conformal Prediction (SAOCP; @algo-saocp) algorithm was proposed as an improvement over the extant ACI algorithms in that it features stronger theoretical guarantees.
+SAOCP works similarly to AgACI and FACI in that it maintains a library of candidate online learning algorithms that generate prediction intervals which are then aggregated using a meta-algorithm [@bhatnagar2023saocp].
 The candidate algorithm was chosen to be SF-OGD, although any algorithm that features anytime regret guarantees can be chosen.
-As opposed to AgACI and FACI, in which each candidate has a different learning rate, here each candidate has the same learning rate but is active over a different interval.
-At each time point, a new expert is instantiated which is active over a finite lifetime.
+As opposed to AgACI and FACI, in which each candidate has a different learning rate but is always able to contribute to the final prediction intervals, here each candidate has the same learning rate but only has positive weight over a specific interval of time.
+New candidate algorithms are continually being spawned in order that, if the distribution shifts rapidly, the newer candidates will be able to react quickly and receive positive weight.
+Specifically, at each time point, a new expert is instantiated which is active over a finite ``lifetime".
 Define the _lifetime_ of an expert instantiated at time $t$ as
@@
 $$
-where $g \in \mathcal{Z}^*$ is a _lifetime multiplier_ parameter. The active experts are weighted according to their empirical performance with respect to the pinball loss function.
+where $g \in \mathbb{Z}^*$ is a _lifetime multiplier_ parameter.
+The active experts are weighted according to their empirical performance with respect to the pinball loss function.
+The authors show that this construction results in intervals that have strong regret guarantees.
 
 ### Theoretical Guarantees
-The following bound for the strongly adaptive regret holds for all lengths $k \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
-$$
-\begin{aligned}
-  \mathrm{SAReg}(T, k) \leq 15 D \sqrt{k(\log T + 1)} \leq \tilde{\mathcal{O}}(D \sqrt k).
-\end{aligned}
-$$
-It should be emphasized that this regret bounds holds simultaneously across all $k$, as opposed to FACI, where a similar bound holds only for a single $k$.
+The theoretical results were established for SAOCP using the linear interval constructor.
+The following bound for the strongly adaptive regret holds for all subperiod lengths $m \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
+$$
+\begin{aligned}
+  \mathrm{SAReg}(T, m) \leq 15 D \sqrt{m(\log T + 1)} \leq \tilde{\mathcal{O}}(D \sqrt m).
+\end{aligned}
+$$
+It should be emphasized that this regret bounds holds simultaneously across all $m$, as opposed to FACI, where a similar bound holds only for a single $m$.
 A bound on the coverage error of SAOCP has also been established  as:
@@
 #| fig-height: 4.5
-#| fig-cap: "Example of prediction intervals generated by the SAOCP algorithm with different values of the maximum radius parameter $D$."
+#| fig-cap: "Example 80% prediction intervals generated by the SAOCP algorithm with different values of the maximum radius parameter $D$.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 alpha <- 0.8
@@
 
-# Simulation Study {#sec-simulations}
-We present two empirical studies in order to compare the performance of the ACI algorithms in simple simulated datasets.
-For both simulations we targeted empirical coverage of $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.
+# Simulation Studies {#sec-simulations}
+We present two empirical studies in order to compare the performance of the AgACI, FACI, SF-OGD, and SAOCP algorithms applied to simple simulated datasets.
+The original ACI algorithm was not included as it is not clear how to set the tuning rate $\gamma$, which can have a large effect on the resulting intervals.
+For both simulations we set the targeted empirical coverage to $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.
+For each algorithm, we chose the interval constructor that was used in its original presentation (see @tbl-aci).
 
 ## Time series with ARMA errors
-In this simulation we reproduce the setup described in @zaffran2022agaci (itself based on that of @friedman1983). The time series values $y_t$ for $t = 1, \dots, 600$ are simulated according to
+In this simulation we reproduce the setup described in @zaffran2022agaci (itself based on that of @friedman1983).
+The time series values $y_t$ for $t \in \llbracket T \rrbracket$ ($T = 600$) are simulated according to
 $$
@@
 $$
-where $X_{t,i}$, $i = 1, \dots, 6$ are uniformly distributed on $[0, 1]$. The noise terms $\epsilon_t$ are generated according to an ARMA(1, 1) process:
+where $X_{t,i}$, $i = 1, \dots, 6$, $t \in \llbracket T \rrbracket$ are independently uniformly distributed on $[0, 1]$ and the noise terms $\epsilon_t$ are generated according to an ARMA(1, 1) process:
 $$
@@
 We set $\psi$ and $\theta$ jointly to each value in $\{ 0.1, 0.8, 0.9, 0.95, 0.99 \}$ to simulate time series with increasing temporal dependence.
-The innovation variance was set to $\sigma^2 = (1 - \psi^2) / (1 + 2\psi \xi + \xi^2)$.
+The innovation variance was set to $\sigma^2 = (1 - \psi^2) / (1 + 2\psi \xi + \xi^2)$ (to ensure that the process has constant variance).
 For each setting, 25 simulated datasets were generated.
@@
   D <- max(abs(data$y - preds)[200:249])
-  gamma <- ifelse(method == "ACI", D / sqrt(length(250:nrow(data))), D / sqrt(3))
-
-  parameters <- list(interval_constructor = "linear", D = D, gamma = gamma, gamma_grid = seq(0.1, 1, 0.1))
-  aci(data$y[250:nrow(data)], preds[250:nrow(data)], method = method, alpha = alpha, parameters = parameters)
+  gamma <- D / sqrt(3)
+
+  interval_constructor = case_when(
+    method == "AgACI" ~ "conformal",
+    method == "FACI" ~ "conformal",
+    method == "SF-OGD" ~ "linear",
+    method == "SAOCP" ~ "linear"
+  )
+
+  if(interval_constructor == "linear") {
+    gamma_grid = seq(0.1, 1, 0.1)
+  }
+  else {
+    gamma_grid <- c(0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128)
+  }
+
+  parameters <- list(
+    interval_constructor = interval_constructor, 
+    D = D, 
+    gamma = gamma, 
+    gamma_grid = gamma_grid
+  )
+
+  aci(
+    data$y[250:nrow(data)], 
+    preds[250:nrow(data)], 
+    method = method, 
+    alpha = alpha, 
+    parameters = parameters
+  )
 }
@@
   alpha = c(0.8, 0.9, 0.95),
-  method = c("ACI", "AgACI", "SF-OGD", "SAOCP", "FACI")
+  method = c("AgACI", "SF-OGD", "SAOCP", "FACI")
 )
@@
 
-The coverage error, mean interval width, and path lengths of each of the algorithms for $\alpha = 0.9$ are shown in @fig-simulation-one-results (results for $\alpha \in \{ 0.8, 0.95 \}$ were similar and are available in the appendix).
+The coverage errors, mean interval widths, and path lengths of each of the algorithms for $\alpha = 0.9$ are shown in @fig-simulation-one-results (results for $\alpha \in \{ 0.8, 0.95 \}$ were similar and are available in the appendix).
 All methods achieved near optimal empirical coverage, although SAOCP tended to slightly undercover.
-The mean interval widths were similar across methods, although again SAOCP had slightly shorter intervals (as expected given its tendency to under cover).
+The mean interval widths were similar across methods, although again SAOCP had slightly shorter intervals (as could be expected given its tendency to undercover).
 The path length of SAOCP was larger than any of the other methods.
 To investigate why, @fig-simulation-one-example plots $w_t - w_{t-1}$, the difference in interval width between times $t-1$ and $t$, for each method in one of the simulations.
-The interval widths for ACI, AgACI, and FACI change slowly relative to those for SF-OGD and SAOCP.
-For SAOCP, we can see the interval widths have larger fluctuations than the other methods, explaining its higher path width.
+The interval widths for AgACI and FACI change slowly relative to those for SF-OGD and SAOCP.
+For SAOCP, we can see the interval widths have larger fluctuations than for the other methods, explaining its higher path width.
 
@@
 #| label: fig-simulation-one-results
-#| fig-height: 5.5
+#| fig-height: 6.5
 #| fig-cap: "Coverage errors, mean interval widths, and path lengths for the first simulation study with target coverage $\\alpha = 0.9$."
@@
 #| label: fig-simulation-one-example
-#| fig-height: 6
+#| fig-height: 5
 #| fig-cap: "Difference in successive interval widths ($w_t - w_{t-1}$) from an illustrative simulation from the first simulation study."
@@
 
-par(mfrow = c(3, 2), mar = c(3, 4, 2, 1))
-for(i in 1:5) {
+par(mfrow = c(2, 2), mar = c(3, 4, 2, 1))
+for(i in 1:4) {
   plot(
@@
 The setup is quite simple in order to probe the basic performance of the methods in response to distribution shift.
-As a baseline, we simulate time series according to
+As a baseline, we simulate time series of independent data with
 $$
@@
 $$
-for $t = 1, \dots, T$ and $T = 500$. In the second type of time series, the variance of the observations increases halfway through the time series:
+for all $t \in \llbracket T \rrbracket$ ($T = 500$). In the second type of time series, the observations are still independent but their variance increases halfway through the time series:
 $$
@@
 y_t &\sim N(0, \sigma_t^2), \\
-\sigma_t^2 &= 0.2 + 0.5 \mathbb{I}[t > 250]
-\end{aligned}
-$$
-In each case, the ACI algorithms are provided with the unbiased predictions $\hat{\mu}_t = 0$, $t = 1, \dots, T$. 50 simulated datasets were generated for each type of time series.
+\sigma_t &= 0.2 + 0.5 \mathbb{I}[t > 250].
+\end{aligned}
+$$
+In each case, the ACI algorithms are provided with the unbiased predictions $\hat{\mu}_t = 0$, $t \in \llbracket T \rrbracket$. Fifty simulated datasets were generated for each type of time series.
 
@@
   if(!is.null(p)) p()
+
+  interval_constructor = case_when(
+    method == "AgACI" ~ "conformal",
+    method == "FACI" ~ "conformal",
+    method == "SF-OGD" ~ "linear",
+    method == "SAOCP" ~ "linear"
+  )
+
+  if(interval_constructor == "linear") {
     D <- max(abs(data$y - data$muhat)[1:50])
-
-  if(method == "ACI") {
-    gamma <- D / sqrt(nrow(data))
   }
   else {
+    D <- 1
+  }
+
   gamma <- D / sqrt(3)
+
+  if(interval_constructor == "linear") {
+    gamma_grid <- seq(0.1, 2, 0.1)
+  }
+  else {
+    gamma_grid <- c(0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128)
   }
@@
   parameters <- list(
-    interval_constructor = "linear", 
+    interval_constructor = interval_constructor, 
     D = D, 
     gamma = gamma, 
-    gamma_grid = seq(0.1, 2, 0.1)
+    gamma_grid = gamma_grid
   )
@@
   N = 500,
-  method = c("ACI", "AgACI", "SF-OGD", "SAOCP", "FACI")
+  method = c("AgACI", "SF-OGD", "SAOCP", "FACI"),
 ) %>%
@@
 
-The coverage error, mean interval widths, and path lengths of each of the algorithms are shown in @fig-simulation-two-results.
+The coverage error, mean path length, and mean interval widths of the algorithms are summarized in @fig-simulation-two-joint (an alternative plot is included in the appendix as @fig-simulation-two-results).
 The coverage error of all the algorithms is near the desired value in the absence of distribution shift.
-On the contrary, all of the algorithms with the exception of ACI undercover when there is distributional shift.
-An illustrative example of prediction intervals generated by each ACI method for one of the simulated time series with distribution shift is shown in @fig-simulation-two-example.
-
-```{r simulation_two_plot}
-#| label: fig-simulation-two-results
-#| fig-height: 9
-#| fig-cap: "Coverage error, mean interval width, and path length for $\\alpha = 0.8, 0.9, 0.95$ and simulations with and without distributional shift."
-
-simulation_two_plot(simulation_study2$results)
+On the contrary, all of the algorithms except AgACI and FACI undercover when there is distributional shift.
+SAOCP tends to have higher average path lengths than the other methods.
+An illustrative example of prediction intervals generated by each method for one of the simulated time series with distribution shift is shown in @fig-simulation-two-example.
+The SAOCP prediction intervals in the example before the distribution shift are more jagged than those produced by the other methods, which illustrates why SAOCP may have longer path lengths.
+
+```{r simulation_two_plot_joint, echo=FALSE}
+#| label: fig-simulation-two-joint
+#| fig-cap: "Mean interval width vs coverage error (top) and Mean Path Length vs.
+coverage error (bottom) for the second simulation study.
+The error bars represent the 10% to 90% quantiles of the metrics over the simulation datasets."
+#| fig-height: 6
+simulation_two_joint_plot(simulation_study2$results)
 ```
@@
 #| fig-height: 6
-#| fig-cap: "Example prediction intervals (target coverage $\\alpha = 0.9$) from the second simulation study of time series with distributional shift."
+#| fig-cap: "Example prediction intervals (target coverage $\\alpha = 0.9$) from the second simulation study of time series with distributional shift, in which the shift occurs at time 250.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 fits <- simulation_study2$example_fits
 
-coverage <- scales::percent_format()(unlist(lapply(fits$fit, function(result) result$metrics$coverage)))
-path_length <- scales::number_format(accuracy = 0.1)(unlist(lapply(fits$fit, function(result) result$metrics$path_length)))
-
-par(mfrow = c(3, 2), mar = c(3, 3, 2, 1))
-for(i in 1:5) {
+coverage    <- format_coverage(extract_metric(fits$fit, "coverage"))
+path_length <- format_path_length(extract_metric(fits$fit, "path_length"))
+
+par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
+for(i in 1:4) {
   plot(fits$fit[[i]], legend = FALSE, main = fits$method[[i]], index = 51:500)
@@
 # Case Study: Influenza Forecasting {#sec-case-study}
-Influenza is a highly infectious disease that is estimated to infect ~1 billion individuals each year around the world [@krammer2018influenza].
+Influenza is a highly infectious disease that is estimated to infect approximately one billion individuals each year around the world [@krammer2018influenza].
 Influenza incidence in temperate climates tends to follow a seasonal pattern, with the highest number of infections during what is commonly referred to as the \textit{flu season} [@lofgren2007influenza].
 Accurate forecasting of influenza is of significant interest to aid in public health planning and resource allocation.
-To investigate the accuracy of influenza forecasts, The US Centers for Disease Control (CDC) initiated a challenge, referred to as FluSight, in which teams from multiple institutions submitted weekly forecasts of influenza incidence [@biggerstaff2016flusight].
+To investigate the accuracy of influenza forecasts, the US Centers for Disease Control (CDC) initiated a challenge, referred to as FluSight, in which teams from multiple institutions submitted weekly forecasts of influenza incidence [@biggerstaff2016flusight].
 @reich2019influenza evaluated the accuracy of the forecasts over seven flu seasons from 2010 to 2017.
@@
 For this case study, we focus on national (US) one-week ahead forecasts of weighted influenza-like illness (wILI), which is a population-weighted percentage of doctors visits where patients presented with influenza-like symptoms [@biggerstaff2016flusight].
-The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, including both mechanistic and statistical approaches [@tushar2018flusightnetwork].
+The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, from both mechanistic and statistical viewpoints [@tushar2018flusightnetwork, @tushar2019flusight].
 For our purposes, we treat the way the forecasts were produced as a black box.
 
-Formally, let $y_{t}$, $t \in \llbracket T \rrbracket$ be the observed national wILI at time $t$, and let $\hat{\mu}_{j,t}$, $j = \llbracket J \rrbracket$ be the one-week ahead forecast of the wILI from model $j$ at time $t$.
+Formally, let $y_{t}$, $t \in \llbracket T \rrbracket$ be the observed national wILI at time $t$, and let $\hat{\mu}_{j,t}$, $j \in \llbracket J \rrbracket$, be the one-week ahead forecast of the wILI from model $j$ at time $t$.
 Two of the original 21 forecasting methods were excluded from this case study due to poor predictive performance (\texttt{Delphi\_Uniform} and \texttt{CUBMA}).
+In addition, six methods had identical forecasts (\texttt{CU\_EAKFC\_SIRS}, \texttt{CU\_EKF\_SEIRS}, \texttt{CU\_EKF\_SIRS}, \texttt{CU\_RHF\_SEIRS}, \texttt{CU\_RHF\_SIRS}), and therefore we only included one (\texttt{CU\_EAKFC\_SEIRS}) in the analysis.
 The ACI methods were then applied to the log-observations and log-predictions, where the log-transformation was used to constrain the final prediction intervals to be positive.
@@
 The ACI algorithms target prediction intervals with coverage of $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.
+As in the simulation study, we used the interval constructor corresponding to the original presentaiton of each algorithm (see @tbl-aci).
 
 ```{r case_study, cache = TRUE}
-raw_data <- read_csv("data/point_ests.csv", show_col_types = FALSE)
+raw_data <- read_csv("https://raw.githubusercontent.com/FluSightNetwork/cdc-flusight-ensemble/master/scores/point_ests.csv", show_col_types = FALSE)
 
@@
 
-  gamma <- ifelse(method == "ACI", D / sqrt(nrow(data)), D / sqrt(3))
+  interval_constructor = case_when(
+    method == "AgACI" ~ "conformal",
+    method == "FACI" ~ "conformal",
+    method == "SF-OGD" ~ "linear",
+    method == "SAOCP" ~ "linear"
+  )
+
+  gamma <- D / sqrt(3)
+
+  if(interval_constructor == "linear") {
+    gamma_grid = seq(0.1, 1, 0.1)
+  }
+  else {
+    gamma_grid <- c(0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128)
+  }
 
   parameters <- list(
-    interval_constructor = "linear", 
+    interval_constructor = interval_constructor,
     D = D, 
     gamma = gamma, 
-    gamma_grid = seq(0.01, 1, 0.05)
+    gamma_grid = gamma_grid
   )
@@
     Location == "US National", 
-    !(model_name %in% c("Delphi_Uniform", "CUBMA"))
+    !(model_name %in% c("Delphi_Uniform", "CUBMA", "CU_EAKFC_SIRS", "CU_EKF_SEIRS", "CU_EKF_SIRS", "CU_RHF_SEIRS", "CU_RHF_SIRS"))
   ) %>%
@@
   analysis_data, 
-  tibble(method = c("ACI", "AgACI", "FACI", "SF-OGD", "SAOCP")), 
+  tibble(method = c("AgACI", "FACI", "SF-OGD", "SAOCP")), 
   tibble(alpha = c(0.8, 0.9, 0.95))
@@
 The coverage errors, mean interval widths, and path lengths of the prediction intervals for each of the underlying forecast models is shown in @fig-case-study-metrics.
-In almost all cases the absolute coverage error was less than $0.1$.
+In all cases the absolute coverage error was less than $0.1$.
 SF-OGD performed particularly well, with coverage errors close to zero for all forecasting models.
 Interval widths were similar across methods, with SAOCP slightly shorter.
-Path Lengths were shorter for ACI and SF-OGD and longer for SAOCP.
+Path Lengths were shorter for AgACI and FACI and longer for SAOCP.
 
@@
 #| label: fig-case-study-metrics
-#| fig-height: 6
-#| fig-cap: "Coverage error, mean interval width, and path lengths of prediction intervals generated with each ACI method based on forecasts from each of the 19 underlying influenza forecasting models."
+#| fig-height: 6.5
+#| fig-cap: "Coverage errors, mean interval widths, and path lengths of prediction intervals generated with each ACI method based on forecasts from each of the 19 underlying influenza forecasting models."
 
@@
 As an illustrative example, in @fig-case-study-example we plot the point forecasts from one of the forecasting models (based on SARIMA with no seasonal differencing) and the associated ACI-generated 90\% prediction intervals for each season from 2011-2017.
+In general, in this practical setting all of the ACI algorithms yield quite similar prediction intervals.
 Interestingly, the forecasts in 2011-2012 underpredicted the observations for much of the season.
@@
 The results of our simulations and case study show that, when tuning parameters are chosen well, Adaptive Conformal Inference algorithms yield well-performing prediction intervals.
-On the other hand, poor choice of tuning parameters can lead to intervals of very low utility.
-Furthermore, in some cases the prediction intervals may appear to perform well with respect to metrics like the empirical coverage error, while simultaneously being practically useless.
-The original ACI algorithm illustrates this phenomena: too small a value of its learning rate $\gamma$ yields prediction intervals that are not reactive enough, and too large yields intervals that change too fast.
+On the contrary, poor choice of tuning parameters can lead to intervals of low utility.
+Furthermore, in some cases the prediction intervals may appear to perform well with respect to metrics like the empirical coverage error, while simultaneously being useless in practice.
+The original ACI algorithm illustrates this phenomenon: too small a value of its learning rate $\gamma$ yields prediction intervals that are not reactive enough, while too large a value yields intervals that change too fast.
 In both cases, the empirical coverage may appear well-calibrated, while the prediction intervals will not be useful.
 Thus, the core challenge in designing an ACI algorithm is in finding an optimal level of reactivity for the prediction intervals.
-As users of these algorithms, the challenge is in finding values for the tuning parametesr that avoid pathological behaviors.
+As users of these algorithms, the challenge is in finding values for the tuning parameters that avoid pathological behaviors.
 
@@
 Our results show the algorithms can perform well in multiple difficult scenarios.
-However, the overall effect of these approaches is to shift the problem to a higher level of abstraction: we still need to set tuning parameters that control the amount of reactivity, we just do so at a higher level than the original ACI algorithm.
+However, the overall effect of these approaches is to shift the problem to a higher level of abstraction: we still need to set tuning parameters that control the amount of reactivity, but do so at a higher level than the original ACI algorithm.
 It is desirable that these tuning parameters be easily interpretable, with simple strategies available for setting them.
@@
 In practice, it is possible the second simulation study does not accurately reflect real-world scenarios.
-Indeed, the benchmarks presented in [@bhatnagar2023saocp] using the datasets from the M4 competition, and using point predictions generated by diverse prediction algorithms, found that ACI algorithms exhibited good performance in terms of empirical coverage.
-Regardless, our recommendation for future papers in this line of research is to include simulation studies for simple distributional shift scenarios as a benchmark.
+Indeed, the benchmarks presented in @bhatnagar2023saocp using the datasets from the M4 competition [@makridakis2020m4], and using point predictions generated by diverse prediction algorithms, found that ACI algorithms exhibited good performance in terms of empirical coverage.
+Nevertheless, our recommendation for future papers in this line of research is to include simulation studies for simple distributional shift scenarios as a benchmark.
 
@@
 If the predictions are biased, for example, then the prediction intervals may be able to achieve optimal coverage at the expense of larger interval widths.
-Using ensemble methods to combine forecasts from several flexible machine-learning models is one strategy that can be used to hedge against model misspecification and improve the quality of forecasts.
+Using ensemble methods to combine forecasts from several flexible machine learning models is one strategy that can be used to hedge against model misspecification and improve the quality of forecasts [@makridakis2020m4].
 
@@
 The algorithms presented in this work primarily consider symmetric intervals evaluated using the pinball loss function (AgACI can yield asymmetric intervals because the aggregation rule is applied separately to the lower and upper bounds from the underlying experts, but those underlying experts only produce symmetric intervals).
-A simple extension would switch to using the interval loss function, which would allow for asymmetric intervals where two parameters are learned for the upper and lower bounds, respectively.
-It may also be of interest to generate prediction intervals that have regret bounds and coverage guarantees for arbitrary subsets of observations (for example, we may seek prediction intervals that have near optimal coverage for every day of the week, or month of the year).
-Another avenue of theoretical research is to find algorithms with provable bounds for the coverage and regret that do not depend on the outcome being bounded.
-
--- Different AgACI weights for lower and upper bounds -- a step towards asymmetry, more flexible
-
-_Final wrap up paragraph?_
+A simple extension would switch to using the interval loss function [@gneiting2007scoring], which would allow for asymmetric intervals where two parameters are learned for the upper and lower bounds, respectively.
+It may also be of interest to generate prediction intervals that have coverage guarantees for arbitrary subsets of observations (for example, we may seek prediction intervals for daily observations that have near optimal coverage for every day of the week, or month of the year), similar to guarantees provided by the MultiValid Prediction method described in [@bastani2022practical].
+Another avenue for theoretical research is to propose algorithms with provable bounds for the coverage and regret that do not depend on the outcome being bounded.
+
+## Acknowledgements {.unnumbered}
+This research is partially supported by the Agence Nationale de la Recherche as part of the “Investissements d’avenir” program (reference ANR-19-P3IA-0001; PRAIRIE 3IA Institute).
+We would like to thank Margaux Zaffran for providing helpful comments on the manuscript.
 
@@
 
+## Additional simulation study results
 ```{r simulation_one_plot_appendix}
@@
 
-## Additional `R` code
-```{r}
-```
+```{r simulation_one_plot_joint_appendix}
+#| label: fig-simulation-one-joint
+#| fig-cap: "Mean Interval Width vs Coverage Error for the first simulation study. The error bars represent the 10% to 90% quantiles of the metrics over the simulation datasets."
+#| fig-height: 4
+
+simulation_one_joint_plot(simulation_study1$results)
+```
+
+```{r simulation_two_plot}
+#| label: fig-simulation-two-results
+#| fig-height: 9
+#| fig-cap: "Coverage error, mean interval width, and path length for $\\alpha = 0.8, 0.9, 0.95$ and simulations with and without distributional shift."
+
+simulation_two_plot(simulation_study2$results)
+```

### 2023-10-09 helpers.R

@@ -1,5 +1,9 @@
 label_alpha = label_bquote(cols = {alpha~"="~.(alpha)})
 
+extract_metric <- function(results, metric) {
+  unlist(lapply(results, function(result) result$metrics[[metric]]))
+}
+
 extract_intervals <- function(fit) {
   tibble(
     lower = fit$intervals[,1],
@@ -7,6 +11,10 @@ extract_intervals <- function(fit) {
     pred  = fit$predictions[,1]
   )
 }
+
+format_coverage <- scales::percent_format()
+format_path_length <- scales::number_format(accuracy = 0.1)
+format_mean_width <- scales::number_format(accuracy = 0.01)
 
 aci_theme <- theme(axis.title.y = element_text(size = 8))
 
@@ -27,8 +35,13 @@ simulation_one_plot <- function(results) {
     geom_boxplot() +
     geom_hline(yintercept = 0, lty = 2, alpha = 0.5) +
     facet_grid(alpha ~ param, labeller = labels) +
-    labs(x = "", y = expression(CovErr(T)), subtitle = "Coverage Error", title = "Simulation study: ARMA errors") +
-    basic_theme +
+    labs(
+      x = "", 
+      y = expression(CovErr(T)), 
+      subtitle = "Coverage Error", 
+      title = "Simulation study: ARMA errors"
+    ) +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
     aci_theme
   
   plot_piwidth <- results %>%
@@ -36,7 +49,7 @@ simulation_one_plot <- function(results) {
     geom_boxplot() +
     facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
     labs(x = "", y = expression(MeanWidth(T)), subtitle = "Mean Interval Width") +
-    basic_theme +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
     aci_theme
   
   plot_pathlength <- results %>%
@@ -44,11 +57,31 @@ simulation_one_plot <- function(results) {
     geom_boxplot() +
     facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
     labs(x = "", y = expression(PathLength(T)), subtitle = "Path Length") +
-    basic_theme +
     aci_theme +
-    theme(axis.text.x = element_text(angle = 90, hjust = 1))
[…]

### 2023-10-09 paper.qmd
Updated JJ affiliation

@@
     affiliations:
-      - name: Inria PreMeDICaL team, Université de Montpellier
+      - name: Inria PreMeDICaL team, Idesp, Université de Montpellier
         url: https://team.inria.fr/premedical/

### 2023-10-09 helpers.R
Added simulation study data

@@ -197,7 +197,7 @@ simulation_two_joint_plot <- function(results) {
 
 
 run_simulation_study1 <- function(data, setup, estimate_model, fit, workers = 8) {
-  cache <- "~/aci-paper/cache/simulation_study1.rds"
+  cache <- here::here("./simulation_studies/simulation_study1.rds")
   if(file.exists(cache)) return(read_rds(cache))
   
   plan(multisession, workers = workers)
@@ -238,7 +238,7 @@ run_simulation_study1 <- function(data, setup, estimate_model, fit, workers = 8)
 }
 
 run_simulation_study2 <- function(setup, fit, workers = 8) {
-  cache <- "~/aci-paper/cache/simulation_study2.rds"
+  cache <- here::here("./simulation_studies/simulation_study2.rds")
   if(file.exists(cache)) return(read_rds(cache))
   
   plan(multisession, workers = workers)

### 2023-10-25 paper.qmd
Updated paper

--- 2023-10-09 paper.qmd
@@
     affiliations:
-      - name: MAP5 (UMR 8145), Université Paris Cité
+      - name: Université Paris Cité, CNRS, MAP5, F-75006 Paris, France
         department: 
@@
 
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 

### 2023-11-17 paper.qmd
Updated references

@@
 bibliography: references.bib
-github-user: computorg
+github-user: herbps10
 repo: "template-computo-r"

### 2023-11-29 paper.qmd
Commit to trigger rebuild check

@@
 date-modified: last-modified
+logo: "true"
 description: |

### 2023-11-30 paper.qmd
Fixed references

@@
 For this case study, we focus on national (US) one-week ahead forecasts of weighted influenza-like illness (wILI), which is a population-weighted percentage of doctors visits where patients presented with influenza-like symptoms [@biggerstaff2016flusight].
-The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, from both mechanistic and statistical viewpoints [@tushar2018flusightnetwork, @tushar2019flusight].
+The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, from both mechanistic and statistical viewpoints [@flusight2020;@tushar2018flusightnetwork;@tushar2019flusight].
 For our purposes, we treat the way the forecasts were produced as a black box.

### 2023-11-30 paper.qmd
Updated references

@@
 ```{r case_study, cache = TRUE}
-raw_data <- read_csv("https://raw.githubusercontent.com/FluSightNetwork/cdc-flusight-ensemble/master/scores/point_ests.csv", show_col_types = FALSE)
+url <- "https://raw.githubusercontent.com/FluSightNetwork/cdc-flusight-ensemble/master/scores/point_ests.csv"
+raw_data <- read_csv(url, show_col_types = FALSE)
 

### 2024-06-13 paper.qmd
Revisions

--- 2023-11-29 paper.qmd
@@
     corresponding: true
-    email: herbert.susmann@dauphine.psl.eu
+    email: herbps10@gmail.com
     url: https://herbsusmann.com
@@
 date-modified: last-modified
-logo: "true"
 description: |
@@
 github-user: herbps10
+logo: "true"
 repo: "template-computo-r"
@@
   computo-pdf: 
+    keep-tex: true
     include-in-header:
@@
 The critical dependence of the original ACI algorithm on proper choice of its learning rate spurred subsequent research into meta-algorithms that learn the correct learning rate (or an analogue thereof) in various ways, typically drawing on approaches from the online learning literature.
-In this paper, we present four such algorithms: Aggregated ACI [AgACI, @zaffran2022agaci], Fully Adaptive Conformal Inference [FACI, @gibbs2022faci], Scale-Free Online Gradient Descent [SF-OGD, @bhatnagar2023saocp], and Strongly Adaptive Online Conformal Prediction [SAOCP, @bhatnagar2023saocp].
-We note that the adaption of conformal inference techniques is an active area of research and the algorithms we focus on in this work are not exhaustive; see among others @feldman2023achieving, @bastani2022practical, @xu2021enbpi, and @xu2023spci.
+In this paper, we present four such algorithms: Aggregated ACI [AgACI, @zaffran2022agaci], Dynamically-tuned Adaptive ACI [DtACI, @gibbs2022faci], Scale-Free Online Gradient Descent [SF-OGD, @bhatnagar2023saocp], and Strongly Adaptive Online Conformal Prediction [SAOCP, @bhatnagar2023saocp].
+We note that the adaption of conformal inference techniques is an active area of research and the algorithms we focus on in this work are not exhaustive; see among others @feldman2023achieving, @bastani2022practical, @xu2021enbpi, @xu2023spci, @angelopoulos2024online, @zhang2024discounted, and @gasparin2024conformal.
 
@@
 
-Define $r_t := \inf\{\theta \in \mathbb{R} : \mathbb{I}(y_t \in \widehat{C}_t(\theta)) \}$ to be the _radius_ at time $t$.
+Define $r_t := \inf\{\theta \in \mathbb{R} : y_t \in \widehat{C}_t(\theta) \}$ to be the _radius_ at time $t$.
 The radius is the smallest possible $\theta$ such that the prediction interval covers the observation $y_t$.
@@
 
-**Assumption**: there exists a $D > 0$ such that $r_t < D$ for all $t$.
-
-Next, we describe two existing definitions of interval construction functions.
+**Assumption**: there exists a finite $D > 0$ such that $r_t < D$ for all $t$.
+
+If the outcome space is bounded, then $D$ can be easily chosen to cover the entire space. Next, we describe two existing definitions of interval construction functions.
 
@@
 Note that $\widehat{C}_t$ is indeed nested in $\theta_t$ because the Quantile function is non-decreasing in $\theta$.
+Note we define $\widehat{C}_t(1) = \max\{s_1, \dots, s_{t-1}\}$ rather than $\widehat{C}_t(1) = \infty$ in order to avoid practical problems with trivial prediction intervals [@zaffran2022agaci].
+Note that we can always choose $D = 1$ to satisfy the outcome boundedness assumption given above.
+
+We focus on the above definition for the quantile interval construction function which is designed to be symmetric around the point prediction $\hat{\mu}_t$.
+However, we note it is possible to take a more general definition, such as
+$$
+\widehat{C}_t(\theta_t) := \{ y : S(\hat{\mu}_t, y) \leq \mathrm{Quantile(\theta, \{ s_1, \dots, s_{t-1}\})} \}
+$$
+Such an approach allows for prediction intervals that may not be centered on $\hat{\mu}_t$.
 
@@
 (\theta, r) \mapsto L^\alpha(\theta, r) := \begin{cases}
-  \alpha(\theta - r), & \theta \geq r \\
-  (1-\alpha)(r - \theta), & \theta < r.
+  (1 - \alpha)(\theta - r), & \theta \geq r \\
+  \alpha(r - \theta), & \theta < r.
 \end{cases}
@@
 As such, stronger notions of regret have been defined.
-The _strongly adaptive regret_ is the largest regret over any subperiod of length $k \in \llbracket T \rrbracket$:
+The _strongly adaptive regret_ is the largest regret over any subperiod of length $m \in \llbracket T \rrbracket$:
 $$
@@
 
-| Algorithm                                             | Tuning Parameters                                           | Original interval constructor | Citation            |
-| ------------------------------------------------------|-------------------------------------------------------------|---------------------|
-| Adaptive Conformal Inference (ACI)                    | Learning rate $\gamma$                                      | Quantile | @gibbs2021adaptive  |
-| Aggregated Adaptive Conformal Inference (AgACI)       | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | Quantile | @zaffran2022agaci   |
-| Fully Adaptive Conformal Inference (FACI)             | Candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ | Quantile | @gibbs2022faci      |
-| Scale-Free Online Gradient Descent (SF-OGD)           | Learning rate $\gamma$ or maximum radius $D$             | Linear | @bhatnagar2023saocp |
-| Strongly Adaptive Online Conformal Prediction (SAOCP) | Learning rate $\gamma$, lifetime multiplier $g$          | Linear | @bhatnagar2023saocp |
-
-  : Summary of ACI algorithms {#tbl-aci}
+| Algorithm |
+| ------------------------------------------------------|
+| **Adaptive Conformal Inference (ACI)** |
+| - Tuning parameters: learning rate $\gamma$ |
+| - Original interval constructor: quantile |
+| - Theoretical guarantees: coverage error, regret |
+| - Citation: @gibbs2021adaptive |
+| **Aggregated Adaptive Conformal Inference (AgACI)** |
+| - Tuning parameters: candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ |
+| - Original interval constructor: quantile |
+| - Citation: @zaffran2022agaci  |
+| **Dynamically-tuned Adaptive Conformal Inference (DtACI)** |
+| - Tuning parameters: candidate learning rates $(\gamma_k)_{1 \leq k \leq K}$ |
+| - Original interval constructor: quantile |
+| - Theoretical guarantees: coverage error, strongly adaptive regret, dynamic regret |
+| - Citation | @gibbs2022faci |
+| **Scale-Free Online Gradient Descent (SF-OGD)** |
+| - Tuning parameters: learning rate $\gamma$ or maximum radius $D$ |
+| - Original interval constructor: linear  |
+| - Theoretical guarantees: coverage error, anytime regret  |
+| - Citation: @bhatnagar2023saocp  |
+| **Strongly Adaptive Online Conformal Prediction (SAOCP)** |
+| - Tuning parameters: learning rate $\gamma$, lifetime multiplier $g$  |
+| - Original interval constructor: linear  |
+| - Theoretical guarantees: coverage error, strongly adaptive regret  |
+| - Citation: @bhatnagar2023saocp  |
+: Summary of ACI algorithms. Only the theoretical guarantees discussed in this paper are shown for each algorithm.  {#tbl-aci}
 
@@
 $$
-y_t \sim N(0, 0.2^2), \quad t \in \llbracket T \rrbracket.
+\begin{aligned}
+y_t &\sim N(0, \sigma_t^2), \quad t \in \llbracket T \rrbracket, \\
+\sigma_t &= \begin{cases}
+  0.2, &t \leq 250, \\
+  0.1, &t > 250.
+\end{cases}
+\end{aligned}
 $$
@@
 N <- 5e2
-mu <- rep(0, N)
-muhat <- mu
-y <- rnorm(N, mean = mu, sd = 0.2)
+
+running_example <- running_example_data(N)
+y <- running_example$y
+yhat <- running_example$yhat
 ```
@@
 ### Theoretical Guarantees
-The ACI algorithm has the following finite sample bound on the coverage error  [@gibbs2021adaptive]. For all $\gamma > 0$, 
+With the choice of the quantile interval structure, the ACI algorithm has the following finite sample bound on the coverage error  (@gibbs2021adaptive; Proposition 4.1).
+For all $\gamma > 0$ (and so long as $\gamma$ does not depend on $T$),
+$$
+|\mathrm{CovErr}(T)| \leq \frac{\max\{\theta_1, 1 - \theta_1\} + \gamma}{\gamma T}.
+$$
+This result was originally shown for ACI with the choice of the quantile interval constructor, although it can also be extended to other interval constructors [@bhatnagar2023saocp, @feldman2023achieving].
+More generally, the algorithm has the following coverage error bound in terms of the radius bound $D$ [@bhatnagar2023saocp]:
 $$
@@
 $$
-This result was originally shown for ACI with the choice of the quantile interval constructor, although it can also be extended to other interval constructors [@bhatnagar2023saocp, @feldman2023achieving].
 In addition, standard results for online subgradient descent yield the following regret bound with the use of the linear interval constructor, assuming that the true radii are bounded by $D$:
@@
 $$
-where the second inequality follows if the optimal choice of $\gamma = D/\sqrt{T}$ is used [@bhatnagar2023saocp].
+where the second inequality follows if the optimal choice (with respect to long-term regret) of $\gamma = D/\sqrt{T}$ is used [@bhatnagar2023saocp].
 Taken together, these theoretical results imply that while the coverage error is guaranteed to converge to zero for any choice of $\gamma$, achieving sublinear regret requires choosing $\gamma$ more carefully.
@@
 Thus, choosing a good value for $\gamma$ is essential.
-However, the optimal choice $\gamma = D / \sqrt{T}$ cannot be used directly in practice unless the time series length $T$ is fixed in advance, or the so called "doubling trick" is used to relax the need to know $T$ in advance [@cesabianchi2006games].
+However, the optimal choice $\gamma = D / \sqrt{T}$ cannot be used directly in practice unless the time series length $T$ is fixed in advance, or the so called "doubling trick" is used to relax the need to know $T$ in advance (@cesabianchi2006games; Section 2.3).
 
@@
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[1])),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[2])),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[3])),
-  aci(y, muhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[4]))
+  aci(y, yhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[1])),
+  aci(y, yhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[2])),
+  aci(y, yhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[3])),
+  aci(y, yhat, alpha = alpha, method = "ACI", parameters = list(interval_constructor = "conformal", theta0 = alpha, gamma = gamma_grid[4]))
 )
@@
 
-### Tuning Parameters
+### Tuning Parameters {#sec-agaci-tuning}
 The main tuning parameter for AgACI is the set of candidate learning rates.
@@
 @fig-agaci illustrates AgACI applied to the running example with two sets of learning grids.
-The first grid is the default, and the second grid includes the additional value $\gamma = \{ 0.0005 \}$.
-For the first grid, we can see that for the lower bound AgACI assigns high weight to the lowest learning rate ($\gamma = 0.001$).
-Based on this observation, as a sensitivity check, we reran the algorithm with the second learning grid, which yields weights that are less concentrated on a single learning rate.
-The output prediction intervals are similar, although slightly smoother in the second grid, reflecting the lack of distributional shift in the simulated data.
+The first grid is $\gamma = \{ 0.032, 0.064, 0.128, 0.256 \}$, and the second grid includes the additional values $\{ 0.008, 0.016 \}$.
+For the first grid, we can see that for the lower bound AgACI assigns high weight to the lowest learning rate ($\gamma = 0.032$).
+The second grid yields weights that are less concentrated on a single learning rate, and the output prediction intervals are smoother.
 
@@
 #| fig-height: 4.5
-#| fig-cap: "Example 80% prediction intervals from the AgACI algorithm with starting values $\\theta_1 = 0.8$ and two different learning rate grids, where the second grid is double the size of the first.
+#| fig-cap: "Example 80% prediction intervals from the AgACI algorithm with starting values $\\theta_1 = 0.8$ and two different learning rate grids.
 In the left column, blue and red points are observations that fell inside and outside the prediction intervals, respectively."
-grid1 <- 2^(seq(0, 7, 1)) / 1e3
-grid2 <- 2^(seq(-1, 5, 1)) / 1e3
+grid1 <- 2^(seq(5, 8, 1)) / 1e3
+grid2 <- 2^(seq(1, 8, 1)) / 1e3
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid1, theta0 = 0.8)),
-  aci(y, muhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid2, theta0 = 0.8))
+  aci(y, yhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid1, theta0 = 0.8)),
+  aci(y, yhat, alpha = alpha, method = "AgACI", parameters = list(interval_constructor = "conformal", gamma_grid = grid2, theta0 = 0.8))
 )
@@
 
-  plot_agaci_weights(results[[i]], main = "Final Aggregation Weights", legend = "topright")
+  plot_agaci_weights(results[[i]], main = "Final Aggregation Weights", legend = "topright", ylim = c(0, 1))
 }
@@
 
-## Fully Adaptive Conformal Inference (FACI)
+## Dynamically-tuned Adaptive Conformal Inference (DtACI)
 ```pseudocode
@@
 \begin{algorithm}
-\caption{Fully Adaptive Conformal Inference}
+\caption{Dynamically-tuned Adaptive Conformal Inference}
 \begin{algorithmic}
@@
 \For{$t = 1, 2, \dots, T$}
-  \State Define $p_t^k := w_t^k / \sum_{i=1}^K w_t^i$, for all $1 \leq k \leq K$.
+  \State Define $p_t^k := p_t^k / \sum_{i=1}^K p_t^i$, for all $1 \leq k \leq K$.
   \State Set $\theta_t = \sum_{k=1}^K \theta_t^k p_t^k$.
@@
   \State Observe $y_t$ and compute $r_t$.
-  \State $\bar{w}_{t}^k \gets w_t^k \exp(-\eta L^\alpha(\theta_t^k, r_t))$, for all $1 \leq k \leq K$.
+  \State $\bar{w}_{t}^k \gets p_t^k \exp(-\eta L^\alpha(\theta_t^k, r_t))$, for all $1 \leq k \leq K$.
   \State $\bar{W}_t \gets \sum_{i=1}^K \bar{w}_t^i$.
-  \State $w_{t+1}^k \gets (1 - \sigma) \bar{w}_t^k + \bar{W}_t \sigma / K$.
+  \State $p_{t+1}^k \gets (1 - \sigma) \bar{w}_t^k + \bar{W}_t \sigma / K$.
   \State Set $\mathrm{err}_t := \mathbb{I}[y_t \not\in \widehat{C}_t(\theta_t)]$.
@@
 
-The Fully Adaptive Conformal Inference (FACI; @algo-faci) algorithm was developed by the authors of the original ACI algorithm in part to address the issue of how to choose the learning rate parameter $\gamma$.
+The Dynamically-tuned Adaptive Conformal Inference (DtACI; @algo-faci) algorithm was developed by the authors of the original ACI algorithm in part to address the issue of how to choose the learning rate parameter $\gamma$.
 In this respect the goal of the algorithm is similar to that of AgACI, although it is achieved slightly differently.
-FACI also aggregates predictions from multiple copies of ACI run with different learning rates, but differs in that it directly aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci] using an exponential reweighting scheme [@gradu2022adaptive].
+DtACI also aggregates predictions from multiple copies of ACI run with different learning rates, but differs in that it directly aggregates the estimated radii emitted from each algorithm based on their pinball loss [@gibbs2022faci] using an exponential reweighting scheme [@gradu2022adaptive].
 As opposed to AgACI, this construction allows for more straightforward development of theoretical guarantees on the algorithm's performance, because the upper and lower bounds of the intervals are not aggregated separately.
@@
 ### Theoretical Guarantees
-FACI was originally proposed with the choice of the quantile interval constructor.
-FACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]: for all $\gamma > 0$ and subperiod lengths $m$,
-$$
-\begin{aligned}
-  \mathrm{SAReg}(T, m) \leq \widetilde{\mathcal{O}}(D^2 / \gamma + \gamma m).
-\end{aligned}
-$$
-If $m$ is fixed a-priori, then choosing $\gamma = D/\sqrt{m}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{m})$ (for a single choice of $m$).
+DtACI was originally proposed with the choice of the quantile interval constructor.
+DtACI has the following strongly-adaptive regret bound [@bhatnagar2023saocp]: for all $\eta > 0$ and subperiod lengths $m$,
+$$
+\begin{aligned}
+  \mathrm{SAReg}(T, m) \leq \widetilde{\mathcal{O}}(D^2 / \eta + \eta m).
+\end{aligned}
+$$
+If $m$ is fixed a-priori, then choosing $\eta = D/\sqrt{m}$ yields a strongly adaptive regret bound of order $\widetilde{\mathcal{O}}(D \sqrt{m})$ (for a single choice of $m$).
 Practically, this result implies that, if we know in advance the time length for which we would like to control the regret, it is possible to choose an optimal tuning parameter value.
@@
 
-To establish a bound on the coverage error, the authors investigated a slightly modified version of FACI in which $\theta_t$ is chosen randomly from the candidate $\theta_{t_k}$ with weights given by $p_{t,k}$, instead of taking a weighted average.
+To establish a bound on the coverage error, the authors investigated a slightly modified version of DtACI in which $\theta_t$ is chosen randomly from the candidate $\theta_{t_k}$ with weights given by $p_{t,k}$, instead of taking a weighted average.
 This is a common trick used in the literature as it facilitates theoretical analysis.
-In practice, the authors comment that this randomized version of FACI and the deterministic version lead to very similar results.
+In practice, the authors comment that this randomized version of DtACI and the deterministic version lead to very similar results.
 The coverage error result also assumes that the hyperparameters can change over time: that is, we have $t$-specific $\eta_{t}$ and $\sigma_t$, rather than fixed $\eta$ and $\sigma$.
-The coverage error then has the following bound [@gibbs2022faci], where $\gamma_{\min}$ and $\gamma_{\max}$ are the smallest and largest learning rates in the grid, respectively:
-$$
-|\mathrm{CovErr}(T)| \leq \frac{1 + 2\gamma_{\max}}{T \gamma_{\min}} + \frac{(1 + 2\gamma_{\max})^2}{\gamma_{\min}} \exp(\eta_t(1 + 2\gamma_{\max})) \frac{1}{T}\sum_{t=1}^T \eta_t + 2 \frac{1+\gamma_{\max}}{\gamma_{\min}} \frac{1}{T} \sum_{t=1}^T \sigma_t.
+The coverage error then has the following bound (@gibbs2022faci; Theorem 3.2), where $\gamma_{\min}$ and $\gamma_{\max}$ are the smallest and largest learning rates in the grid, respectively:
+$$
+|\mathrm{CovErr}(T)| \leq \frac{1 + 2\gamma_{\max}}{T \gamma_{\min}} + \frac{(1 + 2\gamma_{\max})^2}{\gamma_{\min}}  \frac{1}{T}\sum_{t=1}^T \eta_t \exp(\eta_t(1 + 2\gamma_{\max})) + 2 \frac{1+\gamma_{\max}}{\gamma_{\min}} \frac{1}{T} \sum_{t=1}^T \sigma_t.
 $$
@@
 
-We note one additional result established by @gibbs2022faci on a slightly different regret bound in terms of the pinball loss, as it informs the choice of tuning parameters.
+We note one additional result established by @gibbs2022faci (their Theorem 3.1) on a slightly different dynamic regret bound in terms of the pinball loss, as it informs the choice of tuning parameters.
 Let $\gamma_{\mathrm{max}} = \max_{1 \leq k \leq K} \gamma_k$ be the largest learning rate in the grid and assume that $\gamma_1 < \gamma_2 < \cdots < \gamma_K$ with $\gamma_{k+1}/\gamma \leq 2$ for all $1 \leq k < K$.
@@
 The parameter $|I|$, the time interval of interest for which we would like to control, can be chosen arbitrarily.
+This dynamic regret bound can be converted to a strongly adaptive regret bound by choosing $\theta^*_t$ to be constant.
 
@@
 The recommended settings for the tuning parameters depend on choosing a time interval length $|I|$ for which we would like to control the pinball loss. The choice of $|I|$ can be chosen arbitrarily.
-For the tuning parameter $\sigma$, the authors suggest the optimal choice $\sigma = 1 / (2 |I|)$.
+For the tuning parameter $\sigma$, the authors suggest the optimal choice (with respect to the dynamic regret) $\sigma = 1 / (2 |I|)$.
 Choosing $\eta$ is more difficult.
@@
 
-@fig-faci-example illustrates FACI with the quantile interval construction function and with the learning rate grid $\gamma \in \{ 0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128 \}$.
+@fig-dtaci-example illustrates DtACI with the quantile interval construction function and with the learning rate grid $\gamma \in \{ 0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128 \}$.
+The tuning parameter $\eta$ was set to $0.001$, $1$, and $100$ to show how the algorithm responds to extreme choices of the parameter, and to $\eta \approx 3.19$ according to the optimal choice recommendation with $I = 100$ as described in the previous section.
+The results show that, in this simple example, high values of $\eta$ may lead to intervals that are too reactive to the data, as seen in the longer path length.
+The algorithm appears more robust, however, to small choices of $\eta$.
 
 ```{r faci_example_plot, echo = FALSE}
-#| label: fig-faci-example
+#| label: fig-dtaci-example
 #| fig-height: 4.5
-#| fig-cap: "Example 80% prediction intervals generated by the FACI algorithm with starting values $\\theta_1 = 0.8$.
+#| fig-cap: "Example 80% prediction intervals generated by the DtACI algorithm with starting values $\\theta_1 = 0.8$ and with several values of the tuning parameter $\\eta$.
 Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
 gamma_grid <- c(0.001, 0.002, 0.004, 0.008, 0.016, 0.032, 0.064, 0.128)
-result <- aci(y, muhat, alpha = alpha, method = "FACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal"))
-
-coverage    <- format_coverage(extract_metric(list(result), "coverage"))
-path_length <- format_path_length(extract_metric(list(result), "path_length"))
-
-plot(result, ylim = c(-0.9, 0.8), legend = FALSE)
-text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[1]]) ), pos = 4)
-text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[1]])), pos = 4)
+results <- list(
+  aci(y, yhat, alpha = alpha, method = "DtACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal", eta = 0.001)),
+  aci(y, yhat, alpha = alpha, method = "DtACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal", eta = 1)),
+  aci(y, yhat, alpha = alpha, method = "DtACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal", eta = 100)),
+  aci(y, yhat, alpha = alpha, method = "DtACI", parameters = list(gamma_grid = gamma_grid, interval_constructor = "conformal"))
+)
+
+coverage    <- format_coverage(extract_metric(results, "coverage"))
+path_length <- format_path_length(extract_metric(results, "path_length"))
+
+par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
+for(i in 1:4) {
+  plot(results[[i]], legend = FALSE, predictions = FALSE, cex = 0.5, main = bquote(eta==.(results[[i]]$parameters$eta)), ylim = c(-0.9, 0.8))
+  text(x = -10, y = -0.75, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
+  text(x = -10, y = -0.9, labels = bquote(PathLength == .(path_length[[i]])), pos = 4)
+}
+par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
 ```
@@
 $$
-A bound for the coverage error has also been established [@bhatnagar2023saocp].
+A bound for the coverage error has also been established (@bhatnagar2023saocp; Theorem 4.2).
 For any learning rate $\gamma = \Theta(D)$ (where $\gamma = D / \sqrt{3}$ is optimal) and any starting value $\theta_1 \in [0, D]$, then it holds that for any $T > 1$,
@@
 @fig-sf-ogd illustrates SF-OGD for several values of $\gamma$.
+In the example, the prediction intervals are not reactive enough and do not achieve optimal coverage when $\gamma$ is small.
+As $\gamma$ increases, the coverage error is near optimal, although the path length becomes larger.
 
@@
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.01)),
-  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.1)),
-  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.25)),
-  aci(y, muhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.5))
+  aci(y, yhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.01)),
+  aci(y, yhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.1)),
+  aci(y, yhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.25)),
+  aci(y, yhat, alpha = alpha, method = "SF-OGD", parameters = list(gamma = 0.5))
 )
@@
 \For{$t = 1, 2, \dots, T$}
-  \State Initialize expert $\mathcal{A}_t = \texttt{SF-OGD}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma, \theta_1 \leftarrow \theta_{t-1})$, set weight $w_t^t = 0$.
+  \State Initialize expert $\mathcal{A}_t = \texttt{SF-OGD}(\alpha \leftarrow \alpha, \gamma \leftarrow \gamma, \theta_1 \leftarrow \theta_{t-1})$, set weight $p_t^t = 0$.
   \State Compute active set $\mathrm{Active}(t) = \{ i \in \llbracket T \rrbracket : t - L(i) < i \leq t \}$ (see below for definition of $L(t)$).
   \State Compute prior probability $\pi_i \propto i^{-2} (1 + \lfloor \log_2 i \rfloor )^{-1} \mathbb{I}[i \in \mathrm{Active}(t)]$.
-  \State Compute un-normalized probability $\hat{p}_i = \pi_i [w_{t,i}]_+$ for all $i \in \llbracket t \rrbracket$.
+  \State Compute un-normalized probability $\hat{p}_i = \pi_i [p_{t,i}]_+$ for all $i \in \llbracket t \rrbracket$.
   \State Normalize $p = \hat{p} / \| \hat{p} \|_1 \in \Delta^t$ if $\| \hat{p} \|_1 > 0$, else $p = \pi$.
@@
     \State Compute $g_t^i = \begin{cases}
-      \frac{1}{D}\left(L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t)\right) & w_t^i > 0 \\
-      \frac{1}{D}\left[L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t))\right]_+ & w_t^i \leq 0 \\
+      \frac{1}{D}\left(L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t)\right) & p_t^i > 0 \\
+      \frac{1}{D}\left[L^\alpha(\theta_t, r_t) - L^\alpha(\theta_t^i, r_t))\right]_+ & p_t^i \leq 0 \\
     \end{cases}$.
-    \State Update expert weight $w_{t+1}^i = \frac{1}{t - i + 1}\left( \sum_{j=i}^t g_j^i \right) \left(1 + \sum_{j=i}^t w_j^i g_j^i \right)$.
+    \State Update expert weight $p_{t+1}^i = \frac{1}{t - i + 1}\left( \sum_{j=i}^t g_j^i \right) \left(1 + \sum_{j=i}^t p_j^i g_j^i \right)$.
   \EndFor
@@
 The Strongly Adaptive Online Conformal Prediction (SAOCP; @algo-saocp) algorithm was proposed as an improvement over the extant ACI algorithms in that it features stronger theoretical guarantees.
-SAOCP works similarly to AgACI and FACI in that it maintains a library of candidate online learning algorithms that generate prediction intervals which are then aggregated using a meta-algorithm [@bhatnagar2023saocp].
+SAOCP works similarly to AgACI and DtACI in that it maintains a library of candidate online learning algorithms that generate prediction intervals which are then aggregated using a meta-algorithm [@bhatnagar2023saocp].
 The candidate algorithm was chosen to be SF-OGD, although any algorithm that features anytime regret guarantees can be chosen.
-As opposed to AgACI and FACI, in which each candidate has a different learning rate but is always able to contribute to the final prediction intervals, here each candidate has the same learning rate but only has positive weight over a specific interval of time.
+As opposed to AgACI and DtACI, in which each candidate has a different learning rate but is always able to contribute to the final prediction intervals, here each candidate has the same learning rate but only has positive weight over a specific interval of time.
 New candidate algorithms are continually being spawned in order that, if the distribution shifts rapidly, the newer candidates will be able to react quickly and receive positive weight.
@@
 The authors show that this construction results in intervals that have strong regret guarantees.
+The form of the lifetime interval function $L(t)$ is due to the use of geometric covering intervals to partition the input time series, and other choices may be possible [@jun2017coinbetting].
 
@@
 The theoretical results were established for SAOCP using the linear interval constructor.
-The following bound for the strongly adaptive regret holds for all subperiod lengths $m \in \llbracket T \rrbracket$ [@bhatnagar2023saocp]:
+The following bound for the strongly adaptive regret holds for all subperiod lengths $m \in \llbracket T \rrbracket$ (@bhatnagar2023saocp; Proposition 4.1):
 $$
@@
 $$
-It should be emphasized that this regret bounds holds simultaneously across all $m$, as opposed to FACI, where a similar bound holds only for a single $m$.
+It should be emphasized that this regret bounds holds simultaneously across all $m$, as opposed to DtACI, where a similar bound holds only for a single $m$.
 A bound on the coverage error of SAOCP has also been established as:
@@
 $$
-where $S_{\beta}(T)$ is a technical measure of the smoothness of the cumulative gradients and expert weights for each of the candidate experts [@bhatnagar2023saocp].
+where $S_{\beta}(T)$ is a technical measure of the smoothness of the cumulative gradients and expert weights for each of the candidate experts (@bhatnagar2023saocp; Theorem 4.3).
+For some intuition, $S_{\beta}$ can be expected to be small when the weights placed on each algorithm do change quickly, as would be the case under abrupt distributional shifts.
 
 ### Tuning Parameters
-The main tuning parameter for SAOCP is the learning rate $\gamma$ of the SF-OGD sub-algorithms, which we saw in the previous section has for optimal choice $\gamma = D / \sqrt{3}$.
+The primary tuning parameter for SAOCP is the learning rate $\gamma$ of the SF-OGD sub-algorithms, which we saw in the previous section has for optimal choice $\gamma = D / \sqrt{3}$.
 Values for $D$ that are too low lead to intervals that adapt slowly, and values that are too large lead to jagged intervals.
@@
 @fig-saocp illustrates the SAOCP algorithm for choices of $D \in \{0.01, 0.1, 0.25, 0.5 \}$.
-
-```{r saocp_example_plot, echo = FALSE}
+Similarly to SF-OGD, the prediction intervals tend to undercover for small $D$, and achieve near-optimal coverage for larger $D$ at the expense of larger path lengths.
+
+```{r saocp_example_plot, echo = FALSE, cache = TRUE}
 #| label: fig-saocp
@@
 results <- list(
-  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.01)),
-  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.1)),
-  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.25)),
-  aci(y, muhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.5))
+  aci(y, yhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.01)),
+  aci(y, yhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.1)),
+  aci(y, yhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.25)),
+  aci(y, yhat, alpha = alpha, method = "SAOCP", parameters = list(D = 0.5))
 )
@@
 
+# `AdaptiveConformal` `R` package
+The ACI algorithms described in the previous section have been implemented in the open-source and publically available `R` package `AdaptiveConformal`, available at [https://github.com/herbps10/AdaptiveConformal](https://github.com/herbps10/AdaptiveConformal).
+CIn this section, we briefly introduce the main functionality of the package.
+Comprehensive documentation is, including several example vignettes, is included with the package.
+
+The `AdaptiveConformal` package can be installed using the `remotes` package:
+```{r, eval = FALSE}
+remotes::install_github("herbps10/AdaptiveConformal")
+```
+
+The ACI algorithms are accessed through the `aci` function, which takes as input a vector of observations ($y_t$) and a vector or matrix of predictions ($\hat{y}_t$).
+Using the data generating process from the running example to illustrate, we can fit the original ACI algorithm with learning rate $\gamma = 0.1$:
+```{r}
+set.seed(532)
+data <- running_example_data(N = 5e2)
+fit <- aci(data$y, data$yhat, alpha = 0.8, method = "ACI", parameters = list(gamma = 0.1))
+```
+The available parameters for each method can be found in the documentation for the `aci` method, accessible with the command `?aci`.
+The resulting conformal prediction intervals can then be plotted using the `plot` function:
+```{r, fig.height = 4}
+plot(fit)
+```
+The properties of the prediction intervals can also be examined using the `summary` function:
+```{r}
+summary(fit)
+```
+
 # Simulation Studies {#sec-simulations}
-We present two empirical studies in order to compare the performance of the AgACI, FACI, SF-OGD, and SAOCP algorithms applied to simple simulated datasets.
+We present two empirical studies in order to compare the performance of the AgACI, DtACI, SF-OGD, and SAOCP algorithms applied to simple simulated datasets.
 The original ACI algorithm was not included as it is not clear how to set the tuning rate $\gamma$, which can have a large effect on the resulting intervals.
@@
     method == "AgACI" ~ "conformal",
-    method == "FACI" ~ "conformal",
+    method == "DtACI" ~ "conformal",
     method == "SF-OGD" ~ "linear",
@@
 
-N_sims <- 25
+N_sims <- 100
 simulation_data <- expand_grid(
@@
   alpha = c(0.8, 0.9, 0.95),
-  method = c("AgACI", "SF-OGD", "SAOCP", "FACI")
-)
-
+  method = c("AgACI", "SF-OGD", "SAOCP", "DtACI")
+)
+
+# run_simulation_study1 function is defined in helpers.R
 simulation_study1 <- run_simulation_study1(
@@
 
-The coverage errors, mean interval widths, and path lengths of each of the algorithms for $\alpha = 0.9$ are shown in @fig-simulation-one-results (results for $\alpha \in \{ 0.8, 0.95 \}$ were similar and are available in the appendix).
+The coverage errors, mean interval widths, path lengths, and strongly adaptive regret (for $m = 20$) of each of the algorithms for $\alpha = 0.9$ are shown in @fig-simulation-one-results (results for $\alpha \in \{ 0.8, 0.95 \}$ were similar and are available in the appendix).
 All methods achieved near optimal empirical coverage, although SAOCP tended to slightly undercover.
-The mean interval widths were similar across methods, although again SAOCP had slightly shorter intervals (as could be expected given its tendency to undercover).
+The mean interval widths re similar across methods, although again SAOCP had slightly shorter intervals (as could be expected given its tendency to undercover).
+The strongly adaptive regret was similar for all methods.
 The path length of SAOCP was larger than any of the other methods.
-To investigate why, @fig-simulation-one-example plots $w_t - w_{t-1}$, the difference in interval width between times $t-1$ and $t$, for each method in one of the simulations.
-The interval widths for AgACI and FACI change slowly relative to those for SF-OGD and SAOCP.
+To investigate why, @fig-simulation-one-widths plots $w_t - w_{t-1}$, the difference in interval width between times $t-1$ and $t$, for each method in one of the simulations.
+The interval widths for AgACI and DtACI change slowly relative to those for SF-OGD and SAOCP.
 For SAOCP, we can see the interval widths have larger fluctuations than for the other methods, explaining its higher path width.
+The prediction intervals themselves for the same simulation are shown in @fig-simulation-one-example, which shows that although the path lengths are quite different, the output prediction intervals are quite similar.
 
@@
 #| label: fig-simulation-one-results
-#| fig-height: 6.5
-#| fig-cap: "Coverage errors, mean interval widths, and path lengths for the first simulation study with target coverage $\\alpha = 0.9$."
+#| fig-height: 9
+#| fig-cap: "Coverage errors, mean interval widths, path lengths, and strongly adaptive regret (for $m = 20$) for the first simulation study with target coverage $\\alpha = 0.9$."
 
@@
 
-```{r simulation_one_example_plot}
-#| label: fig-simulation-one-example
+```{r simulation_one_width_plot}
+#| label: fig-simulation-one-widths
 #| fig-height: 5
@@
 
+```{r simulation_one_example_plots}
+#| label: fig-simulation-one-example
+#| fig-height: 6
+#| fig-cap: "Example prediction intervals (target coverage $\\alpha = 0.9$) from the first simulation study for time points 300 to 350; metrics shown are for all time points $t \\geq 300$.
+Blue and red points are observations that fell inside and outside the prediction intervals, respectively."
+fits <- simulation_study1$example_fits
+
+coverage    <- format_coverage(map_dbl(map(fits$fit, metrics), `[[`, "coverage"))
+path_length <- format_path_length(map_dbl(map(fits$fit, metrics), `[[`, "path_length"))
+
+par(mfrow = c(2, 2), mar = c(3, 3, 2, 1))
+for(i in 1:4) {
+  plot(fits$fit[[i]], legend = FALSE, main = fits$method[[i]], predictions = FALSE, ylim = c(-20, 35), index = 50:100)
+  text(x = -0, y = -7.5, labels = bquote(EmpCov == .(coverage[[i]]) ), pos = 4)
+  text(x = -0, y = -17.5, labels = bquote(PathLength == .(path_length[[i]]) ), pos = 4)
+}
+par(mfrow = c(1, 1), mar = c(5.1, 4.1, 4.1, 2.1))
+```
+
 ## Distribution shift
@@
   shift <- 1:N > (N / 2)
-  muhat <- mu
+  yhat <- mu
   y <- rnorm(n = length(mu), mean = mu, sd = sigma + ifelse(shift, distribution_shift, 0))
 
-  tibble(y = y, muhat = muhat)
+  tibble(y = y, yhat = yhat)
 }
@@
   if(interval_constructor == "linear") {
-    D <- max(abs(data$y - data$muhat)[1:50])
+    D <- max(abs(data$y - data$yhat)[1:50])
   }
@@
 
-  aci(data$y, data$muhat, method = method, alpha = alpha, parameters = parameters)
-}
-
-N_sims <- 5e1
+  aci(data$y, data$yhat, method = method, alpha = alpha, parameters = parameters)
+}
+
+N_sims <- 100
 simulation_study_setup2 <- expand_grid(
@@
   N = 500,
-  method = c("AgACI", "SF-OGD", "SAOCP", "FACI"),
+  method = c("AgACI", "SF-OGD", "SAOCP", "DtACI"),
 ) %>%
@@
 
+# run_simulation_study2 function is defined in helpers.R
 simulation_study2 <- run_simulation_study2(simulation_study_setup2, fit, workers = 8)
@@
 
-The coverage error, mean path length, and mean interval widths of the algorithms are summarized in @fig-simulation-two-joint (an alternative plot is included in the appendix as @fig-simulation-two-results).
+The coverage error, mean path length, mean interval widths, and strongly adaptive regret (for $m = 20$) of the algorithms are summarized in @fig-simulation-two-results (an alternative plot is included in the appendix as @fig-simulation-two-joint).
 The coverage error of all the algorithms is near the desired value in the absence of distribution shift.
-On the contrary, all of the algorithms except AgACI and FACI undercover when there is distributional shift.
+On the contrary, all of the algorithms except AgACI and DtACI undercover when there is distributional shift.
 SAOCP tends to have higher average path lengths than the other methods.
+In the distribution shift setting, SF-OGD and SAOCP tended to have smaller strongly adaptive regret than the other methods.
 An illustrative example of prediction intervals generated by each method for one of the simulated time series with distribution shift is shown in @fig-simulation-two-example.
@@
 
-```{r simulation_two_plot_joint, echo=FALSE}
-#| label: fig-simulation-two-joint
-#| fig-cap: "Mean interval width vs coverage error (top) and Mean Path Length vs.
-coverage error (bottom) for the second simulation study.
-The error bars represent the 10% to 90% quantiles of the metrics over the simulation datasets."
-#| fig-height: 6
-simulation_two_joint_plot(simulation_study2$results)
+```{r simulation_two_plot}
+#| label: fig-simulation-two-results
+#| fig-height: 6.5
+#| fig-cap: "Coverage error, mean interval width, path length, and strongly adaptive regret ($m = 20$) for $\\alpha = 0.8, 0.9, 0.95$ and simulations with and without distributional shift."
+
+simulation_two_plot(simulation_study2$results)
 ```
@@
 For this case study, we focus on national (US) one-week ahead forecasts of weighted influenza-like illness (wILI), which is a population-weighted percentage of doctors visits where patients presented with influenza-like symptoms [@biggerstaff2016flusight].
-The FluSight dataset, which is publicy available, include forecasts derived from 21 different forecasting models, from both mechanistic and statistical viewpoints [@tushar2018flusightnetwork, @tushar2019flusight].
+The FluSight dataset, which is publicly available, include forecasts derived from 21 different forecasting models, from both mechanistic and statistical viewpoints [@flusight2020;@tushar2018flusightnetwork;@tushar2019flusight].
 For our purposes, we treat the way the forecasts were produced as a black box.
@@
 The ACI algorithms target prediction intervals with coverage of $\alpha = 0.8$, $\alpha = 0.9$, and $\alpha = 0.95$.
-As in the simulation study, we used the interval constructor corresponding to the original presentaiton of each algorithm (see @tbl-aci).
+As in the simulation study, we used the interval constructor corresponding to the original presentation of each algorithm (see @tbl-aci).
 
 ```{r case_study, cache = TRUE}
-raw_data <- read_csv("https://raw.githubusercontent.com/FluSightNetwork/cdc-flusight-ensemble/master/scores/point_ests.csv", show_col_types = FALSE)
+# Paste together URL so it is not cut off in PDF
+url <- paste0("https://raw.githubusercontent.com/FluSightNetwork/",
+ "cdc-flusight-ensemble/master/scores/point_ests.csv")
+raw_data <- read_csv(url, show_col_types = FALSE)
 
@@
     Location == "US National", 
-    !(model_name %in% c("Delphi_Uniform", "CUBMA", "CU_EAKFC_SIRS", "CU_EKF_SEIRS", "CU_EKF_SIRS", "CU_RHF_SEIRS", "CU_RHF_SIRS"))
+    !(model_name %in% c("Delphi_Uniform", "CUBMA", "CU_EAKFC_SIRS", "CU_EKF_SEIRS", 
+                        "CU_EKF_SIRS", "CU_RHF_SEIRS", "CU_RHF_SIRS"))
   ) %>%
@@
   analysis_data, 
-  tibble(method = c("AgACI", "FACI", "SF-OGD", "SAOCP")), 
+  tibble(method = c("AgACI", "DtACI", "SF-OGD", "SAOCP")), 
   tibble(alpha = c(0.8, 0.9, 0.95))
@@
 
-The coverage errors, mean interval widths, and path lengths of the prediction intervals for each of the underlying forecast models is shown in @fig-case-study-metrics.
+The coverage errors, mean interval widths, path lengths, and strongly adaptive regret (for $m = 20$) of the prediction intervals for each of the underlying forecast models is shown in @fig-case-study-metrics.
 In all cases the absolute coverage error was less than $0.1$.
@@
 Interval widths were similar across methods, with SAOCP slightly shorter.
-Path Lengths were shorter for AgACI and FACI and longer for SAOCP.
+Path Lengths were shorter for AgACI and DtACI and longer for SAOCP.
 
@@
 #| label: fig-case-study-metrics
-#| fig-height: 6.5
-#| fig-cap: "Coverage errors, mean interval widths, and path lengths of prediction intervals generated with each ACI method based on forecasts from each of the 19 underlying influenza forecasting models."
+#| fig-height: 8.5
+#| fig-cap: "Coverage errors, mean interval widths, path lengths, and strongly adaptive regret (for $m = 20$) of prediction intervals generated with each ACI method based on forecasts from each of the 19 underlying influenza forecasting models."
 
@@
 The results of our simulations and case study show that, when tuning parameters are chosen well, Adaptive Conformal Inference algorithms yield well-performing prediction intervals.
-On the contrary, poor choice of tuning parameters can lead to intervals of low utility.
+On the contrary, poor choice of tuning parameters can lead to intervals of low utility: for one example, Figure [@fig-sf-ogd] shows how choosing the tuning parameter for SF-OGD to be too small can lead to intervals that update too slowly and significantly undercover.
 Furthermore, in some cases the prediction intervals may appear to perform well with respect to metrics like the empirical coverage error, while simultaneously being useless in practice.
@@
 It is also straightforward to choose this parameter based on a calibration set, although this strategy does not necessarily work well in cases of distribution shift.
-We also found that an advantage of the AgACI method is its robustness to the choice of its main tuning parameter, the set of candidate learning rates.
-Indeed, if AgACI does not perform well, one can simply increase the number of candidate learning rates.
+We also found that an advantage of the AgACI method is its robustness to the choice of its main tuning parameter, the set of candidate learning rates, in the sense that the grid of candidate learning rates can always be expanded as illustrated in @sec-agaci-tuning.
 
@@
 If the predictions are biased, for example, then the prediction intervals may be able to achieve optimal coverage at the expense of larger interval widths.
-Using ensemble methods to combine forecasts from several flexible machine learning models is one strategy that can be used to hedge against model misspecification and improve the quality of forecasts [@makridakis2020m4].
+This type of underperformance due to biased input predictions can be seen in the 2011-2012 flu season in the case study [@fig-case-study-example].
+One way bias can arise in the underlying predictions is due to model misspecification: for example, if a forecast method assumes a time series will evolve according to a particular parametric model that does not accurately capture the true data generating process, then the forecasts may be systematically biased.
+Using ensemble methods to combine forecasts from several flexible machine learning algorithms is one strategy that can be used to hedge against such model misspecification and improve the quality of forecasts [@makridakis2020m4].
+
+Overall, our findings illustrate strengths and weaknesses of all the considered algorithms.
+The original ACI algorithm is appealing in its simplicity, although its performance depends entirely on a good choice of its tuning parameter.
+AgACI tended to perform well in the simulation studies in terms of coverage error, although it had slightly higher strongly adaptive regret than other algorithms in some settings.
+However, there are relatively fewer theoretical guarantees available for AgACI than the other methods.
+DtACI, SF-OGD, and SAOCP all feature strong theoretical results, although they exhibited some differences in the simulation studies, with SF-OGD and SAOCP slightly undercovering in some scenarios.
+SAOCP also had longer path lengths than other methods in simulations, although in practice in the influenza forecasting task longer path lengths does not seem to effect the plausability of the prediction intervals the algorithm produces.
 
@@
 It may also be of interest to generate prediction intervals that have coverage guarantees for arbitrary subsets of observations (for example, we may seek prediction intervals for daily observations that have near optimal coverage for every day of the week, or month of the year), similar to guarantees provided by the MultiValid Prediction method described in [@bastani2022practical].
-Another avenue for theoretical research is to propose algorithms with provable bounds for the coverage and regret that do not depend on the outcome being bounded.
+Another avenue for theoretical research is to relax the assumption of bounded radii necessary for the theoretical results of algorithms such as SAOCP.
 
@@
 #| fig-cap: "Coverage errors, mean interval widths, and path lengths for the first simulation study with target coverage $\\alpha \\in \\{ 0.8, 0.9, 0.95 \\}$."
-#| fig-height: 9 
+#| fig-height: 9.5
 simulation_one_plot(simulation_study1$results)
@@
 
-```{r simulation_two_plot}
-#| label: fig-simulation-two-results
-#| fig-height: 9
-#| fig-cap: "Coverage error, mean interval width, and path length for $\\alpha = 0.8, 0.9, 0.95$ and simulations with and without distributional shift."
-
-simulation_two_plot(simulation_study2$results)
-```
+```{r simulation_two_plot_joint, echo=FALSE}
+#| label: fig-simulation-two-joint
+#| fig-cap: "Mean interval width vs coverage error (top) and Mean Path Length vs.
+coverage error (bottom) for the second simulation study.
+The error bars represent the 10% to 90% quantiles of the metrics over the simulation datasets."
+#| fig-height: 6
+simulation_two_joint_plot(simulation_study2$results)
+```

### 2024-06-13 helpers.R

@@ -1,4 +1,16 @@
+library(tidyverse)
+library(patchwork)
+
 label_alpha = label_bquote(cols = {alpha~"="~.(alpha)})
+
+running_example_data <- function(N) {
+  N <- 5e2
+  mu <- rep(0, N)
+  yhat <- mu
+  sigma <- ifelse(1:N <= N / 2, 0.2, 0.05)
+  y <- rnorm(N, mean = mu, sd = sigma)
+  list(y = y, yhat = yhat)
+}
 
 extract_metric <- function(results, metric) {
   unlist(lapply(results, function(result) result$metrics[[metric]]))
@@ -55,12 +67,22 @@ simulation_one_plot <- function(results) {
   plot_pathlength <- results %>%
     ggplot(aes(x = method, y = path_length, color = method)) +
     geom_boxplot() +
+    scale_y_log10() +
     facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
     labs(x = "", y = expression(PathLength(T)), subtitle = "Path Length") +
     aci_theme +
     theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
   
-  plot_coverr / plot_piwidth / plot_pathlength & theme(plot.margin = margin(0, 0, 0, 0, "cm"))
+  plot_sareg <- results %>%
+    ggplot(aes(x = method, y = strongly_adaptive_regret, color = method)) +
+    geom_boxplot() +
+    scale_y_log10() +
+    facet_grid(alpha ~ param, scale = "free_y", labeller = labels) +
+    labs(x = "", y = expression(SAReg(T, m)), subtitle = "Strongly Adaptive Regret") +
+    aci_theme +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
+  
+  plot_coverr / plot_piwidth / plot_pathlength / plot_sareg & theme(plot.margin = margin(0, 0, 0, 0, "cm"))
 }
 
 simulation_one_joint_plot <- function(results) {
@@ -101,8 +123,8 @@ simulation_two_plot <- function(results) {
     geom_hline(yintercept = 0, lty = 2, alpha = 0.5) +
     facet_grid(distribution_shift ~ alpha, labeller = labels) +
     labs(x = "", y = expression(CovErr(T)), subtitle = "Coverage Error", 
-         title = "Simulation Study: Distribution Shift") +
-    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
[…]

### 2024-06-13 paper.qmd
Fixed typo, updated lock file

@@
 DtACI, SF-OGD, and SAOCP all feature strong theoretical results, although they exhibited some differences in the simulation studies, with SF-OGD and SAOCP slightly undercovering in some scenarios.
-SAOCP also had longer path lengths than other methods in simulations, although in practice in the influenza forecasting task longer path lengths does not seem to effect the plausability of the prediction intervals the algorithm produces.
+SAOCP also had longer path lengths than other methods in simulations, although in practice in the influenza forecasting task longer path lengths does not seem to effect the plausibility of the prediction intervals the algorithm produces.
 
